### 2017-05-08 Publications/Journals & Articles/Concurrency and Computation Practice and Experience/Reply Letter/REPLY.tex
uploading previous publications
Signed-off-by: feliperodri <rms.felipe@gmail.com>

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
%\usepackage{mathtools}
\usepackage{enumerate}
\usepackage[nodayofweek]{datetime}

%\mathtoolsset{showonlyrefs}

\usepackage{xcolor}
%\input{rgbcolors}
%\usepackage{color}
\usepackage{natbib}
% \usepackage[style=authoryear-comp,dashed=false]{biblatex}
% \bibliography{piecewise_lyapunov}
\newcommand{\Real}{\mathbb{R}}

%\definecolor{DarkGreen}{rgb}{0.2, 0.4, 0.2}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{Manuscript Reference Number: CPE-16-0056}
\rhead{Page \thepage\ of \pageref{LastPage}}
\lfoot{}
\rfoot{}

%\pagestyle{headings}
\pagenumbering{roman}
\begin{document}
\setcounter{page}{1}
\thispagestyle{empty}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{logo_ufam.png}}
\end{minipage}}} \hspace*{0.5cm}
\begin{minipage}{11.0cm}{\sf \large \textbf{FEDERAL UNIVERSITY OF AMAZONAS}}\\
{\sc Department of Electronics and Computing}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Dr. Lucas C. Cordeiro\\
Federal University of Amazonas\\
Department of Electronics and Computing\\
Avenida General Rodrigo Ot\'avio, 6200, Coroado\\
69077-000 - Manaus - AM - BRAZIL\\
{\em e-mail: lucascordeiro@ufam.edu.br}
\end{minipage}
\end{flushright}

\vi \vi
\hspace*{\fill}{\small Manaus, June 2, 2016.}
\vi \vi

\hoffset -1.5cm \voffset .5cm

\begin{flushleft}
Dr. Alfredo Goldman\\
   Editor\\
   Concurrency and Computation: Practice and Experience
    \end{flushleft}
\vi

\begin{flushleft}
\textbf{REF.:} CPE-16-0056
    \end{flushleft}
\vi \vi

\indent Dear Dr. Goldman,
\vi

We thank you for asking us to comply with the Referees' reports. We are submitting the revised version of our paper entitled as ``{\em SMT-Based
Context-Bounded Model Checking for CUDA Programs}''  by Phillipe Pereira, Higo Albuquerque, Isabela da Silva, Hendrio Marques, Felipe Monteiro, Ricardo dos Santos, and Lucas Cordeiro.

The details of all changes made to our paper and also our response to each point raised by the reviewers have been included at the end of this letter.

%\vi

%Thank you very much in advance.

\vi\vi

\indent
~~~~~~~~~~~~~~~~~~~~~~~~~~~Sincerely,\\

\begin{quote}
\begin{quote}
\begin{flushright}

\vi
\vi

Lucas C. Cordeiro~~~~~~~
\end{flushright}
\end{quote}
\end{quote}

\hoffset -1.5cm \voffset .5cm

% =============================================
\newpage
\subsection*{Authors'   Response to the Review Comments on manuscript ``{SMT-Based Context-Bounded Model Checking for CUDA Programs}'' -- Manuscript Reference Number: CPE--16--0056}

\vi

The authors thank the anonymous reviewers for their valuable and constructive comments, which helped us improve significantly the quality of our manuscript.

\vi

According to the reviewers comments, our manuscript has been carefully revised and the detailed changes are highlighted in blue color along the original text. Our responses to all comments, which were raised by the reviewers, are given in the sequel using the blue color.

\vi

We hope that these revisions in the manuscript and our responses will be sufficient to make our manuscript suitable for publication in the Journal of Concurrency and Computation: Practice and Experience.

%\subsection*{Editor-in-Chief comments:}

%\begin{quote}
%The reviewers generally agree that the paper contains some potentially publishable results. They also mention a number of technical issues. The paper should be carefully revised by well addressing all these issues mentioned by the reviewers.
%%The review of this paper is now complete, and I am including the Reviewers' reports (directly at the end of this email and sometimes as attachments).  As you can see, one Reviewer recommends rejection, and the others recommend possible reconsideration after minor revisions. If you are able to revise the manuscript thoroughly in response to the Reviewers' comments, I believe that it would be suitable for second review.  I should add that, in view of the strong criticisms, anything but a thorough response and revision would just be an unfortunate waste of time.
%\end{quote}

%\subsection*{Answer to the Editor-in-Chief comments:}

%Thanks for the AE's   comments. Following six reviewers'  comments and
%suggestions, the manuscript has been significantly revised and rewritten to deal with the technical issues arose by reviewers.
%Following the suggestions of the reviewers, the results have been extended to handle a more general class of agents, namely, we have extended the model of the agent from high-order integrator dynamics to any linear or linearizable system. This has been made possible by the explicit need to provide results in a more general practical application, as was requested by the reviewers. Nevertheless, the methodology proposed still remains the same.
%For more details, refer to the  replies to reviewers'  comments and the revised manuscript.

% =============================================
%\newpage

\vspace*{1cm}

%\subsection*{Associate Editor Comments}

%It was not that easy to make a final decision for this article. Reading the reviews, it becomes clear that the reviewers were not at all happy with the article, so that rejecting the article would be justified. However, a lot of the criticism is because of the presentation, and the unclarity of the added values, while the experimental results attracted the interest of all reviewers. I therefore suggest a major revision and hope that the authors can improve the presentation of their article based on the given reviews.

%The first reviewer recommended a rejection of the article. The main arguments for that recommendation are lack of detail of the verification method, and therefore the impossibility of checking its correctness by a reviewer. Moreover, there was some criticism on the presentation and references made in the article.

%The second reviewer recommends a minor revision, but when reading the review, I found some hard critiques: It seems to me that the reviewer was impressed by the experimental results, but has not seen new verification methods in the article. The main contribution seen by that reviewer is to deal with modeling uncertainty and the effect of finite precision arithmetic in one tool that uses BMC as the core verification techniques and its interesting experimental results.

%The third reviewer recommends a major revision, and complains about unclarity of the technical contributions of the paper and the absence of technical details like the first reviewer.

%The fourth reviewer recommends a major revision, and explains in detail what should be changed with the article. The role of bounded model checking seems also to be pretty unclear.

%\subsection*{Response to Associate Editor}

% =============================================
%\newpage

\vspace*{1cm}

\vi
\subsection*{Reviewer \#1 comments followed by the authors' answers:}

\begin{quote}

{\bf Comments to the Authors}

This paper extends the ESBMC bounded model checker to support CUDA constructs and simulate the execution of CUDA programs in the CPU. Specifically, it models CUDA APIs using usual CPU functions that can be processed by the existing ESBMC checker, and turns GPU threads into CPU threads that are examined and optimized by the ESBMC checker.

This paper's main contributions include: (a) a set of CUDA API models based on CUDA's semantics, (b) the application of a partial order reduction algorithm to curb the thread interleaving explosion issue. In addition, the authors performed a quite comprehensive comparison of their tool ESBMC-GPU with other existing GPU verifiers and checkers.

It is interesting to see how to extend a Pthread-based bounded model checker to handing a different parallel programming model like CUDA. In general, the extension needs to address two main challenges:

  (A) model the data structures and the APIs in the new programming model;
  (B) resolve the difference between PThread and CUDA (i.e. handle the inherent characteristics of how threads are executed and synchronized).

It seems to me that the authors did a good job in (A), e.g. the model of dim3, cudaMalloc, cudaMemcpy, and cudaFree. The model is written to cater for the need of the ESBMC checker. With the model, ESBMC can check CUDA programs like Pthread programs.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We really appreciate the reviewer's summary about our work, which fairly describes the main contributions of our manuscript.}
\end{quote}

{\bf Reviews:}

{\bf (1)} However, the second challenge is handled less swiftly. In the proposed method, each CUDA thread is mapped to a Pthread (in ESBMC). This is fine since the official CUDA simulator uses a similar approach. The main problem is on thread synchronization. CUDA uses lock-step execution within warps (as mentioned in the paper on Page 12), and barrier synchronization within blocks. In contrast, Pthread uses schemes such as mutexes, wait-notification, etc. Bridging the gap between these two different synchronization schemes is tricky. Unfortunately I didn't see how the authors address this gap carefully. They should give some formal descriptions and examples to illustrate that ESBMC-GPU can preserve the CUDA's synchronization scheme precisely.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Actually, each CUDA thread is mapped to a thread using the internal representation of ESBMC (via Pthread library). As pointed out by the reviewer, the thread representation of ESBMC follows a similar approach to the official CUDA simulator. One feature of CUDA related to thread synchronization supported by ESBMC-GPU is the \emph{\tt \_\_syncthreads()} intrinsic function, which specifies synchronization points in the GPU kernel. Such alternative guarantees that all threads in a certain block wait before any of them is allowed to proceed. As described in Algorithm $7$, we implement this feature in our operational model using Pthread functions. First of all, a $count$ global variable is initialized to control how many times the \emph{\tt syncthreads} function is actually called. In its first call, a mutex is initialized and a condition variable is set up (steps from $2$ to $5$ of Algorithm $7$). Then, on each subsequent call, the $count$ variable is incremented (step $6$) and the calling thread is put on hold until it receives a signal (step $8$). When all GPU threads are on hold, the $pthread\_cond\_signal$ function sends a signal to all threads waiting on a specific condition variable (step $9$) and the mutex is then unlocked (step $10$). We address this important discussion in Section 3.2.1. about ``Thread Synchronization'' by describing the \emph{\tt syncthreads} algorithm.}
\end{quote}

{\bf (2)} The proposed Monotonic Partial Order Reduction (MPOR) method deserves more investigation and details. First of all, an existing GPU checker GKLEE checks whether the threads access the same memory location and uses this information to execute only one thread interleaving called canonical schedule. It won't incur any unnecessary thread interleavings, while the proposed MPOR may still explore these interleavings. In other words, GKLEE has implemented, with respect to the CUDA model, what MPOR try to achieve. So do other GPU verifiers like GPUVerify and PUG (which explore one thread interleaving of two symbolic threads).

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have now considerably improved the MPOR section by describing its reduction and the main differences between our technique and GKLEE canonical scheduling, as pointed out by the reviewer (see Section 3.3.1. about ``Differences from MPOR to other interleaving reduction methods''). There are some sublte differences between both approaches, as described in Sections 3.3.1 and 2.2, where we also compare to GPUVerify and PUG.}
\end{quote}

{\bf (3)} Second, it is unclear how new the MPOR is. It seems to me that the authors directly reuse the MPOR algorithm already included in ESBMC [23].

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have highlighted our improvements to the MPOR algorithm implemented in ESBMC. In particular, we have modified the MPOR algorithm to identify transitions between threads that accessed different memory locations in the same array, where the previous MPOR implementation of ESBMC was unable to handle it. In addition, we have improved the description about the MPOR reduction process in Section 3.3 by adding (1) more details about the algorithm steps and (2) some examples of the algorithm's execution.}
\end{quote}

{\bf (4)} Third, while GKLEE and other existing tools explores only one thread interleaving, it is unclear how many interleavings are explored by ESBMC+MPOR although the authors show that the performance is boosted significantly with the MPOR optimization.

\begin{quote}
\textcolor{blue}{\textbf{Response:} MPOR techniques explores all (possible) interleavings that produce new program states; it builds an SMT formula, for a given interleaving, only when the MPOR analysis is concluded to avoid checking redundant states. Thus, the verification time is directly affected by the size of this formula and its satisfiability checking by a particular SMT solver. In addition, it is worth notice that GKLEE explores schedules of full thread execution in a barrier interval without interleaving statements, while ESBMC-GPU employs BMC and Pthread using MPOR.}
\end{quote}

{\bf (5)} The experimental result part is perhaps the most valuable one in this paper. Experimental results show that ESBMC-GPU is able to detect more properties violations than other existing GPU verifiers, while keeping lower rates of false results. This indicates that the proposed approach merits being published. However this part can be improved as well. Particularly, the authors of this paper should contact the authors of other tools to make sure (1) the right configurations are used; (2) the assumptions for these tools to run properly are satisfied; (3) confirm that the findings (e.g. the false positives and false negatives) are real. For example, Page 15 mentions "GKLEE has 24 benchmarks that were not supported, which are due to the use of the memset function, which is a specific function of the C/C++ libraries".
In fact GKLEE supports most of the C/C++ library functions; but the option ``-libc'' may need to be set properly so as to link the definitions of the C/C++ library functions including memset. With this, GKLEE supports all the benchmarks while the proposed ESBMC-GPU doesn't support 23.

\begin{quote}
\textcolor{blue}{\textbf{Response:} In our (first) submission to CPE, we had previously contacted all authors of the GPU verifiers that we used in our experimental evaluation. However, only the developers of GPUVerify (Dr. Alastair Donaldson) and CIVL (Dr. Stephen Siegel) have replied to our email (we still keep those exchanged emails). The developers of GKLEE and PUG did not reply to our email, which thus makes it difficult to set the most appropriate configuration parameters for those tool. We thank the reviewer for point out the right configuration for GKLEE. We have now re-run GKLEE using the option ``-libc'', as recommended by the reviewer, and then updated the experimental results accordingly. However, even running GKLEE with the option ``-libc'', GKLEE is unable to verify all benchmarks, as reported in Section 4.3. about the ``Experimental Results''. Additionally, all verification results were carefully analyzed by GPU experts in order to confirm the encountered false positives and false negatives.}
\end{quote}

{\bf (6)} Finally, the presentation of the paper can be improved. Despite including the new materials, it is still in a workshop format. Specifically, more space should be given to the CUDA synchronization modeling and the detailed description of MPOR.

\begin{quote}
\textcolor{blue}{\textbf{Response:} According to all reviewers' comments, we believe we have considerably improved the presentation of our manuscript by adding more details about the ESBMC tool architecture (including the link between CUDA operational model and ESBMC), the CUDA operational model (and its organization), the MPOR implementation (including the one-schedule mechanism), and the experimental evaluation (using the most appropriate configuration parameters and benchmarks description). Additionally, we have substantially expanded the number of references in our manuscript to fairly describe the most up-do-date related studies; in particular, we have added approximately $23$ new references to our manuscript, which substantially increased the total number of references to $46$.}
\end{quote}

{\bf (7)} strength

\begin{itemize}
\item Reuse the existing ESBMC to perform bounded model checking on mutli-thread programs
\item Provide a set of CUDA API models for ESBMC
\item Evaluate extensively the tool and compare it with other existing tools
\item Achieve good evaluation results
\end{itemize}

\begin{quote}
\textcolor{blue}{\textbf{Response:} We appreciate the reviewer's effort to highlight the strengths of our manuscript.}
\end{quote}

{\bf (8)} weakness:

\begin{itemize}
\item Do not describe how many CUDA APIs are modeled and how other important APIs are modeled

\begin{quote}
\textcolor{blue}{\textbf{Response:} We support most CUDA Driver and CUDA Runtime APIs; in particular, ESBMC-GPU supports the CUDA call kernel (\textit{device\_launch\_parameters} and \textit{vector\_types}), driver\_types (\textit{math\_functions}, \textit{cuda\_runtime\_api}, and \textit{host\_definitions}), sm\_atomic\_functions (\textit{vector\_types}), and curand\_kernel (curand). We address this important discussion in Section 3.1. about ``Operational Models for CUDA Libraries''.}
\end{quote}

\item Do not address the thread synchronization scheme in CUDA
\item Do not elaborate the MPOR mechanism
\item Lack of awareness of existing tools' one-schedule mechanism (that beats MPOR)
\item Inconclusive experimental comparisons with existing tools
\end{itemize}

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have tackled each point, as aforementioned in the previous questions (from 1 to 6).}
\end{quote}

\end{quote}

%--------------------------------------------------------------------

\newpage

\subsection*{Reviewer \#2 comments followed by the authors' answers:}
\begin{quote}

{\bf Comments to the Authors}

In this paper the authors present ESBMC-GPU, which is an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC) tool aimed at verifying SMT programs written for the CUDA framework. This paper is an extension of a previous paper published in a conference, and the authors claim that the version of ESBMC-GPU presented in this paper has been optimized and extended when compared to previous versions. All in all, the paper is well written and the results presented are consistent and sound. Yet, the detailed description of the tool and the writing of the paper can be improved to better reflect the optimizations presented in this work. Some comments and suggestions are below:

\begin{quote}
\textcolor{blue}{\textbf{Response:} We appreciate the reviewer's comments. We have carefully revised our manuscript considering all valuable and constructive comments/suggestions.}
\end{quote}

{\bf Reviews:}

{\bf(1)} The abstract can be considerably improved. The contributions of this paper should be highlighted at the abstract.

\begin{quote}
\textcolor{blue}{\textbf{Response:} The abstract was substantially revised, in order to give more emphasis on the main contributions of our manuscript. Additionally, we have now included more details about the experimental results.}
\end{quote}

{\bf(2)} The authors should mention, in the Introduction, that the ESBMC-GPU tool is available for download. The authors mention this possibility in a footnote at section 4.1, but it is important to highlight this possibility at the Introduction.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have now highlighted in the introduction section that readers (or users) can download our tool and all benchmarks from http://esbmc-gpu.org/ webpage.}
\end{quote}

{\bf(3)} At subsection 2.1, the description of ESBMC should be augmented and improved. Only to illustrate this point, the authors do not even clarify what the acronym ?VC? means.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have augmented and considerably improved the description of ESBMC according to the reviewer comments. In particular, we have defined the acronym VC and its interpretation; we have also described in more details the (most up-to-date) ESBMC software architecture and the link between CUDA operational model and ESBMC.}
\end{quote}

{\bf(4)} It is important, to improve readability, to include a figure presenting ESBMC-GPU architecture. It is also important to use this figure to improve the discussion on architectural differences to the original ESBMC tool.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have added a new subsection (2.3.1) to explain in more details about the ESBMC's architecture. In addition, we have now added two new images to improve readability of our manuscript: a sketch of the ESBMC's architecture and a draft that explains the connection of CUDA operational model into the verification process, which is performed by ESBMC. We have also included the CUDA operational model organization to emphasize which CUDA libraries are actually supported by ESBMC-GPU.}
\end{quote}

{\bf(5)} The reference section can be considerably improved. For instance, seminal works on ?Bounded Model Checking? and ?Satisfiability Module Theories? (section 1) should be referenced. The authors should also include references for the SMT solvers listed in subsection 4.2.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have now included the following references for Bounded Model Checking, Satisfiability Modulo Theories, and the respective SMT solvers that we currently cite in the paper:} 
\end{quote}
\textcolor{blue}{A. Biere. Bounded Model Checking. Handbook of Satisfiability. IOS Press 2009; 457--481.}

\textcolor{blue}{Barrett C, Sebastiani R, Seshia S, Tinelli C. Satisfiability Modulo Theories. Handbook of Satisfiability. IOS Press 2009; 825--885.}

\textcolor{blue}{Moura L, Bjorner N. Z3: An Efficient SMT Solver. TACAS 2008; 337--340.}

\textcolor{blue}{Brummayer R, Biere A. Boolector: An Efficient SMT Solver for Bit-Vectors and Arrays. TACAS 2009; 174--177.}

\textcolor{blue}{B. Dutertre. Yices 2.2. CAV 2014; 737--744.}

\textcolor{blue}{Cimatti A, Griggio A, Schaafsma B, Sebastiani R. The MathSAT5 SMT Solver. TACAS 2013; 93--107.}

\textcolor{blue}{Deters M, Reynolds A, King T, Barrett C, Tinelli C. A tour of CVC4: How it works, and how to use it. FMCAD 2014; 7.}

{\bf(6)} The legends for tables I e II should explain what the numbers in the tables mean.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have now thoroughly explained in each legend the data from Tables I and II.}
\end{quote}

\end{quote}

%---------------------------------------------------------------------------

\newpage
\subsection*{Reviewer \#3 comments followed by the authors' answers:}
\begin{quote}

{\bf Comments to the Author}

The paper proposes a context-bounded model checking for CUDA programs. In particular,
the paper leverages on the advances of satisfiability modulo theory (SMT). Compared
to previously proposed approaches, this work could verify more properties in CUDA
programs chosen from a standard pool of benchmarks. It also shows the application of
monotonic partial order reduction (MPOR) in verifying GPU programs. 
The paper is oriented towards a tool. However, such publications are encouraged for the journal. 
Specifically, the work shows the application of several well-known
model checking techniques, such as SAT-based and explicit state-space based model
checking, for CUDA programs. The paper does not propose any novel technique. However,
it has designed a tool, which is evaluated well and it has also been made available
in public. I appreciate such effort by authors.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We appreciate the reviewer's comments. We have emphasized four major contributions of our manuscript in Section 1. We agree with the reviewer that SMT-based context-bounded model checking is not a novel technique (it has been around for nearly one decade), but we have not seen yet its application to verify CUDA programs in the literature. Additionally, we have extended the original MPOR algorithm implemented in ESBMC, in order to identify array accesses that are independent in CUDA programs (which is very common in GPU applications); this extension has led to significant performance improvements in ESBMC-GPU, as described in the experimental evaluation of our manuscript.}
\end{quote}

{\bf Reviews:}

{\bf(1)} The organization of the paper is somewhat very informal. The current presentation
mostly revolves around showing code snippets. However, there is a lack of formal
presentation, especially for a paper that discusses model checking and verification.

\begin{quote}
\textcolor{blue}{\textbf{Response:} In order to improve the organization of our manuscript, we have tackled its informality by replacing code snippets by (detailed) algorithms that reliably represent the CUDA operational model. We believe this alternative form of presentation has also improved the readability of our manuscript. Note that we have described pre- and postconditions of the algorithms, where possible, in order to ensure the correct operation of the CUDA libraries' real behavior and also to consequently improve the manuscript's readability and trustfulness. These pre- and postconditions formally ensure the CUDA libraries semantics by means of assertions properly inserted into the code.}
\end{quote}

{\bf(2)} It would be good to highlight, very concretely in a separate section, the reasons
for the tool being more effective compared to any other existing verifiers for
CUDA programs. Is it due to the operational model being rich for ESBMC-GPU? Is it
due to the encoding via SMT solvers? I would suggest to take a simple example of
CUDA programs and show (with sufficient detail) the reason for ESBMC-GPU being
more effective than other existing tools. Whereas the evaluation is promising, but
we need a general idea of the reason for such positive evaluation.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have included a new subsection at the end of Section 4.3. about the ``Experimental Results'', which properly describes the ESBMC-GPU effectiveness over other existing GPU verifiers. Additionally, we describe a simple example in Section 3.1.3. ``Illustrative Example'', which shows the reason for ESBMC-GPU being more effective than other existing verification tools.} 
\end{quote}

{\bf(3)} Authors use ``pthread\_create'' to use different threads for CUDA programs. Whereas
CUDA threads are different in the sense that they operate in lockstep fashion
within a warp (i.e. the SIMD feature of GPU). It is unclear whether warp-level
threads in GPU can be simulated by ``pthread''. Authors should provide sufficient
justification that such a transformation via pthread does not affect the soundness
of the verifier.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Note that in our present approach, each CUDA thread is actually mapped to a thread using the internal representation of ESBMC (via Pthread library). In ESBMC, a thread $t$ is a sublist of commands between \emph{begin\_thread} and \emph{end\_thread}, which represent just scoping constructs and do not contribute to the expansion of the RT (described in Section 2.3). Threads are created via asynchronous procedure calls (\emph{start\_thread}), which return an integer that can be used as thread identifier for synchronization (\emph{join\_thread}); hence, dynamic thread creation is allowed. The thread representation of ESBMC follows a similar approach to the official CUDA scheduler. 
One feature of CUDA related to thread synchronization supported by ESBMC-GPU is the \emph{\tt \_\_syncthreads()} intrinsic function, which specifies synchronization points in the GPU kernel. Such alternative guarantees that all threads in a certain block wait before any of them is allowed to proceed. As described in Algorithm 7, we implement this feature in our operational model using Pthread functions. First of all, a $count$ global variable is initialized to control how many times the \emph{\tt syncthreads} function is actually called. In its first call, a mutex is initialized and a condition variable is set up (steps from $2$ to $5$ of Algorithm $7$). Then, on each subsequent call, the $count$ variable is incremented (step $6$) and the calling thread is put on hold until it receives a signal (step $8$). When all GPU threads are on hold, the $pthread\_cond\_signal$ function sends a signal to all threads waiting on a specific condition variable (step $9$) and the mutex is then unlocked (step $10$). We address this important discussion in Section 3.2.1. about ``Thread Synchronization''. }
\end{quote}

{\bf(4)} Since the approach described in the paper is quite informal, I could not see
any proof regarding the soundness of the operational model. I do not think it
is straightforward. Authors could try to provide some sketches of proof for
the soundness.

\begin{quote}
\textcolor{blue}{\textbf{Response:} The usefulness of our approach relies on the fact that COM correctly represents the original CUDA libraries. In that sense, all developed COM modules were manually verified and exhaustively compared to the original ones to ensure the same behavior; they all contain precondition and postcondition, where possible, to ensure that a predicate holds before and after the execution of a given function, respectively. Additionally, although COM is a new implementation, it consists in (reliably) constructing a simplified model of the related CUDA libraries, using the same language and by means of the original code and documentation (pre- and postconditions are formally verified by the model checker using assertions within the code itself), which thus tends to decrease the resulting number of errors. Additionally, the behavior of the CUDA libraries functions are represented in C/C$++$ programming languages using native functions, which are already supported by ESBMC ({\it e.g.}, \textit{malloc}, \textit{free}, \textit{assert}). The soundness proof for those native functions implemented in ESBMC can be found in Cordeiro {\it et al.}. Conformance testing regarding operational models is commonly employed in practice; however, that option is not available in the present case, albeit it is an interesting possibility for future work. We discuss this important issue raised by the reviewer, at the end of Section 3.1. ``Operational Models for CUDA Libraries'' and Section 5. ``Conclusions''.}
\end{quote}

{\bf(5)} Authors should provide some summary of these benchmarks in the paper. I know
that the benchmarks are available in the website. However, providing some
salient features of these benchmarks in the paper will help improve the quality.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Note that we used benchmarks from a variety of sources to evaluate ESBMC-GPU's precision and performance, which include CUDA benchmark suites to check basic functions that are commonly used by CUDA applications. Our benchmarks test suite comprises of 20 CUDA kernels from NVIDIA GPU Computing SDK v2.0, 20 CUDA kernels from Microsoft C++ AMP Sample Projects, and 104 CUDA programs that explore a wide range of CUDA functionalities. Although some of these CUDA benchmarks do not implement any real CUDA application, we used them to analyze to which extent ESBMC-GPU is able to handle and detect errors; in particular, those benchmarks have been previously used for comparing the performance and precision of different GPU verifiers. It is worth notice that the benchmarks related to OpenCL were not applied to our experimental evaluation, once ESBMC-GPU does not support it yet. We briefly describe the benchmarks in the first two paragraphs of Section 4.1. about the ``Experimental Setup''.}
\end{quote}

{\bf(6)} A minor comment: Why do we even need Figure 11? I think it's trivial, unless it
is expanded enough with some details of two-thread analysis.

\begin{quote}
\textcolor{blue}{\textbf{Response:} We have modified this specific figure and included more details about the two-thread analysis, as suggested by the reviewer.}
\end{quote}

\end{quote}
%-----------------------------------------------------------------

\label{LastPage}

% \printbibliography

\end{document}

### 2017-05-08 Publications/Journals & Articles/Concurrency and Computation Practice and Experience/cpedoc.tex
uploading previous publications
Signed-off-by: feliperodri <rms.felipe@gmail.com>

% cpedoc.tex V2.0, 13 May 2010

\documentclass[times, doublespace]{cpeauth}

\usepackage{color}
\usepackage{listings}
\usepackage{multirow}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[caption=false]{subfig}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algpseudocode}
\algdef{SE}[FUNCTION]{Function}{EndFunction}
[2]{\algorithmicfunction\ \textproc{#1}\ifthenelse{\equal{#2}{}}{}{(#2)}}%
   {\algorithmicend\ \algorithmicfunction}%

\lstset{language=C,basicstyle=\small}
\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
\lstset{tabsize=2}
\lstset{firstnumber=1}
\lstset{frame=single}
\lstset{
  language={C++},
  morekeywords={define,public,class,this,assert,exit,assume,then,end-if,end,input,output,true,false,bool,malloc,using,namespace,std,cudaError_t
		}
}

\hyphenation{
GKLEE 
property 
operational 
developed
ap-proxi-mates
pro-grams
inter-face
errors
par-ticu-lar
ver-ifiers
veri-fi-ca-tion
ori-gin-al
hier-archy
last-Error
in-de-pend-ent
pos-itions
array
ana-lysis
unsyn-chron-ized
values
total
ex-plor-ation
add-ition-al-ly
}

\usepackage{moreverb}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2016}

\begin{document}

\runningheads{P.~A.~Pereira et al.}{SMT-Based Context-Bounded Model Checking for CUDA Programs}

\title{SMT-Based Context-Bounded Model Checking\\for CUDA Programs}

\author{P.~A.~Pereira\affil{1}, H.~F.~Albuquerque\affil{1}, I.~S.~da Silva\affil{1}, H.~M.~Marques\affil{1}, F.~R.~Monteiro\affil{1}, R.~dos S.~	Ferreira\affil{2}, and L.~C.~Cordeiro\affil{1}\corrauth}

\address{
\affilnum{1}Faculty of Technology, Federal University of Amazonas, Amazonas, Brazil\break
\affilnum{2}Department of Informatics, Federal University of Vi\c{c}osa, Minas Gerais, Brazil}

\corraddr{3000 Av. Rodrigo Ot{\'a}vio, 69077--000, Electronic and Information Research Center at Federal University of Amazonas, Manaus, Amazonas, Brazil.\\E-mail: lucascordeiro@ufam.edu.br}

\begin{abstract}
We present ESBMC-GPU tool, an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC), which is aimed at verifying Graphics Processing Unit (GPU) programs written for the Compute Unified Device Architecture (CUDA) platform. ESBMC-GPU uses an operational model, {\it i.e.}, an abstract representation of the standard CUDA libraries, which conservatively approximates their semantics, in order to verify CUDA-based programs. It then explicitly explores the possible interleavings (up to the given context bound), while treats each interleaving itself symbolically. Additionally, ESBMC-GPU employs the monotonic partial order reduction and the two-thread analysis to prune the state space exploration. Experimental results show that ESBMC-GPU can successfully verify $82\%$ of all benchmarks, while keeping lower rates of false results. Going further than previous attempts, ESBMC-GPU is able to detect more properties violations than other existing GPU verifiers due to its ability to verify errors of the program execution flow and to detect array out-of-bounds and data race violations.
\end{abstract}

\keywords{model checking; GPU; CUDA programs; formal verification;}

\maketitle

\vspace{-6pt}
%------------------------------
\section{Introduction}
\vspace{-2pt}
%------------------------------

{\bf C}ompute {\bf U}nified {\bf D}evice {\bf A}rchitecture (CUDA) is a parallel computing platform and {\bf A}pplication {\bf P}rogramming {\bf I}nterface (API) model created by NVIDIA~\cite{cuda:2012}, which extends C/C$++$ and Fortran, in order to create a computational model that aims to harness the computational power of {\bf G}raphical {\bf P}rocessing {\bf U}nits (GPUs)~\cite{gpu:2010}.\ As in other programming languages, errors in CUDA programs eventually occur, in particular, due to array out-of-bounds, arithmetic overflow, and division by zero violations.\ Additionally, since CUDA is a platform that deals with parallel programming, specific concurrency errors related to data race and barrier divergence can be exposed due to the non-deterministic behaviour of the threads interleavings~\cite{betts:2012}.

Here, we describe and evaluate an approach for verifying CUDA programs based on the {\bf E}fficient {\bf S}MT-Based Context-{\bf B}ounded {\bf M}odel {\bf C}hecker (ESBMC)~\cite{cordeiro:2011,cordeiro:2012,ramalho:2013}, named as ESBMC-GPU, using a CUDA operational model (COM), which is an abstract representation of the standard CUDA libraries ({\it i.e.}, the native API) that conservatively approximates their semantics. We describe the implementation of COM, its pre- and postconditions, and simulation features ({\it e.g.}, how elements are manipulated), and how these are applied to verify CUDA applications. In contrast to previous attempts~\cite{betts:2012,Li:2010,Li:2012,civl:2015}, we combine symbolic model checking, based on {\bf B}ounded {\bf M}odel {\bf C}hecking (BMC)~\cite{Biere:2009} and {\bf S}atisfiability {\bf M}odulo {\bf T}heories (SMT)~\cite{Barrett:2009} techniques, with explicit state-space exploration, similar to Cordeiro {\it et al.}~\cite{cordeiro:2011}. In particular, we explicitly explore the possible interleavings (up to the given context bound), while we treat each interleaving itself symbolically w.r.t. a given property.

To prune the state-space exploration, we apply {\bf M}onotonic {\bf P}artial {\bf O}rder {\bf R}eduction (MPOR)~\cite{KahlonWG09} to CUDA programs, which eliminates redundant interleavings without missing any behavior that can be exhibited by the program. We have modified the MPOR algorithm to identify transitions between threads that accessed different memory locations in the same array. Since CUDA kernels typically produce regular and independent access to explore the benefits of the GPU execution model, the application of MPOR routinely leads to substantial performance improvements in most benchmarks. Thus, using operational models that simulate CUDA libraries, together with MPOR implementation in ESBMC-GPU, we achieve significant (correct) results of CUDA kernels verification, primarily when compared to other state-of-the-art GPU verifiers~\cite{betts:2012,Li:2012,Li:2010,civl:2015}. Additionally, the present approach considers low-level aspects related to dynamic memory allocation, data transfer, memory deallocation, and overflow. Such violations are typically present in CUDA programs, however, they are routinely ignored by most GPU verifiers. Thus, we provide a more precise verification than other existing approaches, considering soundness of data passed by the main program to the kernel, with the drawback of leading to a higher verification time.

%---------------------------------
\subsection{Contributions}
\vspace{-2pt}
\label{sec:contributions}
%---------------------------------

We make four major contributions:
\vspace{-4pt}
\begin{itemize}
\item we extend benefits of SMT-based context-bounded model checking for CUDA programs, in the context of parallel programming for GPUs, to detect more failures than other existing approaches, while keeping lower rates of false results; although SMT-based context-bounded model checking is not a novel technique, we have not seen in the literature its application to verify CUDA programs.

\item this work marks the first application of MPOR to CUDA programs to identify array accesses, which are independent, leading to significant performance improvements.

\item we provide an effective and efficient tool implementation (ESBMC-GPU) to support the checking of several CUDA programs. ESBMC-GPU tool and all benchmarks used during the evaluation process are available at \url{http://esbmc-gpu.org}.

\item we provide an extensive experimental evaluation of our approach against GKLEE~\cite{Li:2012}, GPUVerify~\cite{betts:2012}, PUG~\cite{Li:2010}, and CIVL~\cite{civl:2015} using standard CUDA benchmarks, which are extracted from the literature~\cite{cuda:2012,betts:2012,cudaproguide:2015}. Experimental results also show that our present approach outperforms all existing GPU verifiers with respect to the number of correct results.
\end{itemize}

This paper extends our previous work~\cite{Pereira15,Pereira16}. The version of ESBMC-GPU described and evaluated here has been optimized and extended. It now expands our operational model to support more CUDA libraries ({\it e.g.}, math functions) and includes new SMT solvers as back-end for ESBMC-GPU ({\it e.g.}, Boolector, Yices, MathSAT, and CVC4). We also provide additional details about our operational model, the application of MPOR to CUDA programs, and the two-thread analysis. We have also significantly expanded the experimental basis, including a new GPU verifier (CIVL), and evaluate ESBMC-GPU with (respectively, compare it against) the most recent stable versions of the GPU verifiers.

%---------------------------------
\subsection{Organization of this work}
\vspace{-2pt}
\label{sec:organization}
%---------------------------------

The remainder of the paper is organized as follows: In Section~\ref{sec:fundamentacao_teorica}, we first give a brief introduction to the CUDA programming language, existing verifiers for GPU, and the ESBMC verifier. In Section~\ref{sec:methodology}, we describe our operational models for CUDA libraries, the application of MPOR to CUDA programs, and our two-threads analysis. In Section~\ref{sec:resultadosExperimentais}, we present the results of our experiments using several publicly available benchmarks. In Section~\ref{sec:conclusions}, we conclude and describe future work.

%---------------------------------
\vspace{-6pt}
\section{Preliminaries}
\vspace{-2pt}
\label{sec:fundamentacao_teorica}
%---------------------------------

%---------------------------------
\subsection{CUDA Programming Language}
\vspace{-2pt}
\label{ssec:linguagem_cuda}
%---------------------------------

CUDA is a general-purpose parallel computing platform (developed by NVIDIA) to represent a programming model for GPUs~\cite{cuda:2012,cudaproguide:2015}. In the CUDA programming model, the kernel concept is used for a function that runs $n$ copies concurrently in the GPU, where $n$ is the product between the number of blocks and threads. A kernel is defined by a \emph{\tt \_\_global\_\_} specifier and uses the notation \emph{\tt kernel$<<<$ B,T $>>>$}, where $B$ and $T$ are the number of blocks and threads per block, respectively. Each kernel runs in the GPU as thread and each thread receives an unique identifier (ID), which is formed by thread and block numbers. The thread ID is used to index its tasks ({\it i.e.}, memory positions and cooperation); and threads are typically organized by blocks in the GPU.
Inside a block, the thread hierarchy is defined by a variable called \emph{\tt threadIdx}. This variable is a vector of three components, which allows the use of uni-, two-, and three-dimensional indexes~\cite{cudaproguide:2015}.

Blocks can also be defined in three dimensions, where each dimension can be accessed by the \emph{\tt blockIdx} variable. This variable is also composed by three components that allow CUDA programs to use uni-, two-, and three-dimensional blocks.
The maximum number of threads per block depends on the hardware platform, but it usually ranges from $1024$ to $2048$~\cite{cudaproguide:2015}. In addition, blocks have a feature that permits them to be executed in any order; and they can also be allocated in any processor. As a result, a kernel may also be executed by multiple blocks, and the total number of threads represents the number of blocks multiplied by the number of threads per block.

In the CUDA programming model, the GPU is refereed as a \textit{device} and the \textbf{C}entral \textbf{P}rocessing \textbf{U}nit (CPU) is refereed as a \textit{host}. $\_\_device\_\_$ is a specifier for functions, which are executed and called only by the GPU, while $\_\_host\_\_$ is a specifier for functions, which are executed and called only by the CPU. $\_\_global\_\_$ is a function specifier that operates as an entry point for executing the kernels. The data allocation in the \textit{device} is carried out by the \textit{host}, using the \textit{cudaMalloc}, \textit{cudaFree}, and \textit{cudaMemcpy} functions. Importantly, these are essential functions for CUDA programs, to transfer data from the \textit{host} to the \textit{device} and vice-versa.

%---------------------------------
\subsection{Existing GPU Verifiers}
\vspace{-2pt}
\label{ssec:trabalhos_relacionados}
%---------------------------------

GPUVerify~\cite{betts:2012} proposes a new operational semantics, known as \textbf{S}ynchronous \textbf{D}elayed \textbf{V}isibility (SDV) semantics to verify kernels, which aims to detect data race and barrier divergence. It uses the Boogie verification system~\cite{boogie:2011} to generate verification conditions, which are solved by Z3~\cite{Moura:2008} or CVC4~\cite{CVC4Barrett11} SMT solvers. GPUVerify accepts only kernel function as input, and it disregards the main function, which thus exposes incorrect results for verifying low-level aspects of CUDA programs.

\textbf{S}ymbolic \textbf{E}xecutor with \textbf{S}tatic \textbf{A}nalysis (SESA)~\cite{sesa:2014} and GPU + KLEE (GKLEE)~\cite{Li:2012} are verifiers based on the \textit{concolic} (concrete, but symbolic) execution of CUDA programs; nevertheless they use different approaches to determine symbolic variables. While SESA performs an automatic evaluation, GKLEE needs inputs from the user to define those variables. On the one hand, SESA checks real applications using original configuration of the number of threads and its focus is on data race detection, but it presents inconclusive results regarding access to memory positions.\ On the other hand, GKLEE supports checks related to barrier synchronization, functional correctness, performance, and data race. Importantly, SESA does not verify the main function, while GKLEE considers both kernel and main functions.

\textbf{P}rover of \textbf{U}ser \textbf{G}PU Programs (PUG)~\cite{Li:2010} analyzes kernels automatically using SMT solvers and it detects data race, barrier synchronization, and conflicts on shared memory. PUG faces problems with invariant derivation for loops, which can lead to incorrect results, and thus requires the user to provide those invariants.\ In addition, problems are also found in arithmetic operations of pointers and advanced C$++$ features.

CIVL~\cite{civl:2015} is a framework for static analysis and concurrent program verification, which supports MPI, POSIX, OpenMP, and CUDA. CIVL recognizes programs written in C++$11$, CUDA--C, and CIVL--C, which is an intermediate language that represents ANSI-C programs based on the aforementioned libraries. CIVL is the tool that is closer to ESBMC-GPU in terms of verification techniques, verifying the concurrent program with partial order reduction, in order to eliminate unnecessary interleavings. The verified properties include: user-specified assertions, deadlocks, memory leaks, invalid pointer dereference, array out-of-bounds, and division by zero. The support of CIVL for CUDA libraries is still in progress, being effective and efficient in programs with kernels that do not use many CUDA functionalities.

%---------------------------------
\subsection{ESBMC}
\vspace{-2pt}
\label{ssec:esbmc}
%---------------------------------

ESBMC is an open source, permissively licensed, award-winning context-bounded model checker for C/C$++$ programs based on SMT solvers~\cite{cordeiro:2011,cordeiro:2012,Morse11,Morse:15}. It is able to verify single- and multi-threaded software with shared variables and locks using the Phtread/POSIX. In particular, ESBMC is one of the most prominent BMC tools for verifying ANSI-C programs according to the last editions of the International \textbf{Comp}etition on \textbf{S}oftware \textbf{V}erification (SV-COMP)~\cite{svcomp2014,svcomp2015}. ESBMC can verify programs that contain bit-level, arrays, structs/unions, pointers, dynamic memory allocation, as well as fixed-point arithmetic. It can reason about arithmetic overflows, pointer safety, memory leaks, array out-of-bounds, atomicity and order violations, local and global deadlocks, data races, and user-specified assertions.

In ESBMC, ANSI-C/C++ programs are typically represented as a state transition system $M = (S, R, s_{0})$, which are extracted from the \textbf{C}ontrol-\textbf{F}low \textbf{G}raph (CFG). $S$ represents the set of states, $R \subseteq S \times S$ represents the set of transitions, and $s_{0} \subseteq S$ represents the set of initial states.\ A state $s \in S$ consists of the value  of the program counter \emph{pc} and the values of all program variables.\ An initial state $s_{0}$ assigns the initial program location of the CFG to \emph{pc}.\ Each transition $\gamma=(s_i,s_{i+1}) \in R$ between two states $s_{i}$ and $s_{i+1}$ is identified with a logical formula $\gamma(s_i,s_{i+1})$, which captures constraints on the corresponding values of the program counter and the program variables.

In particular, consider the transition system \textit{M}, a safety property $\phi$, a context bound $C$, and a bound $k$, ESBMC constructs a \textbf{R}eachability \textbf{T}ree (RT), which represents the program unfolding for $C$, $k$, and $\phi$. A \textbf{v}erification \textbf{c}ondition (VC) $\psi^{\pi}_{k}$ is then derived for each given interleaving $\pi = \{\nu_1,\ldots, \nu_k\}$, such that $\psi^{\pi}_{k}$ is satisfiable \textit{iff} $\phi$ has a counterexample of depth \textit{k}, which is exhibited by $\pi$.\ $\psi^{\pi}_{k}$ is given as follows
%
\begin{equation} \label{bounded-model-checking} \psi^{\pi}_{k} =
I(s_{0}) 
\wedge 
  \bigvee^{k}_{i=0} \bigwedge^{i-1}_{j=0} \gamma(s_{j},s_{j+1})
\wedge 
  \neg \phi(s_i)
\end{equation}
%
\noindent $I$ characterizes the set of initial states of $M$ and $\gamma(s_{j},s_{j+1})$ is the transition relation of $M$ between time steps $j$ and $j+1$. Hence, $I(s_0)\wedge\bigwedge^{i-1}_{j=0} \gamma(s_{j},s_{j+1})$ represents executions of $M$ of length $i$ and $\psi^{\pi}_{k}$ can be satisfied if and only if for some $i \leq k$ there exists a reachable state along $\pi$ at time step $i$ in which $\phi$ is violated. $\psi^{\pi}_{k}$ is a quantifier-free formula in a decidable subset of first-order logic, which is checked for satisfiability by an SMT solver. If $\psi^{\pi}_{k}$ is satisfiable, then $\phi$ is violated along $\pi$ and the SMT solver provides a satisfying assignment, from which the  program variables values can be extracted, in order to construct a counterexample. A counterexample for a particular property $\phi$ is a sequence of states $s_{0}, s_{1},\ldots, s_{k}$ with $s_{0} \in S_{0}$, $s_{k} \in S$, and $\gamma\left(s_{i}, s_{i+1}\right)$ for $0 \leq i < k$. If $\psi^{\pi}_{k}$ is unsatisfiable, one can conclude that no error state is reachable in $k$ steps or less along $\pi$. Finally, we can define $\psi_{k} = \bigwedge_{\pi}\psi^{\pi}_{k}$ and use this to check all paths.

ESBMC combines symbolic model checking with explicit state space exploration; in particular, it explicitly explores the possible interleavings (up to the given context bound) while it treats each interleaving itself symbolically. ESBMC simply traverses the RT depth-first, and calls the single-threaded BMC procedure on the interleaving whenever it reaches an RT leaf node. It stops when it finds a bug, or has systematically explored all possible RT interleavings.

%---------------------------------
\subsubsection{Software Architecture}
\vspace{-2pt}
\label{ssec:SoftwareArchitecture}
%---------------------------------

\begin{figure*}[htb]
\includegraphics[width=\textwidth]{images/esbmc-arch-2}
\caption{ESBMC architecture.}
\label{figure:esbmc-arch}
\end{figure*}

Fig.~\ref{figure:esbmc-arch} shows the ESBMC architecture. Rounded white
rectangles represent input and output; Squared gray rectangles represent the
verification steps.

ESBMC offers two front-ends to parse and generate the \textbf{I}ntermediate
\textbf{R}epresentation (IR) of the program. On the old front-end, based on
one from CBMC~\cite{Clarke:2004}, ESBMC would first
parse the program using flex/bison~\cite{Levine:2009:FB:1696439} to generate an
\textbf{A}bstract \textbf{S}yntax \textbf{T}ree (AST), which is converted to IR. The IR is then type-checked,
which is language-specific. During the type-checking, the code is statically
analysed, including assignment checks, type-cast checks, pointer
initialisation checks, function call checks and template instantiation.

The new front-end offers a much simpler and stronger alternative, but it does not support C++ yet. 
In particular, the new front-end uses clang~\cite{Lopes:2014:GSL:2692607} to generate an AST, which is
then converted to an IR. Clang is a state-of-art compiler for
C/C++/ObjectiveC/ObjectiveC++ that it is already heavily used in
industry~\cite{clang-usage}. The compiler offers an industry level static
analyser and an API to access and traverse its internal AST, which is used by
ESBMC to generate its IR. By using clang, we avoid the need to maintain our own
front-end (a real challenge, given that the ANSI-C and C++ standards are rapidly
evolving nowadays), and can focus on the main objective of the tool, the formal
verification of a program. ESBMC currently can only convert the AST generated
for ANSI-C programs.

The subsequent steps are the same as already described us in previous
papers~\cite{cordeiro:2012}. Regardless of the chosen front-end, the output is the
IR that will be used by the GOTO converter to generate the GOTO program, which
is a simplified version of the original program. The symbolic execution executes
the GOTO program (unrolling loops up to bound \emph{k}) and converts the
expressions to \textbf{S}tatic \textbf{S}ingle \textbf{A}ssignments (SSA)~\cite{cytron89}. Finally, two
sets of quantifier-free formula are created based on the SSA expressions
(\textit{C} for the constraints and \textit{P} for the properties), and used as
input for an SMT solver that will produce a counterexample if there is a
violation of a given property, or an unsatisfiable answer (\textit{i.e.}, verification
successful) if the property holds.

%----------------------------------------------	    	
\vspace{-6pt}
\section{Verifying CUDA Programs}
\label{sec:methodology}
\vspace{-2pt}
\label{sec:metodologiaProposta}
%----------------------------------------------

%----------------------------------------------
\subsection{Operational Models for CUDA Libraries}
\vspace{-2pt}
\label{sec:modeloOperacional}
%----------------------------------------------

In order to correctly support CUDA functionalities, operational models are developed to reliably simulate the behavior of the CUDA libraries. Such approach was previously attempted in the formal verification of C++ programs~\cite{ramalho:2013}, Qt-based applications~\cite{monteiro:2015,garcia:2016}, and Android mobile applications~\cite{android:2012,android:2015}. In particular, our operational model consists of an abstract representation of a set of methods and data structures, which conservatively approximate the CUDA libraries semantics; every method simulates the library's real behavior, including pre- and post-conditions by means of assertions to ensure the correct operation.\ Thus, the operational model contains only methods for verification, ignoring irrelevant calls ({\it e.g.}, screen-printing methods), where there is no relevant property to be checked in terms of software. As a result, our verification focuses on the operational model of the CUDA libraries, and how it is used to verify real-world CUDA programs; this simplifies significantly the model verification and consequently reduces the verification time. The operational model also includes built-in assertions, which check specific properties ({\it e.g.}, array out-of-bounds, arithmetic overflow, pointer safety, and data races). As shown in Fig.~\ref{figure:com}, the operational model is passed to ESBMC's front-end together with the source code to be verified, in order to produce reliable AST with all relevant information for a trustful verification process.

\begin{figure*}[htb]
\centering
\includegraphics[width=120mm]{images/cuda-operational-model.pdf}
\caption{Link between CUDA operational model and ESBMC.}
\label{figure:com}
\end{figure*}

In our CUDA operational model (COM), methods, data types, qualifiers, and CUDA directives were implemented in C$++$. As an example, Figure~\ref{fig:dim3_implementation} shows the implementation of the data type \emph{\tt dim3}, which is represented by a \emph{\tt struct} with three variables: $x$, $y$, and $z$. The \emph{\tt dim3} model has a default constructor, a constructor that receives a variable \emph{\tt uint3} as parameter (this variable is modelled by a \emph{\tt struct} with three \emph{\tt uint} variables), and a constructor that is initialized by a \emph{\tt dim3} variable. In addition to the (typical) primitive data types of ANSI-C/C$++$, COM supports specific data types of CUDA (\textit{e.g.}, \emph{\tt char1}, \emph{\tt short2}, and \emph{\tt float2}).

\begin{figure} [htb]
\centering
\begin{minipage}{0.95\textwidth}
\begin{lstlisting}
struct __dim3
{
  unsigned int x, y, z;
  __dim3(unsigned int vx=1, unsigned int vy=1, unsigned int vz=1){
  	x=vx; y=vy; z=vz;
  }
  __dim3(uint3 v){
  	x=v.x; y=v.y; z=v.z;
  }
  __dim3(__dim3 d){
  	x=d.x; y=d.y; z=d.z;
  }
  operator uint3(void){
  	uint3 t; t.x=x; t.y=y; t.z=z;
	return t;
  }
};
typedef struct __dim3 dim3;
\end{lstlisting}
\end{minipage}
\caption{\emph{\tt dim3} implementation.}
\label{fig:dim3_implementation}
\end{figure}

COM also models methods typically used to develop real CUDA applications. In this sense, we support most CUDA driver and runtime APIs; in particular, ESBMC-GPU supports the CUDA call kernel (\textit{device\_launch\_parameters} and \textit{vector\_types}), driver\_types (\textit{math\_functions}, \textit{cuda\_runtime\_api}, and \textit{host\_definitions}), sm\_atomic\_functions (\textit{vector\_types}), and curand\_kernel (\textit{curand}). Figure~\ref{figure:comOrganization} shows the CUDA operational model organization (internally) implemented in ESBMC-GPU.

\begin{figure*}[htb]
\centering
\includegraphics[width=130mm]{images/com.pdf}
\caption{CUDA Operational Model Organization.}
\label{figure:comOrganization}
\end{figure*}

Here, we describe three essential functions in CUDA programs: \textit{cudaMalloc}, \textit{cudaMemcpy}, and \textit{cudaFree}. Algorithm~\ref{cudaMalloc-algorithm} shows an example of an operational model developed for the \textit{cudaMalloc} function, which abstracts the GPU memory hierarchy and accepts as input arguments, a pointer to allocate memory on the device ({\it i.e.}, \textit{devPtr}) and the size in \textit{bytes} needed for memory allocation (\textit{i.e.}, $size$). In order to allocate memory for \textit{devPtr} pointer (in step $3$), this algorithm uses the \textit{malloc} function, which represents the memory allocation on \textit{device}, checking whether there is a successful allocation (in step $4$). If so, the function returns \emph{\tt CUDA\_SUCCESS}; otherwise, it returns an error specified by \emph{\tt CUDA\_ERROR\_OUT\_OF\_MEMORY}. The variable \textit{lastError} is global and stores the last \emph{\tt cudaError\_t} value to be used in the \textit{cudaLastError()}. In addition, \textit{cudaMalloc()} has, as precondition, a positive memory size allocation; step $2$ of Algorithm~\ref{cudaMalloc-algorithm} includes an assertion in which the size to be allocated must be greater than zero. If there is a violation in this precondition, then ESBMC-GPU returns an error message, specifying the error together with the counterexample. As postcondition, we check whether the memory has been successfully allocated (in step $7$).
%
\begin{algorithm}
\caption{Algorithm of \textit{cudaMalloc} operation.}
\label{cudaMalloc-algorithm}
\begin{algorithmic}[1]
  \Function{cudaMalloc}{\textit{\textbf{void} ** devPtr, \textbf{size\_t} size}}
  \State Check precondition: $size$ must be greater than zero
  \State Allocate a block of memory, which its magnitude is equal to $size$, for the $devPtr$ pointer
  \State If $*devPtr$ is equal to \emph{ \tt NULL}, go to step 5, otherwise go to step 6
  \State Assign \emph{\tt CUDA\_ERROR\_OUT\_OF\_MEMORY} status to the global variable $lastError$
  \State Assign \emph{\tt CUDA\_SUCCESS} status to the global variable $lastError$
	\State Check postcondition: $devPtr$ must be different than NULL
  \State Return $lastError$ and terminates
  \EndFunction
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{cudaMemcpy-algorithm} presents the $cudaMemcpy()$ function operational model. It checks, as precondition, the memory size to be copied (in step $2$). Then, two local variables named as $cdst$ and $csrc$ are used to receive arguments, which represent destination and source of the data copying (in steps $3$ and $4$). This model defines the number of \textit{bytes} to be copied (in step $5$); the data copy is actually performed (in step $6$) between \textit{device} and \textit{host}. As postcondition, we check whether $cdst$ and $csrc$ contain the same data (in step $7$). Finally, \textit{cudaMemcpy} function returns $CUDA\_SUCCESS$ value.

\begin{algorithm}
\caption{Algorithm of \textit{cudaMemcpy} operation.}
\label{cudaMemcpy-algorithm}
\begin{algorithmic}[1]
  \Function{cudaMemcpy}{\textit{\textbf{void} *dst, \textbf{const void} *src, \textbf{size\_t} size, \textbf{cudaMemcpyKind} kind}}
  \State Check precondition: $size$ must be greater than zero
  \State Initialize $cdst$ variable with $dst$'s content
  \State Initialize $csrc$ variable with $src$'s content
  \State Initialize $numbytes$ variable with the number of bytes to be copied (\textit{i.e.}, $size$)
  \State Copy $numbytes$ positions from $csrc$ to $cdst$
	\State Check postcondition: $cdst$ and $csrc$ must contain the same data
  \State Assign \emph{\tt CUDA\_SUCCESS} status to the global variable $lastError$
  \State Return $lastError$ and terminates
  \EndFunction
\end{algorithmic}
\end{algorithm}

Similarly to \textit{cudaMalloc} and \textit{cudaMemcpy}, we also provide the implementation of the \textit{cudaFree} function. Algorithm~\ref{cudaFree-algorithm} shows the operational model, where a pointer to a variable is passed as an input argument so that the allocated memory is released. As precondition, we check whether the pointer $devPtr$, which points to the memory block previously allocated with \textit{malloc}, is not NULL in order to avoid double free. Additionally, as we can see (in step $3$), the \textit{cudaFree} function is referenced to the \textit{free} function of the ANSI-C programming language, however, in this case the memory is freed by the GPU. Importantly, this situation does not affect the verification results, because the memory allocation and deallocation model is represented by a tuple in the background theories of the (supported) SMT solvers, and properties that are analyzed do not take into account the memory hierarchy~\cite{tacas14}. Finally, we assign \textit{CUDA\_SUCCESS} to \textit{lastError}, which has the same functionality as in \textit{cudaMalloc}.

\begin{algorithm}
\caption{Algorithm of \textit{cudaFree} operation.}
\label{cudaFree-algorithm}
\begin{algorithmic}[1]
  \Procedure{cudaFree}{\textit{\textbf{void} *devPtr}}
	\State Check precondition: $devPtr$ must be different than NULL
  \State Call $free$ ANSI-C function to deallocate the memory from $devPtr$ pointer
  \State Assign \emph{\tt CUDA\_SUCCESS} status to the global variable $lastError$
  \State Return $lastError$ and terminates
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Note that the operational models are implemented according to the NVIDIA Programming Guide~\cite{cudaproguide:2015}. The behavior of those functions can be represented in C/C$++$ programming languages using native functions, which are already supported by ESBMC ({\it e.g.}, \textit{malloc}, \textit{free}, \textit{assert}). The soundness proof for those native functions supported by ESBMC can be found in Cordeiro {\it et al.}~\cite{cordeiro:2012}. In particular, the \textit{cudaMalloc} function operates similarly to the \textit{malloc} function, which accepts as input argument the size of the variable to be allocated; the behavior of this function is in compliance with the ANSI-C semantics.\ However, the conceptual difference for CUDA programs is that the memory allocation is carried out in the GPU, which is abstracted since neither hardware functions nor memory hierarchy are included into our operational model, as also done by~\cite{betts:2012,Li:2010,Li:2012}.\ The \textit{cudaMemcpy} function is implemented similar to \textit{Memcpy} function; the only difference is one additional parameter, which determines whether the operation is from \textit{device} to \textit{host} or vice-versa.

%----------------------------------------------
\subsubsection{Correctness of the CUDA Operational Model (COM)}
\label{sec:CorrectnessCUDAmodeloOperacional}
%----------------------------------------------

The idea of making an environment operational model to perform model checking of real applications has been done before~\cite{monteiro:2015,garcia:2016,android:2012,android:2015}, and the correctness of such model to trust in the verification results is actually the major issue. Consequently, the usefulness of our approach relies on the fact that COM correctly represents the original CUDA libraries. In that sense, all developed COM modules were manually verified and exhaustively compared to the original ones to ensure the same behavior; they all contain precondition and postcondition to ensure that a (given) predicate holds before and after the execution of a (given) function, respectively.

Additionally, although COM is a new implementation, it consists in (reliably) constructing a simplified model of the related CUDA libraries, using the same language and by means of the original code and documentation (pre- and postconditions are tested using assertions within the code itself), which thus tends to decrease the resulting number of errors. Note further that the behavior of the CUDA libraries functions are actually represented in C/C$++$ programming languages using native functions ({\it e.g.}, \textit{malloc}, \textit{free}, \textit{assert}). The soundness proof for those native functions, which are already supported by ESBMC, can be found in Cordeiro {\it et al.}~\cite{cordeiro:2011,cordeiro:2012}. Although proofs regarding the soundness of the entire COM could be carried out, it represents a hard task due to the (adopted) memory model~\cite{Mehtaa:2005}. To further improve the correctness of COM, conformance testing could be employed in practice~\cite{Pedro:2011,Pedro:2009}; however, that option is not available in the present case, although it is an interesting possibility for future work.

%-------------------------------------
\subsection{Modeling Kernel with Pthread Functions}
\vspace{-2pt}
\label{sec:pthreadkernels}
%-------------------------------------

The ESBMC's architecture is designed to handle concurrent programs in the C/C$++$ programming languages using the Pthread/POSIX library~\cite{posix:2008}. Thus, ESBMC's verification model is guided by the processing method used by the CPU using that library, where thread instructions can interleave to form (different) execution paths. To apply this methodology to CUDA kernel verification, it is necessary to apply code transformations to the kernel call using ESBMC's intrinsic functions. In particular, the kernel call provides the thread/block configuration in a CUDA program; the parameters used by the (intrinsic) function are obtained from the same \textit{struct} in that kernel call. As a result, the first step to verify a CUDA kernel is to create a new function \textit{ESBMC\_verify\_kernel} using templates, in order to support different data types and parameters.

\begin{algorithm}
\caption{Algorithm of \textit{ESBMC\_verify\_kernel} operation.}
\label{fig:verify_kernel_template}
\begin{algorithmic}[1]
  \Procedure{ESBMC\_verify\_kernel}{\textit{\textbf{RET} *kernel, \textbf{BLOCK} blocks, \textbf{THREAD} threads, \textbf{T1} arg1, \textbf{T2} arg2, \textbf{T3} arg3}}
	\State Check precondition: $kernel$ must be different than NULL, $blocks$ and $threads$ must be greater than zero
  \State Initialize $gridDim$ variable with threads' dimension using $dim3$ function
  \State Initialize $blockDIM$ variable with blocks' dimension using $dim3$ function
  \State Call $ESBMC\_verify\_kernel\_wta$ function in order to determine the type of the $*kernel$
  \State Call $pthread\_join$ ANSI-C function to each GPU thread and terminates
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{fig:verify_kernel_template} shows the implementation of \textit{ESBMC\_verify\_kernel} function, which supports six input parameters. The \textit{kernel} parameter is a pointer to the original program kernel function, the \textit{blocks} and \textit{threads} parameters receive the blocks and threads configuration, which can be of type \textit{int} or \textit{dim3}, and the \textit{arg1}, \textit{arg2}, and \textit{arg3} parameters correspond to values sent to the (respective) function. As preconditions, we check whether $kernel$ is a valid pointer and $blocks$ and $threads$ are greater than zero. Internally, there are two variables, \textit{gridDim} (step $3$) and \textit{blockDim} (step $4$), which receive the output of the \textit{dim3} constructor used to configure the blocks and threads dimension. The function \textit{ESBMC\_verify\_kernel\_wta} (in step $5$) particularizes the argument type to \textit{int} and determines the type of the kernel function pointer. Lastly, there is a loop (in step $6$), which synchronizes by means of the \textit{pthread\_join} function, the threads that are created at the end of the kernel process. The \textit{ESBMC\_verify\_kernel} function is implemented to operate with a real number of threads in the program. However, with the reduction to the two-threads analysis (see Section~\ref{sec:twothreadsReduction}), the respective loop in step $6$ is bounded to two unwindings.

\begin{algorithm}
\caption{Algorithm of \textit{ESBMC\_verify\_kernel\_wta} operation.}
\label{fig:verify_kernel_wta}
\begin{algorithmic}[1]
  \Procedure{ESBMC\_verify\_kernel\_wta}{\textit{\textbf{void} *(*kernel)(int*,int*,int*), \textbf{int} blocks, \textbf{int} threads, \textbf{void} *arg1, \textbf{void} *arg2, \textbf{void} *arg3}}
  \State Allocate a block of memory for the $threads\_id$ pointer, in order to store thread's ids
  \State Allocate a block of memory for each component of $dev\_three$
	\State Check precondition: $threads\_id$ and $dev\_three$ must be different than NULL
  \State Assign $arg1$, $arg2$, and $arg3$ to each component of $dev\_three$, respectively
  \State Assign $*kernel$ to the function pointer component of $dev\_three$
  \State Call $assignIndexes$ function
  \State Call $pthread\_create$ ANSI-C function to create each GPU thread and terminates
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{fig:verify_kernel_wta} shows the reasoning behind the \textit{ESBMC\_verify\_kernel\_wta} function, which is responsible for translating GPU threads into POSIX threads. The \textit{threads\_id} pointer (in step $2$) stores the thread IDs of \textit{pthread\_t} type, the \textit{dev\_three} variable (in steps $3$ to $6$) is an instance of \textit{struct} and stores a pointer to the kernel function, and the \textit{assignIndexes} function (in step $7$) calculates the thread position in the grid using the ID of a thread (pthread) and a configuration of blocks and threads in the CUDA program; such values are stored in a \textit{dim3} vector and are statically calculated to reduce paths generated during the verification. Lastly, there is a loop (in step $8$), where threads are created by means of \textit{pthread\_create} function call and the function that will correspond to the thread is called via \textit{ESBMC\_execute\_kernel\_t}.

\begin{algorithm}
\caption{Algorithm of \textit{ESBMC\_execute\_kernel\_t} operation.}
\label{fig:execute_kernel_t}
\begin{algorithmic}[1]
  \Procedure{ESBMC\_execute\_kernel\_t}{\textit{\textbf{void} *args}}
	\State Check precondition: $args$ must be different than NULL
  \State Call function pointer component of $dev\_three$ (\textit{i.e.}, $dev\_three.func$) and terminates
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{fig:execute_kernel_t} shows the implementation of \textit{ESBMC\_execute\_kernel\_t} function. It checks the original kernel function using its parameters, which are passed as arguments by the \textit{struct dev\_three}. ESBMC-GPU then verifies this function and interleaves its possible execution paths using its native functions.

%-------------------------------------
\subsubsection{Thread Synchronization}
\vspace{-2pt}
\label{sec:threadsynchronization}
%-------------------------------------

Note that in our proposed approach, each CUDA thread is actually mapped to a thread using the internal representation of ESBMC (via Pthread library)~\cite{cordeiro:2011}. In ESBMC, a thread $t$ is a sublist of commands between \emph{begin\_thread} and \emph{end\_thread}, which represent just scoping constructs and do not contribute to the expansion of the RT. Threads are created via asynchronous procedure calls (\emph{start\_thread}), which return an integer that can be used as thread identifier for synchronization (\emph{join\_thread}); hence, dynamic thread creation is allowed. Thus, the thread representation of ESBMC follows a similar approach to the official CUDA scheduler~\cite{cuda:2012, gpu:2010, betts:2012}.

One feature of CUDA related to thread synchronization supported by ESBMC-GPU is the \emph{\tt \_\_syncthreads()} intrinsic function, which specifies synchronization points in the GPU kernel. Such alternative guarantees that all threads in a certain block wait before any of them is allowed to proceed~\cite{cuda:2012}. As described in Algorithm~\ref{fig:syncthreads}, we implement this feature in our operational model using Pthread functions. First of all, a $count$ global variable is initialized to control how many times the \emph{\tt syncthreads} function is actually called. In its first call, a mutex is initialized and a condition variable is set up (steps from $2$ to $5$). Then, on each subsequent call, the $count$ variable is incremented (step $6$) and the calling thread is put on hold until it receives a signal (step $8$). When all GPU threads are on hold, the $pthread\_cond\_signal$ function sends a signal to all threads waiting on a specific condition variable (step $9$) and the mutex is then unlocked (step $10$).

\begin{algorithm}
\caption{Algorithm of \textit{\_\_syncthreads} intrinsic function.}
\label{fig:syncthreads}
\begin{algorithmic}[1]
  \Procedure{\_\_syncthreads}{}
	\State If $count$ is equal to $0$, execute steps $3$, $4$, and $5$; otherwise go to step $6$ 
	\State Call $pthread\_mutex\_init$ function to initializes a mutex
	\State Call $pthread\_cond\_init$ function to initializes a condition variable 
	\State Call $pthread\_mutex\_lock$ function to obtain a lock on the indicated mutex
	\State Increment the $count$ variable
	\State If $count$ is different than the number of GPU threads, go to step $8$, otherwise go to step $9$
	\State Call $pthread\_cond\_wait$ function to block the indicated thread until it receives a signal
	\State Call $pthread\_cond\_signal$ function for all GPU threads
	\State Call $pthread\_mutex\_unlock$ function and terminates
  \EndProcedure
\end{algorithmic}
\end{algorithm}

%-------------------------------------	
\subsubsection{Illustrative Example}
\vspace{-2pt}
\label{sec:ilustrativeExample}
%-------------------------------------

The code fragment shown in Fig.~\ref{fig:FigExVerify} has $1$ block and $2$ threads, {\it i.e.}, $M=1$ and $N=2$, respectively. This CUDA program has a kernel (lines $4$ to $6$), which assigns thread's index values to an array passed as an input argument. The goal is to instantiate array positions, according to the thread index. Despite that, there is a mistake in the array index, as the value $1$ is accidentally added to the thread index (in line $5$). As shown in the main function, array positions are assigned with value $0$ (line $14$), and after the kernel call (line $16$), it is expected by the programmer that $a[0]==0, a[1]==1$.

In this example, however, ESBMC-GPU detects an array out-of-bounds violation. Indeed, this CUDA program retrieves a memory region that has not been previously allocated, so that when $threadIdx.x = 1$, the program tries to access the position $a[2]$. Analyzing the \textit{cudaMalloc()} function operational model, there is a precondition that checks if the memory size to be allocated is greater than zero. Assertions check if the result matches the expected postconditions (line $19$). The verification of this specific program produces $34$ successful and $6$ failed interleavings in ESBMC-GPU. One possible failed interleaving is represented by the threads executions $t0:{a[1]=0};$ $t1:{a[2]=1}$, where $a[2]=1$ represents an incorrect access to the array index $a$.

 \begin{figure} [ht]
 \centering
 \begin{minipage}{0.9\textwidth}
 \begin{lstlisting}
 #define M 1
 #define N 2

 __global__ void kernel(int *A) {
   A[threadIdx.x + 1] = threadIdx.x;
 }

 int main(){
   int *a; int *dev_a;
   int size = N * sizeof(int);
   a = (int*) malloc(size);
   cudaMalloc((void**)&dev_a, size);
   for (int i = 0; i < N; i++)
      a[i] = 0;
   cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
   ESBMC_verify_kernel(kernel, M, N, dev_a);
   cudaMemcpy(a,dev_a,size, cudaMemcpyDeviceToHost);
   for (int i = 0; i < N; i++)
      assert(a[i]==i);
   cudaFree(dev_a);
   free(a);
 }
 \end{lstlisting}
 \end{minipage}
 \caption{Code snippet to index \textit{array}.}
 \label{fig:FigExVerify}
 \end{figure}

ESBMC-GPU and GKLEE are able to detect this array out-of-bounds violation, but GPUVerify and PUG fail to detect such violation, presenting a true incorrect (missed bug), and CIVL does not support this benchmark since it does not handle the identifier \textit{threadIdx}.

The \textit{cudaMalloc} function (line $12$) verifies the precondition, in order to check whether \textit{size} is greater than zero. Internally, the allocation of this memory area is converted to a \textit{malloc} function, which is interpreted by ESBMC-GPU as an ANSI-C function. The \textit{cudaMemcpy} function (lines $15$ and $17$) also checks whether \textit{size} is greater than zero and, if so, it performs a copy using the \textit{memcpy} function.

\begin{figure*}[htb]
\lstset{basicstyle=\scriptsize}
\center
\includegraphics[width=14.5cm, height=7.5cm]{images/eps/Pereira_09_COM.eps}
\caption{COM conversion steps.}
\label{fig:egpu-verify-arc}
\end{figure*}

The \textit{ESBMC\_verify\_kernel} function receives as parameter the kernel function name (\textit{kernel}), the number of blocks (\textit{M}) and threads (\textit{N}), and the \textit{int} parameter \textit{dev\_a}. This function runs the grid configuration, creating a \textit{dim3} representation and saving it in the \textit{gridDim} and \textit{blockDim} variables. In that particular point, the \textit{assignIndexes} function is executed to create a vector with preprocessed thread IDs, avoiding state overgrowth at the verification paths. A loop triggers those threads with \textit{pthread\_create} and, after concluding its execution, the \textit{pthread\_join} thread synchronizes and finalizes the \textit{ESBMC\_verify\_kernel} execution. The \textit{assert} function checks whether there is any unexpected value in the kernel. Lastly, the \textit{cudaFree} function releases the memory allocated by variables, verifying the occurrence of errors ({\it e.g.}, doublefree).

Figure \ref{fig:egpu-verify-arc} shows the steps executed by the ESBMC-GPU tool, in order to verify the program shown in Figure~\ref{fig:FigExVerify}. Firstly, the \textit{main} function instructions are modeled and converted to C/C$++$ native functions together with COM. C/C$++$ variables and functions are also recognized by the ESBMC-GPU without the need for converting them ({\it e.g.}, lines $10$ and $11$). The CUDA functions \textit{cudaMalloc}, \textit{cudaMemcpy}, \textit{cudaFree}, and \textit{ESBMC\_verify\_kernel}  are part of the COM, which are thus converted to C/C++ and POSIX threads.

%-------------------------------------
\subsection{Monotonic Partial Order Reduction}
\label{sec:mpor}
%-------------------------------------

To reduce the number of threads interleavings in CUDA programs, ESBMC-GPU implements the \textbf{M}onotonic \textbf{P}artial \textbf{O}rder \textbf{R}eduction (MPOR), which was initially proposed by Kahlon {\it et al.}~\cite{KahlonWG09}. In particular, the implementation of MPOR in ESBMC-GPU is inherited from its predecessor (ESBMC)~\cite{morse:2015}. This algorithm classifies transitions inside a multi-threaded program as dependent or independent, which determines whether interleaving pairs always compute the same state, thus removing duplicate states in the RT. For dependent transitions, MPOR considers possible thread execution orders to ensure that all program states are reached. If one transition is independent, then the MPOR algorithm considers only one order, because the program state is the same for other execution orders. As MPOR depends on the next state, ESBMC-GPU then verifies the previous state from the current one to verify whether any dependency between threads exists.

In particular, the MPOR implementation is applied to identify accesses to different positions in (shared) arrays. Typically, threads access unique positions in those global (global) arrays, which do not have dependency between them, thus allowing us to remove redundant states that are generated by the possible thread execution orders. Based on this observation, we have extended the MPOR algorithm in ESBMC to identify transitions in which threads access different array locations. This type of access results in the same state independently of the thread execution order.\ Multiple accesses to specific memory positions in CUDA programs happen due to its concurrent nature, based on the linearized configuration of threads and blocks~\cite{cuda:2012}.\

Next we describe the MPOR algorithm to identify transitions in which threads access different array locations and also outline its execution using a running example. Let $\Pi$ be the RT, each node $\nu$ in $\Pi$ be represented as a tuple $\nu = (A_i, C_i, s_i)$ for a given time step $i$, where $A_i$ represents the currently active thread; $C_i$ represents the context switch number; and $s_i$ represents the current state. Algorithm~\ref{alg:mpor_implementation} shows the main steps of MPOR in ESBMC-GPU.

\begin{algorithm}
\caption{MPOR algorithm to identify accesses to different positions in shared arrays.}
\label{alg:mpor_implementation}
\begin{algorithmic}[1]
  \Function{MPOR}{\textit{$\nu$, $\Pi$}}
  \State Check whether $s_i$ exists in $\Pi$; otherwise, go to step $4$.
  \State Check whether $A_i$ has read/write access to $s_i$; otherwise, go to step $5$
  \State Analyze whether $\gamma\left(s_{i-1},s_i\right)$ is independent on $\Pi$; otherwise, go to step $6$  
  \State Return ``independent'' on $\Pi$ and terminates.
	\State Return ``dependent'' on $\Pi$ and terminates.
  \EndFunction
\end{algorithmic}
\end{algorithm}

As example, Fig.~\ref{fig:kernelMpor}(a) shows the application of MPOR to a CUDA kernel, where the global variable $a$ is written in a position relative to the thread ID;\ in particular, this variable determines the program execution state. On the first interleaving, thread $t_{1}$ reaches $v_1$ resulting in $a=[0,0]$. Thread $t_{2}$ runs and reaches $v_2$ resulting in $a=[0,1]$. The MPOR algorithm then checks whether $v_2$ does not exist (step $2$) and whether transition $t_1 \rightarrow t_2$ is defined as independent (step $4$). Both conditions do not hold and transition $t_1 \rightarrow t_2$ is defined as dependent (step $6$). ESBMC-GPU thus performs all possible interleavings, and the next execution starts with thread $t_2$ from $v_0$, which changes the array's content to $a=[0,1]$. Thread $t_1$ runs and reaches $v_4$, which is similar to $v_2$. Algorithm~\ref{alg:mpor_implementation} then checks whether thread $t_1$ has reached a redundant state (step $2$). As thread $t_1$ has only reached $v_1$, then the algorithm concludes that the transition $t_2 \rightarrow t_1$ is independent, disregarding it (represented in Fig.~\ref{fig:kernelMpor}(a) by dotted lines).

Fig.~\ref{fig:kernelMpor}(b) shows another example in which the thread execution results in different states. On the first interleaving, the thread execution $t_1$ reaches $v_1$ and the thread execution $t_2$ reaches $v_2$. Since the content of $a[1]$ is different in both RT nodes, the condition (step $2$) does not hold. On the second interleaving, thread $t_2$ modifies the array $a$ to $a=[0,1]$ and thread $t_1$ accidentally writes to $a[2]$. Algorithm~\ref{alg:mpor_implementation} then checks whether $v_4$ exists on $\Pi$ (step $2$). Since the condition (step $4$) does not hold, transition $t_2 \rightarrow t_1$ is then defined as dependent. In this particular case, both interleavings are considered and those two thread execution orders result in an array out-of-bounds violation since the array $a$ is of length $2$ (\textit{e.g.}, $v_4$ in dotted lines).

In our experimental evaluation, we observed that the application of MPOR to CUDA programs leads to a substantial performance improvement in ESBMC-GPU. The main reason for this improvement is that our MPOR implementation symbolically encodes the interleavings only if transitions are defined as dependent. In particular, when SMT formulae are built, ESBMC-GPU disregards those interleavings that lead to redundant states and only pass to the SMT solver the ones that have dependent transitions, which thus lead to new program states. With the MPOR analysis, we firstly define the first interleaving and only consider further interleavings if they are not redundant (the proof is described in Kahlon {\it et al.}~\cite{KahlonWG09}).

\begin{figure*}[h]
\lstset{basicstyle=\scriptsize}
\center

\begin{minipage}{.42\textwidth}
  \begin{lstlisting}
__global__ void kernel1(int *a) {
  a[threadIdx.x] = threadIdx.x;
}
  \end{lstlisting}
\end{minipage}
  \hspace{.7cm}

\begin{minipage}{.51\textwidth}
  \begin{lstlisting}
__global__ void kernel2(int *a) {
  if(a[1]==1){ 
    a[threadIdx.x+2] = threadIdx.x; 
  }
  else { a[threadIdx.x] = threadIdx.x; }
}
  \end{lstlisting}
\end{minipage}
\end{figure*}

\begin{figure*}[h]
\lstset{basicstyle=\scriptsize}
\center
\includegraphics[width=14cm, height=4.0cm]{images/eps/Pereira_10_mpor.eps}
\caption{MPOR applied to a kernel with independent (Fig.~\ref{fig:kernelMpor}a) and dependent (Fig.~\ref{fig:kernelMpor}b) transitions.}
\label{fig:kernelMpor}
\end{figure*}

%-------------------------------------
\subsubsection{Differences from MPOR to other interleaving reduction methods}
\label{sec:Differencesfro MPOR}
%-------------------------------------

There are software verifiers that apply interleaving reduction methods to CUDA program verification. As example, GKLEE implements a technique named as ``canonical scheduling'', which records all reads and writes that are run by each thread within barrier interval before continuing to the next thread. Records are related in pairs that can have conflicts and, by means of an SMT solver query, those conflicts are checked to detect whether data race condition exists. If conflicts are not detected, then the canonical schedule is defined to represent the barrier interval for the verification of other properties. The MPOR algorithm performs its analysis and reduction on verification time and checks the properties violation on-the-fly, by encoding (small) SMT formulae. GKLEE first checks data race condition to reduce schedules and then verifies other properties. We can observe that both approaches call several times its back-end verification engine, being GKLEE with canonical scheduling and ESBMC-GPU with BMC and Pthread using MPOR.

%-------------------------------------
\subsection{Two-threads Analysis}
\vspace{-2pt}
\label{sec:twothreadsReduction}
%-------------------------------------

GPU architectures are composed by multiprocessors built upon \textbf{P}rocessing \textbf{E}lements (PE) sets~\cite{cuda:2012,cudaproguide:2015}.\ Those PEs are typically arranged in subgroups (named as warps), which run in the same lockstep, ensuring that threads inside those PEs can synchronize using barriers, while threads from one subgroup run independently~\cite{betts:2012} to threads from another subgroup.

Similar to GPUVerify~\cite{betts:2012} (for checking race- and divergence-freedom) and PUG~\cite{Li:2010}, we also reduce the CUDA program verification to two threads for improving verification time and avoiding the state-space explosion problem. Since CUDA kernels typically manipulate one element of the array, and for each element one thread is used, the two-threads analysis ensures that, errors ({\it e.g.}, data races) that are detected between two threads in a given subgroup, due to unsynchronized accesses to shared variables, are enough to justify the property violation in the program.

   \begin{figure}[!htp]
   \centering
    \subfloat[NVIDIA Fermi GPU\label{subfig-1:fermi}]{%
      \includegraphics[width=0.29\textwidth]{images/eps/fermi-architecture-gray.eps}
    }
    \hspace{2em}
    \subfloat[Two-threads approach\label{subfig-2:two-thread}]{%
      \includegraphics[width=0.39\textwidth]{images/two-thread-approach.pdf}
    }
    \caption{Comparison between the NVIDIA Fermi GPU architecture and the two-threads approach used by used to handle data.}
    \label{fig:comparison}
  \end{figure}

In order to demonstrate the reasoning upon two-threads analysis, a NVIDIA Fermi GPU~\cite{cudaproguide:2015} composed by $32$ processing cores is shown in Figure~\ref{fig:comparison}. Importantly, Figure~\ref{subfig-1:fermi} shows the architecture model of that GPU with $16$ processing cores (highlighted in grey boxes), because the thread processing is performed in half-warps~\cite{fermi:2009}, {\it i.e.}, $16$ threads of the same warp are scheduled, ensuring synchronization when a barrier occurs. Figure~\ref{subfig-2:two-thread} demonstrates the thread representation, which accesses the GPU shared memory. First of all, the numbered white boxes represent the shared memory positions to be accessed, ${t_{r}}^n$ are the thread-readers, and ${t_{w}}^n$ are the thread-writers, where $n$ is the index of each thread; it is worth noticed that the threads are executed on a streaming multiprocessor. In the first case of Figure~\ref{subfig-2:two-thread} (\textit{i.e.}, $S_{1}$), the threads, whose kernel is correctly implemented, do not present a data race condition, once they access different memory positions. In the second case of Figure~\ref{subfig-2:two-thread} (\textit{i.e.}, $S_{2}$), the threads execution of a kernel results in a data race condition, where several threads read or write in the same memory position, at the same lock-step. Finally, the third case of Figure~\ref{subfig-2:two-thread} is quite similar to the situation occurred in $S_{2}$, however, instead of using multiple threads to analyse data-race conditions, ESBMC-GPU ensures that such errors can be detected through the analysis of the behaviour of two threads that are operating over the same memory location.

Note that the two-threads analysis affects mostly the data race verification, where program states must be analyzed with respect to their possible threads interleavings, which lead to an execution order of statements that results in error. In our benchmarks, we observed a substantial performance improvement, considering only two threads, while keeping true incorrect results at low rates.

%---------------------------------------
\vspace{-6pt}
\section{Experimental Evaluation}
\vspace{-2pt}
\label{sec:resultadosExperimentais}
%---------------------------------------

%---------------------------------------
\subsection{Experimental Setup}
\vspace{-2pt}
\label{sec:PlanejamentoeDesenvolvimento}
%---------------------------------------

This section describes the experiments to investigate ESBMC-GPU performance for verifying CUDA programs.\
We also compare ESBMC-GPU to GKLEE~\cite{Li:2012}, GPUVerify~\cite{betts:2012}, PUG~\cite{Li:2010}, and CIVL~\cite{civl:2015}.\ In particular, we evaluate ESBMC-GPU's ability to verify $154$ benchmarks\footnote{ESBMC-GPU and benchmarks are available at http://esbmc-gpu.org}, which are extracted from~\cite{cuda:2012,betts:2012,cudaproguide:2015}; we added a \textit{main} function to those benchmarks that do not contain it. The kernels typically exploit the support for:  arithmetic operations, pointer assignment, $\_\_device\_\_$ function calls, general ANSI-C functions ({\it e.g.}, \textit{memset}, \textit{assert}), general CUDA functions ({\it e.g.}, \textit{atomicAdd}, \textit{cudaMemcpy}, \textit{cudaMalloc}, \textit{cudaFree}, \textit{\_\_syncthreads}), general libraries in CUDA ({\it e.g.}, \textit{curand.h}, \textit{curand\_kernel.h}, \textit{curand\_mtgp32\_host.h}) and the ability to work with variables \textit{int}, \textit{float}, \textit{char} as well as type modifiers ({\it e.g.}, \textit{long} and \textit{unsigned}), pointers to that variables, function pointers, type definitions, and intrinsic CUDA variables ({\it e.g.}, \textit{uint4}).

Note that we used benchmarks from a variety of sources to evaluate ESBMC-GPU's precision and performance, which include CUDA benchmarks suite to check basic functions that are commonly used by real CUDA applications. Our benchmarks suite comprises of $20$ CUDA kernels from NVIDIA GPU Computing SDK v$2$.$0$~\cite{toolkit:2015}, $20$ CUDA kernels from Microsoft C$++$ AMP Sample Projects~\cite{microsoft:2012}, and $104$ CUDA programs that explore a wide range of CUDA functionalities~\cite{cuda:2012,cudaproguide:2015}. Although there are some CUDA benchmarks that do not implement real CUDA applications, they are still valuable to analyze to which extent ESBMC-GPU is able to handle and detect (known) errors; in particular, those benchmarks have been previously used for comparing the performance and precision of different GPU verifiers~\cite{betts:2012}. It is worth notice that the benchmarks related to OpenCL~\cite{Khronos:2011} were not applied to our experimental evaluation, once ESBMC-GPU does not support it yet.

Our experiments answer two research questions: RQ1 (sanity check) which results does ESBMC-GPU obtain upon verifying benchmarks that compose the specified suite? RQ2 (comparison with other tools) what is ESBMC-GPU performance when compared to GKLEE, GPUVerify, PUG, and CIVL?

To answer RQ1, ESBMC-GPU is executed with: \emph{\tt --force-malloc-success}, which considers that there is always enough memory available in the device; \emph{\tt --context-switch C}\footnote{The value of C ranges from $2$ to $4$ context switches.}, which considers a context-switch bound among all threads; and \emph{\tt --I libraries}, which specifies the library directory. We also replace the kernel call by the respective \textit{ESBMC\_verify\_kernel} function using one block and two threads per block. As example, we call ESBMC-GPU as: \emph{\tt esbmc file.cu} \emph{\tt --force-malloc-success --context-switch 2 --I libraries}.

To answer RQ2, we apply GKLEE, GPUVerify, PUG, and CIVL to the ESBMC-GPU benchmarks suite. With GKLEE, we use two commands: \textit{gklee-nvcc} and \textit{gklee}. The first one checks the file to be verified, with the extension ``.cu''. When this command is executed, two new files are generated: a ``.cpp'' file and a runnable file (without extension). Then, the second command is used with the generated runnable file. We call GKLEE as: \emph{\tt gklee-nvcc file.cu -libc=uclibc} or \emph{\tt gklee-nvcc file.cu -libc=klee; gklee file}.

To the verification of GPUVerify, the following modifications are required: 
(a) remove the main function; 
(b) check if the variable initialization performed by the main function is responsible for controlling some conditional declaration inside the kernel; if so, such variable must be initialized by \emph{\tt \_\_requires()} function; 
(c) check if there is any assertion in the kernel; if so, this assertion must be replaced by \emph{\tt \_\_assert()}; 
(d) check if there are specific functions to C/C$++$ libraries; if so, they must be removed, as they are not supported by GPUVerify. 
To run GPUVerify, two options must be used: \emph{\tt --gridDim=M} and \emph{\tt --blockDim=N} to assign the number of blocks and threads (per block), respectively. We call GPUVerify as: \emph{\tt gpuverify file.cu} \emph{\tt --blockDim=2 --gridDim=1}.

Additional options for ESBMC-GPU and GPUVerify are necessary to check for data races and array out-of-bounds, respectively.

Some additional changes to the benchmarks are necessary to use PUG: 
(a) the file extension is changed from ``.cu'' to ``.c''; 
(b) given that PUG is unable to verify main functions, these must be removed, in order to keep the kernel function only; 
(c) PUG's proprietary libraries \textit{my\_cutil.h} and \textit{config.h} must be called inside the ``.c'' file. The first library has definitions of structs, qualifiers, and datatypes. The second one defines the number of blocks and threads (per block); 
(d) The kernel function's name has to be renamed to ``kernel''. As example, we call PUG as: \emph{\tt pug kernel.c}.

To verify CUDA programs with CIVL, we use the command \emph{\tt civl verify file.cu}. If the benchmark contains the \textit{malloc} and \textit{free} functions, we must manually add the standard ANSI-C library \textit{stdlib.h} to the CUDA program, although it is not mandatory to declare it according to the CUDA documentation~\cite{cuda:2012}.

All experiments were conducted on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, $16$ GB of RAM, and Linux OS.\ All times given are wall clock time in seconds as measured by the UNIX time command.

%---------------------------------
\subsection{Evaluating SMT Solvers for Verifying CUDA Programs}
\vspace{-2pt}
%---------------------------------

ESBMC-GPU now supports five different SMT solvers: Z$3$ v$4$.$0$~\cite{Moura:2008}, Boolector v$2$.$0$.$1$~\cite{Brummayer:2009}, Yices v$2$.$3$.$1$~\cite{Dutertre:2014}, MathSAT v$5$~\cite{Cimatti:2013}, and CVC4~\cite{Deters:2014}. The flag $--$solver (\textit{e.g.},$--$boolector) is used to define an SMT solver, which is different from the standard ESBMC-GPU configuration.

The solvers' performance is shown in Table~\ref{table:solvers}, where the 
``tool'' column describes the tool name, 
``true correct'' shows the result in which no error has been found for bug-free benchmarks,
``false correct'' shows the results in which the tool correctly detected a bug,
``true incorrect'' shows the results in which the tool did not detect a bug,
``false incorrect'' shows the results in which the tool incorrectly detected a bug,
``not supported'' shows the results which are not supported by the respective tool,
and ``time'' shows the total verification time.

Table~\ref{table:solvers} shows that Z$3$, Boolector, Yices, and MathSAT produce exactly the same number of true correct results (56). Additionally, Z3, Boolector, and Yices present $63$ false correct results, while MathSAT produced only $61$. These results ensure that Z3, Boolector, and Yices are the most effective SMT solvers for verifying CUDA programs in ESBMC-GPU, where correctness is an important parameter for evaluating the solvers performance. Additionally, if we consider verification time, then Z$3$ has an improved performance of $3,6\%$ over Yices. Thus, we defined Z3 as the (default) SMT solver to verify CUDA programs in ESBMC-GPU.

%---------------------------------
\subsection{Experimental Results}
\vspace{-2pt}
%---------------------------------

Table~\ref{table:nonlin} shows the results of ESBMC-GPU v$1$.$25$, GKLEE (LLVM v$3$.$2$), GPUVerify $2$, PUG (ROSE v$0$.$9$) and CIVL v$1$.$6$; each row means: 
tool name (Tool); 
total number of benchmarks in which the program was analyzed to be free of errors (True Correct); 
total number of benchmarks in which the error in the program was found and an error path was reported (False Correct);
total number of benchmarks in which the program had an error, but the verifier did not find it (True Incorrect); 
total number of benchmarks in which an error is reported for a program that fulfills the specification (False Incorrect);
total number of benchmarks which are not supported (Not Supported); 
verification time, in seconds, for all benchmarks (Time).

ESBMC-GPU is able to correctly verify $82.5\%$ of the benchmarks, GKLEE $71.4\%$, GPUVerify $57.1\%$, PUG $35.1\%$ and CIVL $30.5\%$. Note that ESBMC-GPU produces $1$ true incorrect results, while GKLEE produces $14$, GPUVerify produces $9$, PUG produces $7$ and CIVL does not produce any true incorrect result.
With ESBMC-GPU, this result is due to a null pointer access.\
With GKLEE, errors are due to failure in detecting data race ($10$), unsuccessful detection of attempts to modify constant memory ($2$), incorrect detected assertion ($1$), and null pointer access ($2$).
GPUVerify does not detect data race ($7$), does not detect an array bounds violation ($1$), and does not detect an assertion violation ($1$).

PUG does not detect access to null pointer ($1$), data race ($4$), array bounds violation ($1$), and incorrect detected assertion ($1$). ESBMC-GPU generated $3$ false incorrect results, due to assertions included in the kernel, which should return true, but it fails ($2$), and the partial coverage of the \textit{cudaMalloc} function for copies of float-type variables ($1$).
GKLEE generated $7$ false incorrect results, which are caused by incorrectly detected assertions ($4$), data-races ($1$), array out-of-bounds ($1$) and solver call failure ($1$).
GPUVerify generated $8$ false incorrect results, due to incorrectly detected assertion ($2$) and data-races ($6$).
PUG produces $11$ false incorrect results due to data races incorrectly detected.
CIVL produces $3$ false incorrect results due to detection of memory allocation error ($2$) and assertions ($1$).

ESBMC-GPU had $23$ benchmarks that were not supported. These are related to constant memory access ($3$), the use of CUDA's specific libraries (\textit{e.g., curand.h}) ($7$), and the use of pointers to functions, structures, and char type variables used as kernel call arguments ($13$).
GKLEE has $22$ benchmarks that were not supported, which are due to incorrectly detected assertions ($2$); function pointers ($11$), either used as kernel arguments or as in any other parts of the CUDA program; specific CUDA libraries \textit{curand.h} ($7$), and \textit{switch-case} functions ($2$).

GPUVerify did not support $49$ benchmarks. Since it does not support main functions, this explains most cases ($39$); it also does not support the use of \textit{memset} function ($2$), the array bounds detection ($2$), the use of \textit{math\_functions} ($1$), while  other cases are explained by the absence of support to function pointers, either as kernel function arguments or as in any other part of the CUDA program ($5$).

PUG does not support $82$ benchmarks. As GPUVerify, PUG does not verify main functions and this explains most unsupported cases ($31$), while others are explained by the lack of support to \textit{\_\_syncthreads} function ($12$), function pointers ($9$), and the \textit{curand.h} library ($7$); additionally, PUG does not support the use of unsigned type modifier as argument to the function \textit{atomicAdd} ($6$); changes in variables stored in constant memory ($3$), and inability to handle structs ($2$), variables with \textit{\_\_device\_\_} qualifier ($2$), and \textit{size\_t} type ($1$), in addition to other cases that PUG aborted by returning a false null pointer access ($7$) or because it did not recognize the NULL identifier ($2$).

CILV does not support $104$ benchmarks, this is caused by the use of \textit{atomic functions} ($18$), \textit{cudaThreadSynchronize} ($10$), \textit{threadIdx} ($10$), function \textit{typedef} ($8$), \textit{curand functions} ($7$), \textit{dim3} ($6$), \textit{math\_functions} ($5$), \textit{uint4} ($5$), \textit{\_\_constant\_\_ variables} ($5$), \textit{\_\_attribute\_\_} ($5$), \textit{\_\_restrict}  ($2$),  structs ($2$), scanf ($2$), boolean functions (e.g AND and OR) ($2$), \textit{uint} ($2$), \textit{extern} C ($1$), \textit{\_\_threadfence}  ($1$), typecast ($1$). Other benchmarks could not be verified due to the existence of data-race in benchmarks ($11$) and undefined cases ($1$).

%---------------------------------
\subsubsection{ESBMC-GPU effectiveness over existing GPU verifiers}
\vspace{-2pt}
%---------------------------------

MPOR produced an improved performance of approximately $80\%$ in our benchmarks; in particular, it reduced the verification time from $16$ to $3$ hours. With two-threads analysis, we further reduced the verification time from $3$ hours to $811$ seconds. Although the application of these techniques has considerably improved the ESBMC-GPU performance, it still takes more (verification) time than all other existing tools. This is due to the actual execution of the threads interleavings (which combines symbolic model checking with explicit state space exploration), while in GPUVerify the analysis is fully symbolic, performed only in the kernel level, without considering threads interleavings with the main thread. PUG lower verification time is due to the two-threads analysis and because it does not perform the \textit{main function} verification. GKLEE presents a low verification time due to its directed state/path reduction method.

Additionally, in order to improve the verification performance, ESBMC-GPU uses an abstract representation of the CUDA libraries that properly reflects their semantics. This abstract representation includes pre/post-conditions and simulation features that consider only relevant behavior to be explored from the verification perspective, {\it i.e.}, irrelevant code fragments that unnecessarily complicate the generated VCs are actually abstracted. As a result, during the ESBMC-GPU verification process, the COM replaces the real CUDA libraries, which thus improves its effectiveness to detect more properties violations than other existing GPU verifiers, as shown in our experimental evaluation.

%------------------------------
\vspace{-6pt}
\section{Conclusions}
\label{sec:conclusions}
\vspace{-2pt}
%------------------------------

ESBMC-GPU is able to verify CUDA programs using SMT-based context-bounded model checking and operational models, which recognize CUDA directives and further simplify the verification model.\ This work marks the first application of symbolic model checking with explicit state space exploration using MPOR for verifying CUDA programs. In particular, MPOR led to $80$\% of performance improvement in our benchmarks.\
ESBMC-GPU also presents an improved ability to detect array out-of-bounds and data race violations if compared to GKLEE, GPUVerify, PUG, and CIVL. Additionally, ESBMC-GPU provides fewer incorrect results than all other existing GPU verifiers.
Experimental results also show that ESBMC-GPU presents a successful verification rate of $82.5\%$, compared to $71.4\%$ of GKLEE, $57.1\%$ of GPUVerify, $35.1\%$ of PUG, and $30.5\%$ of CIVL. For future work, we will improve our support to argument types of kernel functions, support stream interleaving, and apply techniques to reduce the number of threads interleavings. Besides, conformance testing procedures will be developed for further validating COM, which could also be applied to CUDA modules.

\noindent \textbf{Acknowledgements.} This research project was supported by the Institute of Technology Development (INDT) and by CNPq 475647/2013-0 (UNIVERSAL).

% ---------------------------------------
\begin{thebibliography}{99}

%1
\bibitem{cuda:2012}
Cheng J, Grossman M, McKercher T.
\newblock {\em {Professional CUDA C Programming}}.
\newblock {John Wiley and Sons, Inc.} 2014.

%2
\bibitem{gpu:2010}
Kirk D, Hwu W.
\newblock {\em {Programming Massively Parallel Processors}}.
\newblock Elsevier Inc. 2010.

%3
\bibitem{betts:2012}
Betts A, Chong N, Donaldson AF, Qadeer S, Thomson P.
%A.~Betts {\it et al.}
\newblock {\em{GPUVerify: A Verifier for {GPU} Kernels}}.
\newblock {OOPSLA} 2012; 113--132.

%4
\bibitem{cordeiro:2011}
Cordeiro L, Fischer B.
\newblock {\em{Verifying Multi-threaded Software using SMT-based Context-Bounded
  Model Checking}}.
\newblock {ICSE} 2011; 331--340.

%5
\bibitem{cordeiro:2012}
Cordeiro L, Fischer B, Marques{-}Silva J.
\newblock {\em{SMT-Based Bounded Model Checking for Embedded {ANSI-C} Software}}.
\newblock {IEEE} Trans. Software Eng. 2012; 38(4):957--974.

%6
\bibitem{ramalho:2013}
%M.~Ramalho {\it et al.}
Ramalho M, Freitas M, Sousa F, Marques H, Cordeiro L, Fischer B.
\newblock {\em{SMT-Based Bounded Model Checking of C++ Programs}}.
\newblock {{ECBS}} 2013; 147--156.

%7
\bibitem{Li:2010}
Li G, Gopalakrishnan G.
\newblock {\em{Scalable SMT-based Verification of GPU Kernel Functions}}.
\newblock {FSE} 2010; 187--196.

%8
\bibitem{Li:2012}
%G.~Li {\it et al.}
Li G, Li P, Sawaya G, Gopalakrishnan G, Ghosh I, Rajan SP.
\newblock {\em{GKLEE: Concolic Verification and Test Generation for GPUs.}}
\newblock {PPoPP} 2012; 215--224.

%9
\bibitem{civl:2015}
%M.~Zheng {\it et al.}
Zheng M, Rogers M, Luo Z, Dwyer M, Siegel S.
\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
\newblock {ASE} 2015.

%10
\bibitem{Biere:2009}
A.~Biere.
\newblock {\em{Bounded Model Checking. Handbook of Satisfiability}}. 
\newblock {IOS Press} 2009; 457--481.

%11
\bibitem{Barrett:2009}
Barrett C, Sebastiani R, Seshia S, Tinelli C.
\newblock {\em{ Satisfiability Modulo Theories. Handbook of Satisfiability}}.
\newblock {IOS Press} 2009; 825--885.

%12
\bibitem{KahlonWG09}
Kahlon V, Wang C, Gupta A.
\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order
  Reduction Technique}}.
\newblock {CAV} 2009; 398--413.

%13
\bibitem{cudaproguide:2015}
NVIDIA.
\newblock {\em {CUDA C Programming Guide}}.
\newblock NVIDIA Corporation 2015.

%14
\bibitem{Pereira15}
Pereira P, Albuquerque H, Marques H, Silva I, Carvalho C, Santos V, Ferreira R, Cordeiro L.
\newblock {\em{Verifica\c{c}\~ao de Kernels em Programas CUDA usando Bounded Model Checking.}} 
\newblock {WSCAD-SSC} 2015; 24--35.

%15
\bibitem{Pereira16}
Pereira P, Albuquerque H, Marques H, Silva I, Carvalho C, Santos V, Ferreira R, Cordeiro L.
\newblock {\em{Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.}} 
\newblock {SAC SVT track 2016; 1648--1653.}

%16
\bibitem{boogie:2011}
Le~Goues C, Leino KRM, Moskal M.
\newblock {\em{The Boogie Verification Debugger}.}
\newblock {{SEFM} 2011; 407--414.}

%17
\bibitem{Moura:2008}
Moura L, Bjorner N.
\newblock {\em {Z3: An Efficient SMT Solver}}. 
\newblock {TACAS} 2008; 337--340.

%18
\bibitem{CVC4Barrett11}
%C. ~Barrett {\it et al.}
Barrett C, Conway C, Deters M, Hadarean L, Jovanovi\'{c} D, King T, Reynolds A, Tinelli C.
\newblock {\em{CVC4.}}
\newblock   {CAV} 2011; 171--177.

%19
\bibitem{sesa:2014}
Li P, Li G, Gopalakrishnan G.
\newblock {\em{Practical Symbolic Race Checking of {GPU} Programs}}.
\newblock {SC} 2014; 179--190.

%20
\bibitem{Morse11}
Morse J, Cordeiro L, Nicole D, Fischer B.
\newblock {\em{Context-Bounded Model Checking of LTL Properties for ANSI-C Software.}} 
\newblock {SEFM} 2011; 302--317.

%21
\bibitem{Morse:15}
Morse J, Cordeiro L, Nicole D, Fischer B.
\newblock {\em{Model checking LTL properties over ANSI-C programs with bounded traces.}} 
\newblock {Software and System Modeling} 2015; 14(1): 65--81.

%22
\bibitem{svcomp2014}
Beyer, D.
\newblock{\em{Status report on software verification.}}
%\newblock In Tools and Algorithms for
%  the Construction and Analysis of Systems, LNCS, vol. 8413, pp. 373--388.
%  Springer Berlin Heidelberg (2014)
\newblock {TACAS} 2014; 373--388.

%23
\bibitem{svcomp2015}
Beyer, D.
\newblock{\em{Software verification and verifiable witnesses.}}
\newblock{TACAS} 2015; 401--416.

%24
\bibitem{Clarke:2004}
Clarke E, Kroening D, Lerda F.
\newblock {\em {A Tool for Checking ANSI-C Programs}}.
\newblock {TACAS} 2004; 168--176.

%25
\bibitem{Levine:2009:FB:1696439}
Levine J, John L.
\newblock {\em {Flex \& Bison}}.
\newblock 1st Edition, O'Reilly Media, Inc., 2009.

%26
\bibitem{Lopes:2014:GSL:2692607}
Lopes B, Auler R 
\newblock {\em {Getting Started with LLVM Core Libraries}}. 
\newblock Packt Publishing, 2014.

%27
\bibitem{clang-usage}
M.~Cade. 
\newblock {\em {Why {A}pple’s swift language will instantly remake computer programming}}.
\newblock http://www.wired.com/2014/07/apple-swift/, 2016.

%28
\bibitem{cytron89}
Cytron R, Ferrante J, Rosen B, Wegman M, Zadeck F. 
\newblock {\em {An Efficient Method of Computing Static Single Assignment Form}}
\newblock {POPL} 1989; 25--35.

%29
\bibitem{monteiro:2015}
Monteiro F, Cordeiro L, de Lima Filho E. 
\newblock {\em {Bounded Model Checking of C++ Programs Based on the Qt Framework}}. 
\newblock {GCCE} 2015; 179--447.
%\newblock {DOI=http://dx.doi.org/10.1109/GCCE.2015.7398699}}

%30
\bibitem{garcia:2016}
Garcia M, Monteiro F, Cordeiro L, de Lima Filho E. 
\newblock {\em {ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
\newblock {SPIN} 2016; 97--103.
%\newblock {DOI=http://dx.doi.org/10.1007/978-3-319-32582-8\_6}}

%31
\bibitem{android:2012}
Heila van der Merwe, Brink van der Merwe, and Willem Visser.
\newblock {\em {Verifying android applications using Java PathFinder}}
\newblock {SIGSOFT Softw. Eng. Notes 37, 6} 2012; 1-5.
%\newblock {DOI=http://dx.doi.org/10.1145/2382756.2382797}}

%32
\bibitem{android:2015}
Heila van der Merwe, Oksana Tkachuk, Brink van der Merwe, and Willem Visser.
\newblock {\em {Generation of Library Models for Verification of Android Applications}}
\newblock {SIGSOFT Softw. Eng. Notes 40, 1} 2015; 1-5.
%\newblock {DOI=http://dx.doi.org/10.1145/2693208.2693247}}

%33
\bibitem{tacas14}
Morse J, Ramalho M, Cordeiro L, Nicole D, Fischer B.
\newblock ESBMC 1.22 - (Competition Contribution).
\newblock {TACAS} 2014; 405--407.

%34
\bibitem{Mehtaa:2005}
Mehtaa F, Nipkowb T.
\newblock {\em {Proving pointer programs in higher-order logic}}.
\newblock {Information and Computation} 2005; 199(1), 200--227.

%35
\bibitem{Pedro:2011}
C\'amara P, Castro J, Gallardo M, Merino P.
\newblock {\em {Verification support for ARINC-653-based avionics software}}.
\newblock {JSTVR} 2011; 21(4): 267--298.

%36
\bibitem{Pedro:2009}
C\'amara P, Gallardo M, Merino P, San\'an D.
\newblock {\em {Checking the reliability of socket based communication software}}. 
\newblock {STTT} 2009; 11(5): 359-374.

%37
\bibitem{morse:2015}
Morse J.
\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
\newblock {University of Southampton, PhD Thesis} 2015.

%38
\bibitem{posix:2008}
Institute of Electrical and Electronics Engineers, Inc.
\newblock {\em {IEEE Standard for Information Technology - Portable Operating System Interface (POSIX) Base Specifications, IEEE Std 1003.1-2008 (Revision of IEEE Std 1003.1-2004)}}.
\newblock {IEEE} 2008.

%39
\bibitem{fermi:2009}
NVIDIA.
\newblock {\em {CUDA Compute Architecture: Fermi}}.
\newblock {NVIDIA Corporation} 2009.

%40
\bibitem{Brummayer:2009}
Brummayer R, Biere A.
\newblock {\em {Boolector: An Efficient SMT Solver for Bit-Vectors and Arrays}}. 
\newblock {TACAS} 2009; 174--177.

%41
\bibitem{Dutertre:2014}
B.~Dutertre.
\newblock {\em {Yices 2.2}}. 
\newblock {CAV} 2014; 737--744.

%42
\bibitem{Cimatti:2013}
Cimatti A, Griggio A, Schaafsma B, Sebastiani R.
\newblock {\em {The MathSAT5 SMT Solver}}.
\newblock {TACAS} 2013; 93--107.

%43
\bibitem{Deters:2014}
Deters M, Reynolds A, King T, Barrett C, Tinelli C.
\newblock {\em {A tour of CVC4: How it works, and how to use it}}. 
\newblock {FMCAD} 2014; 7.

%44
\bibitem{toolkit:2015}
NVIDIA.
\newblock {\em {CUDA Toolkit Release Archive}}. 
\newblock url{https://developer.nvidia.com/cuda-toolkit-archive}, 2015.

%45
\bibitem{microsoft:2012}
Microsoft Corporation.
\newblock {\em {C++ AMP sample projects for download (MSDN blog)}}. 
\newblock url{blogs.msdn.com/b/nativeconcurrency/archive/2012/01/30/c-amp-sample-projects-for-download.aspx}, 2012.

%46
\bibitem{Khronos:2011}
Khronos OpenCL Working Group.
\newblock {\em {The OpenCL Specification - version 1.1}}. 
\newblock Document Revision: 44, 2011.
\end{thebibliography}

\newpage

\begin{table*}[bh]
 \caption{Results of the SMT Solvers Performance. \textbf{True Correct} represents the number of flawless benchmarks correctly verified, while \textbf{True Incorrect} represents the ones verified incorrectly. Similarly, \textbf{False Correct} represents the number of faulty benchmark verified correctly, while \textbf{False Incorrect} represents the ones verified incorrectly. Not Supported represents the number of benchmarks unsuccessfully verified, \textbf{Time} represents the total verification time for each solver, and bold numbers represent the best results in each category.}
\centering
\label{table:solvers}
\begin{tabular}{c c c c c c c}
\hline
\multicolumn{1}{|l|}{\textbf{Result/Tool}}
& \multicolumn{1}{c|}{\textbf{Z3 4.0}} & \multicolumn{1}{c|}{\textbf{Boolector 2.0.1}} & \multicolumn{1}{c|}{\textbf{Yices 2.3.1}}
& \multicolumn{1}{c|}{\textbf{CVC 4}} & \multicolumn{1}{c|}{\textbf{MathSAT 5}} \\ \hline
\multicolumn{1}{|l|}{True Correct}
& \multicolumn{1}{c|}{\textbf{56}} & \multicolumn{1}{c|}{\textbf{56}} 
& \multicolumn{1}{c|}{\textbf{56}} & \multicolumn{1}{c|}{35} & \multicolumn{1}{c|}{\textbf{56}} \\ \hline

\multicolumn{1}{|l|}{False Correct}
& \multicolumn{1}{c|}{\textbf{63}} & \multicolumn{1}{c|}{\textbf{63}} 
& \multicolumn{1}{c|}{\textbf{63}} & \multicolumn{1}{c|}{60} & \multicolumn{1}{c|}{61} \\ \hline

\multicolumn{1}{|l|}{True Incorrect}
& \multicolumn{1}{c|}{\textbf{2}} & \multicolumn{1}{c|}{\textbf{2}} 
& \multicolumn{1}{c|}{\textbf{\textbf{2}}} & \multicolumn{1}{c|}{4} & \multicolumn{1}{c|}{\textbf{2}} \\ \hline

\multicolumn{1}{|l|}{False Incorrect}
& \multicolumn{1}{c|}{5} & \multicolumn{1}{c|}{5} 
& \multicolumn{1}{c|}{4} & \multicolumn{1}{c|}{28} & \multicolumn{1}{c|}{\textbf{2}} \\ \hline

\multicolumn{1}{|l|}{Not supported}
& \multicolumn{1}{c|}{28} & \multicolumn{1}{c|}{28} 
& \multicolumn{1}{c|}{29} & \multicolumn{1}{c|}{\textbf{27}} & \multicolumn{1}{c|}{33} \\ \hline

\multicolumn{1}{|l|}{\textbf{Time (s)}}
& \multicolumn{1}{c|}{\textbf{1828}} & \multicolumn{1}{c|}{4614} 
& \multicolumn{1}{c|}{1894} & \multicolumn{1}{c|}{3326} & \multicolumn{1}{c|}{8398} \\ \hline
\end{tabular}
\end{table*}

\newpage
\begin{table*}[bh]
 \caption{Results of ESBMC-GPU, GKLEE, GPUVerify, PUG, and CIVL. \textbf{True Correct} represents the number of flawless benchmarks correctly verified, while \textbf{True Incorrect} represents the ones verified incorrectly. Similarly, \textbf{False Correct} represents the number of faulty benchmarks verified correctly, while \textbf{False Incorrect} represents the ones verified incorrectly. \textbf{Not Supported} represents the number of benchmarks unsuccessfully verified, \textbf{Time} represents the total verification time for each verifier, and bold numbers represent the best results in each category.}
\centering
\label{table:nonlin}
\begin{tabular}{c c c c c c c}
\hline
\multicolumn{2}{|c|}{\textbf{Result/Tool}}
& \multicolumn{1}{c|}{\textbf{ESBMC-GPU}}	& \multicolumn{1}{c|}{\textbf{GKLEE}}	& \multicolumn{1}{c|}{\textbf{GPUVerify}}	& \multicolumn{1}{c|}{\textbf{PUG}}	& \multicolumn{1}{c|}{\textbf{CIVL}}	\\ \hline
\multicolumn{2}{|c|}{True Correct}
& \multicolumn{1}{c|}{\textbf{60}}		& \multicolumn{1}{c|}{53}	& \multicolumn{1}{c|}{58}	& \multicolumn{1}{c|}{39}	& \multicolumn{1}{c|}{23}	\\ \hline
\multicolumn{2}{|c|}{False Correct}                                               
& \multicolumn{1}{c|}{\textbf{67}} 		& \multicolumn{1}{c|}{57}	& \multicolumn{1}{c|}{30} 	& \multicolumn{1}{c|}{15}	& \multicolumn{1}{c|}{24}	\\ \hline
\multicolumn{2}{|c|}{True Incorrect}
& \multicolumn{1}{c|}{1}			& \multicolumn{1}{c|}{14}       & \multicolumn{1}{c|}{9}	& \multicolumn{1}{c|}{7}	& \multicolumn{1}{c|}{\textbf{0}}	\\ \hline
\multicolumn{2}{|c|}{False Incorrect}
& \multicolumn{1}{c|}{\textbf{3}}		& \multicolumn{1}{c|}{8}        & \multicolumn{1}{c|}{8}	& \multicolumn{1}{c|}{11}	& \multicolumn{1}{c|}{\textbf{3}}	\\ \hline
\multicolumn{2}{|c|}{Not Supported}
& \multicolumn{1}{c|}{23}		& \multicolumn{1}{c|}{\textbf{22}}       & \multicolumn{1}{c|}{49}	& \multicolumn{1}{c|}{82}	& \multicolumn{1}{c|}{104}	\\ \hline
\multicolumn{2}{|c|}{\textbf{Time (s)}}
& \multicolumn{1}{c|}{811}			& \multicolumn{1}{c|}{128}	& \multicolumn{1}{c|}{147}     	& \multicolumn{1}{c|}{\textbf{12}} & \multicolumn{1}{c|}{158}	\\ \hline
\end{tabular}
\end{table*}

\end{document}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/comments/comments.tex
uploading JSCP
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

\documentclass[12pt]{article}

\usepackage[normalem]{ulem}
\usepackage{xcolor}

%please, use \sout{} to indicate that a comment has been implemented

\begin{document}

\setcounter{secnumdepth}{0}

\section{Reviews}

\subsection{Reviewer \#1}

\subsubsection{Summary}

The authors present an extension of the ESBMC tool, to verify CUDA GPU kernels by means of bounded model checking. This is an interesting and relevant topic, since GPUs are used increasingly often to speed up applications significantly. However, programming GPU software can be very tedious and error-prone. A formal verifier of GPU code can make the development process much more efficient.

\subsubsection{Comments}

\begin{itemize}
\item The present submission provides an ok overview of the capabilities of the tool, in my opinion, but the authors never really address the details to fully understand what the tool can and cannot do. For instance, when discussing the experimental evaluation, it is not explained what the source of the bugs is in the various benchmarks.
  \begin{itemize}
  \item Are there clear differences in the types of bugs that the different tools being compared can detect?
  \item Which types could ESBMC-GPU handle in particular better than the other tools, and why?
  \item To what extent does ESBMC-GPU support CUDA?
  \item Can ESBMC-GPU handle variables in different types of memory?
  \item (Shared, global, thread registers?), and inter-warp communication, for instance?
  \item In that respect, what does COM cover, and what remains to be added?
  \item Is the memory model for the different types of memory complete?
  \end{itemize}

\item Another paragraph that raises questions is the second on page 3. It seems to say that COM uses pthread functions to simulate GPU kernels. Is that so, and why is it necessary to simulate them? Is this safe to do, because some features in CUDA are not available in C/C++. How do the authors handle those, or do they not?

\item At some point, the authors mention in passing that Microsoft C++ AMP projects have been used for the evaluation. Does this mean that this language is also supported, or are the benchmarks translated?

\item Concluding, I would like to see an overall more detailed discussion of the capabilities of the tool.

\end{itemize}

\subsubsection{Detailed comments:}

\begin{itemize}
\item Page 2
  \begin{itemize}
  \item space-state exploration $\rightarrow$ state-space exploration;
  \item ESBMC-GPU builds on top of ESBMC $\rightarrow$ ESBMC-GPU is built on top of ESBMC;
  \end{itemize}
\item Page 3
  \begin{itemize}
  \item consists in $\rightarrow$ consists of;
  \item Besides, in: remove ``Besides'';
  \item reduce state space $\rightarrow$ reduce the state space;
  \end{itemize}
\item Page 4
  \begin{itemize}
  \item if multiple threads perform unsynchronized 90 access to shared-memory locations: to the same memory locations, it is crucial to mention that;
  \item has been systematically explored $\rightarrow$ have been systematically explored;
  \item Therefore: the next conclusion does not follow logically from the previous statements, so ``therefore'' should be removed here;
  \end{itemize}
\item Page 5
  \begin{itemize}
  \item Figure 2: between lines 17 and 18, the contents of {\tt dev\_a} should be copied to a with a {\tt cudaMemcpy}, otherwise the subsequent check is quite useless;
  \end{itemize}
\item Page 6
  \begin{itemize}
  \item In addition $\rightarrow$ On the other hand;
  \item Besides, it further simplifies verification models: further compared to what? How does it make it easier?
  \item if compared: remove ``if'';
  \item stream interleaving: what is this?
  \item  to reduce the number of thread interleavings: in what respect? Not reducing the number of threads, right? You already use the 2-thread simplification;
  \end{itemize}
\end{itemize}

\end{document}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/others/cover letter/Cover-Letter.tex
uploading JSCP
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}

%\usepackage{mathtools}
%\mathtoolsset{showonlyrefs}

\usepackage{color}
\usepackage{natbib}
% \usepackage[style=authoryear-comp,dashed=false]{biblatex}
% \bibliography{piecewise_lyapunov}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\newcommand{\Real}{\mathbb{R}}

\pagestyle{empty}

\begin{document}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{logo_ufam.png}}
\end{minipage}}} \hspace*{0.1cm}
\begin{minipage}{11.0cm}{\sf \large \textbf{FEDERAL UNIVERSITY OF AMAZONAS}}\\
{\sc Department of Electronics and Computing}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Mr. Felipe R. Monteiro\\
Federal University of Amazonas\\
Department of Electronics and Computing\\
Avenida General Rodrigo Ot\'avio, 6200, Coroado\\
69077-000 - Manaus - AM - BRAZIL\\
{\em e-mail: felipemonteiro@ufam.edu.br}
\end{minipage}
\end{flushright}

\vi \vi
\hspace*{\fill}{\small Manaus, \today.}
\vi \vi

\begin{flushleft}
    Professor Mark van den Brand\\
    Editor-in-Chief\\
    Science of Computer Programming Software Track
    \end{flushleft}
\vi

\indent Dear Editor-in-Chief,
\vi

Please find enclosed our manuscript entitled as ``{\em ESBMC-GPU: A Context-Bounded Model Checking Tool to Verify CUDA Programs}'' by Felipe R. Monteiro, Erickson H. Alves, Isabela S. Silva, Hussama A. Ismail, Lucas C. Cordeiro, and Eddie B. de Lima Filho, which we would like to submit for publication in the software track of the journal of {\bf Science of Computer Programming}.

\vi

This paper addresses ESBMC-GPU's functionality and discusses its practical capabilities and applications, whose underlying techniques have been previously published at the Journal of Concurrency and Computation: Practice and Experience. We focus here on the usage aspects of ESBMC-GPU, in order to emphasize its design and implementation from the user point of view. In addition, we present a rigorous experimental evaluation of our approach against other state-of-the-art software verifiers ({\it i.e.}, CIVL, GKLEE, GPUVerify) using standard CUDA benchmarks taken from the literature.

\vi

For your reference, we attach the original paper (ccpe2016.pdf) with this submission. These new results have not been published elsewhere and are not under consideration by another journal.

\vi

\begin{quote}
\begin{flushright}
Thank you for your time.~~~~~~~

Sincerely,~~~~~~~

Felipe R. Monteiro~~~~~~~~
\end{flushright}
\end{quote}

\end{document}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
uploading JSCP
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt, a4paper]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{subfig}
\usepackage[american]{babel}
\usetikzlibrary{arrows,automata,positioning}
\usepackage{epstopdf}
\newtheorem{mydef}{Definition}
\newtheorem{mylem}{Lemma}
\usepackage{afterpage}

\renewcommand{\baselinestretch}{0.975}

\lstset{language=C,basicstyle=\tiny}
\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
\lstset{tabsize=2}
\lstset{firstnumber=1}
\lstset{frame=single}
\lstset{
  language={C},
  morekeywords={enquanto,se,fim,entao,senao,retorne,faca,assume,then,end-if,end,input,output,cudaMemcpyHostToDevice,cudaMemcpyDeviceToHost,\_\_global\_\_}
}

\hyphenation{port-a-ble}

\newcommand{\comment}[1]{}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

\journal{Science of Computer Programming}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{ESBMC-GPU \\ A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{Felipe~R.~Monteiro$^{1}$, Erickson~H.~da~S.~Alves$^{1}$, Isabela~S.~Silva$^{1}$,\\Hussama~A.~Ismail$^{1}$, Lucas~C.~Cordeiro$^{1,2}$, and Eddie~B.~de~Lima~Filho$^{1,3}$}

\address{$^{1}$Faculty of Technology, Federal University of Amazonas, Brazil \\ 
$^{2}$Department of Computer Science, University of Oxford, United Kingdom \\
$^{3}$ FPF Tech, Brazil}

\begin{abstract}
The Compute Unified Device Architecture (CUDA) is a programming model used for exploring the advantages of Graphics Processing Unit (GPU) devices, through parallelization and specialized functions and features. Nonetheless, as in other development platforms, errors may occur, due to traditional software creation processes, which may even compromise the execution of an entire system. In order to address such a problem, ESBMC-GPU was developed, as an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). In summary, ESBMC processes input code through ESBMC-GPU and an abstract representation of the standard CUDA libraries, with the goal of checking a set of desired properties. Experimental results showed that ESBMC-GPU was able to correctly verify $85$\% of the chosen benchmarks and it also overcame other existing GPU verifiers.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
GPU verification \sep formal verification \sep model checking \sep CUDA

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

\linenumbers

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Introduction} 
\label{sec:intro}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\vspace{-1mm}

The Compute Unified Device Architecture (CUDA) is a development framework that makes use of the architecture and processing power of Graphics Processing Units (GPUs)~\cite{cuda:2012}. Indeed, CUDA is also an Application Programming Interface (API), through which a GPU's parallelization scheme and tools can be accessed, with the goal of executing kernels. Nonetheless, source code is still written by human programmers, which may result in arithmetic overflow, division by zero, and other violation types. In addition, given that CUDA allows parallelization, problems related to the latter can also occur, due to thread scheduling~\cite{betts:2012}.

In order to address the mentioned issues, an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{cordeiro:2012} was developed, named as ESBMC-GPU~\cite{Pereira15,Pereira16,pereira:2016}, with the goal of verifying CUDA-based programs (available online at \texttt{http://esbmc.org/gpu}). ESBMC-GPU consists of an extension for parsing CUDA source code ({\it i.e.}, a front-end to ESBMC) and a CUDA operational model (COM), which is an abstract representation of the standard CUDA libraries ({\it i.e.}, the native API) that conservatively approximates their semantics.

A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit state-space exploration~\cite{cordeiro:2011,cordeiro:2012}. In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, space-state exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}. As a consequence, redundant interleavings are eliminated, without ignoring a program's behavior.

Finally, existing GPU verifiers often ignore some aspects related to memory leak, data transfer, and overflow, which are normally present in CUDA programs. The proposed approach, in turn, explicitly addresses them, through an accurate checking procedure, which even considers data exchange between main program and kernel. Obviously, it results in higher verification times, but more errors can then be identified and later corrected, in another development cycle.

%\textcolor{red}{seria interessante colocar um resumo sobre o que serah visto no resto do artigo?} \textcolor{blue}{Acho que nao temos espaco para isso} \textcolor{green}{Melhor nao colocarmos este resumo...}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%\subsection{Existing GPU Verifiers}
%\label{sec:related-work}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

%\textbf{Existing GPU Verifiers}. GPUVerify~\cite{betts:2012} is based on Synchronous Delayed Visibility, which focus on detecting data race and barrier divergence, while reducing kernel verification procedures to the analysis of sequential programs. Symbolic Executor with Static Analysis (SESA)~\cite{sesa:2014} and GPU + KLEE (GKLEE)~\cite{Li:2012} are concrete plus symbolic (\textit{concolic}) execution tools. Initially, GKLEE modeled every original thread, which was later improved by taking into account symmetry, and it considers both kernel and main function, while checking barrier synchronization and data race, among others. SESA combines equivalent flows, which results in great performance improvement, focus on data race detection, and does not verify the main function. Prover of User GPU Programs (PUG)~\cite{Li:2010} is based on SMT solvers and applies symbolic static analysis to each kernel, but requires user annotations. CIVL~\cite{civl:2015} generates an abstract syntax tree and supports analysis and transformations. Besides, it also verifies concurrent programs with partial order reduction, which makes it very similar to the solution proposed here. Given that its support to CUDA libraries is still under development, CIVL is an interesting approach for checking simple CUDA programs. \textcolor{red}{$<$---- poderiamos deletar esse paragrafo?}

\vspace{-3mm}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Architecture and Implementation}
\label{sec:arch}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\vspace{-1mm}

ESBMC-GPU builds on top of ESBMC, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.

\vspace{-2mm}

\begin{figure}[htb]
  \centering
  \includegraphics[width=4.8in]{../figures/arch.pdf}
  \caption{Overview of ESBMC-GPU's architecture.}
  \label{figure:arch}
\end{figure}

\vspace{-2mm}

\noindent \textbf{1. CUDA Operational Model.} An operational model for CUDA libraries that provides support to CUDA functionalities, in conjunction with ESBMC, as shown in Fig.~\ref{figure:arch}. Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, consists in an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf. Sec.~\ref{sec:features}). Indeed, COM allows the necessary control for performing code analysis, where both CUDA operation and knowledge for model checking its properties are available.

%\begin{figure}[htb]
%  \centering
%  \includegraphics[width=4.5in]{../figures/cuda-om.png}
%  \caption{CUDA operational model development process.}
%  \label{figure:om}
%\end{figure}

ESBMC was designed to handle multi-threaded software, through the use of an API called Portable Operating System Interface (POSIX -- ISO/IEC 9945)~\cite{posix:2008}. Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).
%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.

\noindent \textbf{2. Two-threads Analysis.} Similarly to GPUVerify~\cite{betts:2012} and PUG~\cite{Li:2010}, ESBMC-GPU also reduces the number of threads (to only two elements), during the verification of CUDA programs, by considering a NVIDIA Fermi GPU architecture, in order to improve verification time and avoid the state-space explosion problem. Besides, in CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check. Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.

%\noindent \textbf{State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and reduce state space, based on SHA256 hashes.\\%, due to the low the likelihood regarding collisions. \\
%

\noindent \textbf{3. State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also reduce state space, based on SHA256 hashes \cite{FIPS:2002}. In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states. Given any state computed during the symbolic execution of a specific CUDA kernel, ESBMC-GPU simply summarizes it and efficiently determines whether it has been explored before or not, along a different computation path. When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT). This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.% \textcolor{red}{similares? entao, qual eh o impacto de se verificar apenas um? eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.

\noindent \textbf{4. Monotonic Partial Order Reduction.} MPOR is used to reduce the number of thread interleavings, by classifying transitions inside a program as dependent or independent. As a consequence, it is possible to determine whether interleaving pairs always lead to the same state and then remove duplicates in a reachability tree, without ignoring any program's behavior~\cite{morse:2015}.

%In addition to both techniques,

\vspace{-4mm}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Functionalities}
\label{sec:features}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\vspace{-1mm}

Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access to shared-memory locations; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.

In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node. Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings has been systematically explored~\cite{pereira:2016}.

%\textcolor{red}{eu nao sei como a disponibilidade de espaco estah, mas seria bom aqui um fechamento de como BMC checa tudo...} \textcolor{blue}{acredito que nao teremos espaco para isso} \textcolor{green}{Felipe, podes referenciar o nosso journal do CPE informando que la explicamos como o BMC checa tudo com mais detalhes.}

\vspace{-4mm}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Illustrative Example}
\label{subsec:usage}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\vspace{-1mm}

In this part, ESBMC-GPU usage is demonstrated, by using the CUDA program shown in Fig.~\ref{fig:illustrative-example}. First of all, users must replace the default kernel call (line $16$) by an intrinsic function of ESBMC-GPU (line $17$). Then, the resulting CUDA program can be passed to the command-line version of ESBMC-GPU, as follows: \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},  where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.

%\begin{center}
%\vspace{-2 mm}
%\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

%\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
%\vspace{-2 mm}
%\end{center}

%\vspace{-0.1cm}

%\textcolor{red}{esta parte ficou solta, pois nao houve explicacao de BMC e, por conseguinte, k (unwind)} \textcolor{blue}{concorco com o comentario, mas como nao temos espaco, estamos considerando que o revisor ja tera conhecimento de tecnicas mais conhecidas, como BMC} \textcolor{green}{sugiro novamente referenciar o journal do CPE.}
%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, respectively. Furthermore, its kernel (lines $8$ to $10$) stores in the variable $c$ the result of the sum of $a$ and $b$, which are passed as input arguments. However, there is a mistake in the assertion (line $26$), as $a + b$ must be equal to $c$, which is correctly detected by ESBMC-GPU detects an assert violation.

%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, and a kernel (lines $12$ to $14$), which assigns thread's index values to an array passed as an input argument. Indeed, its goal is to instantiate array positions according to the thread index. However, there is a mistake in the array index, as the value $1$ is accidentally added to the thread index (line $13$). As shown in the main function, array positions are assigned with value $0$ (line $23$), and after the kernel call (line $26$), it is expected that $a[0] == 0$ and $a[1] == 1$. \textcolor{red}{$<$---- poderiamos deletar esse paragrafo?}

In the mentioned example, ESBMC-GPU detects an array out-of-bounds violation. Indeed, this CUDA-based program retrieves a memory region that has not been previously allocated, {\it i.e.}, when {\tt threadIdx.x = 1}, the program tries to access $a[2]$. Importantly, the {\tt cudaMalloc()} function's operational model has a precondition that checks if the memory size to be allocated is greater than zero. In addition, an assertion checks if the result matches to the expected postcondition (line $19$). Therefore, the verification of this program through ESBMC-GPU produces $54$ successful and $3$ failed interleavings. For instance, one possible failed interleaving is represented by the threads executions $t_0 : a[1] = 0$; $t_1 : a[2] = 1$, where $a[2] = 1$ represents an incorrect access to the array index $a$. It is worth noticing that CIVL, ESBMC-GPU, and GKLEE are also able to detect this array out-of-bounds violation, but GPUVerify fails, as it reports a true incorrect result (missed bug).

\vspace{-3mm}

\begin{figure} [htb]
\centering
\begin{minipage}{8.5cm}
\begin{lstlisting}
#include <...>
#define BLOCKS 1
#define THREADS 2
__global__ void kernel(int *A) {
	A[threadIdx.x + 1] = threadIdx.x;
}
int main(){
	int *a;
	int *dev_a;
	int size = THREADS*sizeof(int);
	a = (int*)malloc(size);
	cudaMalloc((void**)&dev_a, size);
	for (int i = 0; i < THREADS; i++)
		a[i] = 0;
	cudaMemcpy(dev_a, a, size, cudaMemcpyHostToDevice);
	// kernel<<<BLOCKS,THREADS>>>(dev_a);
	ESBMC_verify_kernel(kernel, BLOCKS, THREADS, dev_a);
	for (int i = 0; i < THREADS; i++)
		assert(a[i]==i);	
	cudaFree(dev_a);
	return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Illustrative CUDA code example.}
\label{fig:illustrative-example}
\end{figure}

\vspace{-7mm}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Experimental Evaluation}
\label{sec:effectiveness}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\vspace{-1mm}

In order to evaluate ESBMC-GPU's precision and performance, benchmarks\footnote{A detailed description of all benchmarks is available at {\tt http://esbmc.org/gpu/}} were extracted from the available literature ({\it i.e.}, NVIDIA GPU Computing SDK v$2$.$0$~\cite{toolkit:2015} and Microsoft C$++$ AMP Sample Projects~\cite{microsoft:2012}), which covers basic functions commonly used by real CUDA applications. 
%The chosen benchmark suite comprises $20$ CUDA kernels from NVIDIA GPU Computing SDK v$2$.$0$, $20$ CUDA kernels from Microsoft C$++$ AMP Sample Projects%~\cite{microsoft:2012}
%, and $104$ CUDA-based programs that explore a wide range of CUDA functionalities~\cite{cuda:2012}. In summary, the chosen suite contains $47.4$\% bug-free and $52.6$\% buggy benchmarks, which tackle data race, arithmetic operations, pointer assignment, {\tt \_\_device\_\_} function calls, general ANSI-C functions ({\it e.g.}, \texttt{memset}), general CUDA functions ({\it e.g.}, \texttt{cudaMemcpy}), general libraries in CUDA ({\it e.g.}, \texttt{curand.h}), and the ability to work with variables, type modifiers ({\it e.g.}, \texttt{unsigned}), pointers, type definitions, and intrinsic CUDA variables ({\it e.g.}, \texttt{uint4}) \textcolor{red}{poderiamos colocar toda essa descrição dos benchmarks numa pagina online e cita-la aqui??}.
The present experiments answer two research questions: {\it (i)} How accurate is ESBMC-GPU when verifying the chosen benchmarks? {\it (ii)} How does ESBMC-GPU's performance compare to other existing verifiers?

%\begin{enumerate}

%\item {\bf (sanity check)} Which results does ESBMC-GPU obtain upon verifying benchmarks from the automated test suite?
%\item How accurate is ESBMC-GPU when verifying the chosen test suite?

%\item {\bf (compare software performance)} What is ESBMC-GPU performance when compared to GKLEE, GPUVerify, PUG, and CIVL?
%\item How does ESBMC-GPU compare to other existing verifiers?

%\end{enumerate}

 In order to answer both questions, all benchmarks were verified with $4$ GPU verifiers (ESBMC-GPU $v2.0$, GKLEE $v2012$, GPUVerify $v1811$, and CIVL $v1.7.1$), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM, running Ubuntu 14.04 OS. Importantly, all presented execution times are actually CPU times, {\it i.e.}, only the elapsed time periods spent in the allocated CPUs, which was measured with the {\tt times} system call (POSIX system). An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified ({\it i.e.}, a verification tool reports an unexpected result). As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}. More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%). %\textcolor{red}{e os outros? quais sao as vantagens? esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}

{\bf Limitations.} ESBMC-GPU was unable to correctly verify $24$ benchmarks, which are related to constant memory access ($2$\%), CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when passed as kernel call arguments ($4.5$\%). In addition, it only reported $3$\% of incorrect true and $1$\% of incorrect false results.

%In addition, ESBMC-GPU had $23$ benchmarks that were not supported. These are related to constant memory access ($3$), the use of CUDA's specific libraries (\textit{e.g., curand.h}) ($7$), and the use of pointers to functions, structures, and char type variables used as kernel call arguments ($13$).

{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $789.6$ sec. Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec). This is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}. %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications. 
In addition, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.

\vspace{-2mm}

\begin{figure}[htb]
  \centering
  \includegraphics[width=3.9in]{../figures/experimental-results.pdf}
  \caption{Experimental evaluation of ESBMC-GPU against other verifiers.}
  \label{figure:experimental}
\end{figure}

\vspace{-8mm}

%\noindent \textbf{Availability of Data and Tools.} The performed experiments are based on a set of publicly available benchmarks. All benchmarks, tools, and results, associated with the current evaluation, are available on {\tt http://esbmc-gpu.org/}.

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Conclusions and Future Work}
\label{sec:conc}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

%\vspace{-1mm}

ESBMC-GPU marks the first application of an SMT-based context-BMC tool that recognizes CUDA directives~\cite{pereira:2016}. Besides, it further simplifies verification models and provides fewer incorrect results, if compared with GKLEE, GPUVerify, and CIVL. Finally, it presents improved ability to detect array out-of-bounds and data race violations. Future work aims to support stream interleaving and implement further techniques to reduce the number of thread interleavings, by taking into account GPU symmetry.

%\noindent \textbf{Acknowledgements.} This paper is based on research sponsored by the Institute of Development and Technology (INdT) and by the National Council for Scientific and Technological Development (CNPq) under agreement number $475647/2013-0$.

\section*{References}
\label{}

\begin{thebibliography}{25}

\bibitem{cuda:2012}
Cheng, J., Grossman, M., and McKercher, T.
\newblock {\em {Professional CUDA C Programming}}.
\newblock {John Wiley and Sons, Inc.} 2014.

%\bibitem{gpu:2010}
%Kirk, D. and Hwu, W.
%\newblock {\em {Programming Massively Parallel Processors}}.
%\newblock Elsevier, 2010.

\bibitem{betts:2012}
Betts, A., Chong, N., Donaldson, A., Qadeer, S., and Thomson, P.
\newblock {\em{GPUVerify: A Verifier for {GPU} Kernels}}.
\newblock {OOPSLA} 2012; 113--132.

\bibitem{cordeiro:2011}
Cordeiro, L. and Fischer, B.
\newblock {\em{Verifying Multi-threaded Software using SMT-based Context-Bounded
  Model Checking}}.
\newblock {ICSE} 2011; 331--340.

\bibitem{cordeiro:2012}
Cordeiro, L. and, Fischer, B., and Marques{-}Silva, J.
\newblock {\em{SMT-Based Bounded Model Checking for Embedded {ANSI-C} Software}}.
\newblock {IEEE} Trans. Software Eng. 2012; 38(4):957--974.

\bibitem{Pereira15}
Pereira, P., Albuquerque, H., Marques, H., Silva, I., Carvalho, C., Santos, V., Ferreira, R., and Cordeiro, L.
\newblock {\em{Verifica\c{c}\~ao de Kernels em Programas CUDA usando Bounded Model Checking.}} 
\newblock {WSCAD-SSC} 2015; 24--35.

\bibitem{Pereira16}
Pereira, P., Albuquerque, H., Marques, H., Silva, I., Carvalho, C., Santos, V., Ferreira, R., and Cordeiro, L.
\newblock {\em{Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.}} 
\newblock {SAC SVT track 2016; 1648--1653.}

\bibitem{pereira:2016}
Pereira, P., Albuquerque, H., Silva, I., Marques, H., Monteiro, F., Ferreira, R., and Cordeiro, L.
\newblock {\em{SMT-Based Context-Bounded Model Checking for CUDA Programs}}.
\newblock Concurrency Computat.: Pract. Exper. 2016 (to appear).

\bibitem{Li:2010}
Li, G. and Gopalakrishnan, G.
\newblock {\em{Scalable SMT-based Verification of GPU Kernel Functions}}.
\newblock {FSE} 2010; 187--196.

\bibitem{Li:2012}
Li, G., Li, P., Sawaya, G., Gopalakrishnan, G., Ghosh, I., and Rajan, S.
\newblock {\em{GKLEE: Concolic Verification and Test Generation for GPUs.}}
\newblock {PPoPP} 2012; 215--224.

\bibitem{civl:2015}
Zheng, M., Rogers, M., Luo, Z., Dwyer, M., and Siegel, S.
\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
\newblock {ASE} 2015; 830--835.

\bibitem{Biere:2009}
Biere, A.
\newblock {\em{Bounded Model Checking. Handbook of Satisfiability}}. 
\newblock {IOS Press} 2009; 457--481.

\bibitem{Barrett:2009}
Barrett C, Sebastiani R, Seshia S, Tinelli C.
\newblock {\em{Satisfiability Modulo Theories. Handbook of Satisfiability}}.
\newblock {IOS Press} 2009; 825--885.

\bibitem{KahlonWG09}
Kahlon V, Wang C, Gupta A.
\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
\newblock {CAV} 2009; 398--413.

%\bibitem{sesa:2014}
%Li, P., Li, G., and Gopalakrishnan, G.
%\newblock {\em{Practical Symbolic Race Checking of {GPU} Programs}}.
%\newblock {SC} 2014; 179--190.

%\bibitem{Morse11}
%Morse, J., Cordeiro, L., Nicole, D., and Fischer, B.
%\newblock {\em{Context-Bounded Model Checking of LTL Properties for ANSI-C Software.}} 
%\newblock {SEFM} 2011; 302--317.

%\bibitem{Morse:15}
%Morse, J. and {\it et al.} %, Cordeiro, L., Nicole, D., and Fischer, B.
%\newblock {\em{Model checking LTL properties over ANSI-C programs with bounded traces.}} 
%\newblock {Software and System Modeling} 2015; 14(1): 65--81.

\bibitem{ramalho:2013}
Ramalho, M., Freitas, M., Sousa, F., Marques, H., Cordeiro, L., and Fischer, B.
\newblock {\em{SMT-Based Bounded Model Checking of C++ Programs}}.
\newblock {{ECBS}} 2013; 147--156.

\bibitem{monteiro:2015}
Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {Bounded Model Checking of C++ Programs Based on the Qt Framework}}. 
\newblock {GCCE} 2015; 179--447.
%\newblock {DOI=http://dx.doi.org/10.1109/GCCE.2015.7398699}}

\bibitem{garcia:2016}
Garcia, M., Monteiro, F., Cordeiro, L., and de Lima Filho, E. 
\newblock {\em {ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
\newblock {SPIN} 2016; 97--103.
%\newblock {DOI=http://dx.doi.org/10.1007/978-3-319-32582-8\_6}}

\bibitem{monteiro:2017}
Monteiro, F., Garcia, M., Cordeiro, L., and de Lima Filho, E.
\newblock {\em{Bounded Model Checking of C++ Programs based on the Qt Cross-Platform Framework}}.
\newblock Software Testing, Verification \& Reliability. 2017; ({\it to appear}).

\bibitem{posix:2008}
Institute of Electrical and Electronics Engineers, Inc.
\newblock {\em {IEEE Standard for Information Technology - Portable Operating System Interface (POSIX) Base Specifications, IEEE Std 1003.1-2008 (Revision of IEEE Std 1003.1-2004)}}.
\newblock {IEEE} 2008.

\bibitem{FIPS:2002}
Federal Information Processing Standard 180-2. 
\newblock {\em {Secure Hash Standard.}}
\newblock National Institute of Standards and Technology, 2002.

\bibitem{morse:2015}
Morse, J.
\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
\newblock {University of Southampton, PhD Thesis} 2015.

%\bibitem{cplusplus:2014}
%International Organization for Standardization.
%\newblock {\em {International Standard ISO/IEC 14882:2014(E) -- Programming Language C++}}.
%\newblock {Geneva, Switzerland}, 2014.

%\bibitem{cudaproguide:2015}
%NVIDIA.
%\newblock {\em {CUDA C Programming Guide}}.
%\newblock NVIDIA Corporation 2015.

\bibitem{toolkit:2015}
NVIDIA.
\newblock {\em {CUDA Toolkit Release Archive}}. 
\newblock {\tt https://developer.nvidia.com/cuda-toolkit-archive}, 2015.

\bibitem{microsoft:2012}
Microsoft Corporation.
\newblock {\em {C++ AMP sample projects for download (MSDN blog)}}. 
\newblock {\tt blogs.msdn.com/b/nativeconcurrency/archive/2012/01/30/c-amp-} {\tt sample-projects-for-download.aspx}, 2012.

\end{thebibliography}

\newpage

\section*{Required Metadata}
\label{}

\section*{Current executable software version}
\label{}

Ancillary data table required for sub version of the executable software.%: (x.1, x.2 etc.) kindly replace examples in right column with the correct information about your executables, and leave the left column as it is.

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{(executable) Software metadata description} & \textbf{Please fill in this column} \\
\hline
S1 & Current software version & 2.0 \\
\hline
S2 & Permanent link to executables of this version  & {\tt http://esbmc.org/gpu/} \\
\hline
S3 & Legal Software License & Apache v2.0 \\
\hline
S4 & Computing Operating System & Ubuntu Linux OS\\
\hline
S5 & Installation requirements \& dependencies & GNU Libtool; Automake; Flex \& Bison; Boost C++ Libraries; Multi-precision arithmetic library developers tools (libgmp3-dev package); SSL development libraries (libssl-dev package); CLang 3.8; LLDB 3.8; GNU C++ compiler (multilib files); libc6 and libc6-dev packages\\
\hline
S6 & Link to user manual & {\tt http://esbmc.org/gpu/}  \\
\hline
S7 & Support email for questions & {\tt lucas.cordeiro@cs.ox.ac.uk} \\
\hline
\end{tabular}
\caption{Software metadata (optional)}
\label{} 
\end{table}

\newpage

\section*{Current code version}
\label{}

Ancillary data table required for subversion of the codebase. %Kindly replace examples in right column with the correct information about your current code, and leave the left column as it is.

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this column} \\
\hline
C1 & Current code version & $v2.0$ \\
\hline
C2 & Permanent link to code/repository used for this code version & {\tt https://github.com/ssvlab/esb} {\tt mc-gpu} \\
\hline
C3 & Legal Code License   & GNU Public License \\
\hline
C4 & Code versioning system used & git \\
\hline
C5 & Software code languages, tools, and services used & C++ \\
\hline
C6 & Compilation requirements, operating environments \& dependencies & GNU Libtool; Automake; Flex \& Bison; Boost C++ Libraries; Multi-precision arithmetic library developers tools (libgmp3-dev package); SSL development libraries (libssl-dev package); CLang 3.8; LLDB 3.8; GNU C++ compiler (multilib files); libc6 and libc6-dev packages\\
\hline
C7 & Link to developer documentation & {\tt http://esbmc.org/gpu} \\
\hline
C8 & Support email for questions & {\tt lucas.cordeiro@cs.ox.ac.uk}\\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{} 
\end{table}

\end{document}
\endinput
%%
%% End of file `SoftwareX_article_template.tex'.

### 2017-05-11 Publications/Original Software Publications/SoftwareX/appendix.tex
uploading JSCP
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

% This is LLNCS.DEM the demonstration file of
% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science, version 1.1
\documentclass{llncs}
%
% Remove any % below to load the required packages
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{subfig}
\usepackage[american]{babel}
\usetikzlibrary{arrows,automata,positioning}
\usepackage{epstopdf}
\newtheorem{mydef}{Definition}
\newtheorem{mylem}{Lemma}
\usepackage{afterpage}

\renewcommand{\baselinestretch}{0.96}

\lstset{language=C,basicstyle=\small}
\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
\lstset{tabsize=2}
\lstset{firstnumber=1}
\lstset{frame=single}
\lstset{
  language={C},
  morekeywords={enquanto,se,fim,entao,senao,retorne,faca,assume,then,end-if,end,input,output,cudaMemcpyHostToDevice,cudaMemcpyDeviceToHost,\_\_global\_\_}
}
\hyphenation{port-a-ble}

\newcommand{\comment}[1]{}

\begin{document}

\title{ESBMC-GPU}
\subtitle{A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs}
%\author{Felipe~R.~Monteiro$^{1}$, Isabela~S.~Silva$^{1}$,\\Lucas~C.~Cordeiro$^{1,2}$, and Eddie~B.~de~Lima~Filho$^{1,3}$}

%\institute{$^{1}$Federal University of Amazonas, Brazil \\ 
%$^{2}$University of Oxford, United Kingdom \\
%$^{3}$ FPF Tech, Brazil}% \\

%
%\date{Received: date / Revised version: date}
% The correct dates will be entered by Springer
%

%\authorrunning{F.~R.~Monteiro et al.}
\titlerunning{ESBMC-GPU: A Context-Bounded Model Checking Tool to Verify CUDA Programs}

\maketitle

\begin{abstract}
The Compute Unified Device Architecture (CUDA) is a programming model used for exploring the advantages of Graphics Processing Unit (GPU) devices, through parallelization and specialized functions and features. However, as in other development platforms, errors may occur, due to traditional software creation processes, which may even compromise the execution of an entire system. In order to address such a problem, ESBMC-GPU was developed, as an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). In summary, ESBMC processes input code through ESBMC-GPU and an abstract representation of the standard CUDA libraries, with the goal of checking a set of desired properties. Experimental results showed that ESBMC-GPU was able to correctly verify $82$\% of the chosen benchmarks and also overcame other existing GPU verifiers.%, mainly due to their inability in detecting array out-of-bounds and data race violations.
\end{abstract}

%\keywords{GPU verification; formal methods; model checking; CUDA.}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Introduction} 
\label{sec:intro}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

The Compute Unified Device Architecture (CUDA) is a development framework that makes use of the architecture and processing power of Graphics Processing Units (GPUs)~\cite{cuda:2012}. Indeed, CUDA is also an Application Programming Interface (API), through which a GPU's parallelization scheme and tools can be accessed, with the goal of executing kernels. However, source code is still written by human programmers, which may result in arithmetic overflow, division by zero, and other violation types. In addition, given that CUDA allows parallelization, problems related to the latter can be identified, due to thread scheduling~\cite{betts:2012}.

In order to address that problem, an extension to the Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{cordeiro:2012} was developed, named as ESBMC-GPU~\cite{Pereira15,Pereira16,pereira:2016}, with the goal of verifying CUDA-based programs (available online at \texttt{www.esbmc-gpu.org}). ESBMC-GPU consists of an extension for parsing CUDA source code ({\it i.e.}, a front-end to ESBMC) and a CUDA operational model (COM), which is an abstract representation of the standard CUDA libraries ({\it i.e.}, the native API) that conservatively approximates their semantics.

A distinct difference of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC) allied to Satisfiability Modulo Theories (SMT), with explicit state-space exploration~\cite{cordeiro:2012}. Concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled. However, even with BMC, space-state exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}. As a consequence, redundant interleavings are eliminated, without ignoring a program's behavior.

Finally, existing GPU verifiers normally ignore some aspects related to memory leak, data transfer, and overflow, which are normally present in CUDA programs. The proposed approach, in turn, explicitly addresses them, through an accurate checking procedure, which even considers data exchange between main program and kernel. Of course, that results in higher verification times, but more errors can then be identified and later corrected, in another development cycle.

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%\subsection{Existing GPU Verifiers}
%\label{sec:related-work}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\textbf{Existing GPU Verifiers}. GPUVerify~\cite{betts:2012} is based on Synchronous Delayed Visibility, which focus on detecting data race and barrier divergence, while reducing kernel verification procedures to the analysis of sequential programs. Symbolic Executor with Static Analysis (SESA)~\cite{sesa:2014} and GPU + KLEE (GKLEE)~\cite{Li:2012} are concrete plus symbolic (\textit{concolic}) execution tools. Initially, GKLEE modeled every original thread, which was later improved by taking into account symmetry, and it considers both kernel and main function, while checking barrier synchronization and data race, among others. SESA combines equivalent flows, which results in great performance improvement, focus on data race detection, and does not verify the main function. Prover of User GPU Programs (PUG)~\cite{Li:2010} is based on SMT solvers and applies symbolic static analysis to each kernel, but requires user annotations. CIVL~\cite{civl:2015} generates an abstract syntax tree and supports analysis and transformations. Besides, it also verifies concurrent programs with partial order reduction, which makes it very similar to the solution proposed here. Given that its support to CUDA libraries is still under development, CIVL is an interesting approach for checking simple CUDA programs.

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Architecture and Implementation}
\label{sec:arch}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

ESBMC-GPU builds on ESBMC, which is an open source context-bounded model checker based on SMT solvers~\cite{cordeiro:2012} for ANSI-C/C$++$ programs. ESBMC-GPU adds four essential models to ESBMC, as described below.\\

\noindent \textbf{CUDA Operational Model.} An operational model for CUDA libraries, in conjunction with ESBMC, provides support to CUDA functionalities, as shown in Fig.~\ref{figure:arch}. Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013,garcia:2016}, consists in an abstract representation that reliably approximates the CUDA library's semantics. However, COM incorporates pre- and post-conditions as assertions into verification processes, as shown in Fig.~\ref{figure:om}, which enables ESBMC-GPU to verify specific properties (cf. Sec.~\ref{sec:features}). Indeed, COM allows the necessary control for performing code analysis, where both CUDA operation and knowledge for model checking its properties are available.

\begin{figure}[htb]
  \centering
  \includegraphics[width=5in]{figures/arch.pdf}
  \caption{Overview of ESBMC-GPU's architecture.}
  %\caption{Connecting the operational model for CUDA library to ESBMC architecture.}
  \label{figure:arch}
\end{figure}

ESBMC was designed to handle multi-threaded software, through the use of the Portable Operating System Interface (POSIX -- ISO/IEC 9945).
%~\cite{posix:2008}
Thus, ESBMC-GPU applies a combination between processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths. Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\
%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.

\begin{figure}[htb]
  \centering
  \includegraphics[width=4.5in]{figures/cuda-om.png}
  \caption{CUDA operational model development process.}
  \label{figure:om}
\end{figure}

\noindent \textbf{Two-threads Analysis.} Similarly to GPUVerify~\cite{betts:2012} and PUG~\cite{Li:2010}, ESBMC-GPU also reduces the number of threads, during the verification of CUDA programs, to two, considering a NVIDIA Fermi GPU architecture, in order to improve verification time and avoid the state-space explosion problem. In CUDA programs, whilst threads execute the same parametrized kernel, only two threads are necessary for conflict check. Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation.\\

\noindent \textbf{State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and reduce state space, based on SHA256 hashes.\\%, due to the low the likelihood regarding collisions. \\
%~\cite{FIPS:2002}

\noindent \textbf{Monotonic Partial Order Reduction.} MPOR is used to reduce the number of thread interleavings, by classifying transitions inside a program as dependent or independent. As a consequence, it is possible to determine whether interleaving pairs always lead to the same state and then remove duplicates in a reachability tree, without ignoring any aspect of a program's behavior.

%In addition to both techniques,

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Functionalities}
\label{sec:features}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

Through the integration of COM into ESBMC, ESBMC-GPU is able to analyze CUDA programs and verify the following properties:

\begin{itemize}
	\item \textbf{Data race.\ } ESBMC-GPU checks data race conditions, in order to detect if multiple threads perform unsynchronized access to shared-memory locations.

	\item \textbf{Pointer safety. } ESBMC-GPU also ensures that {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid.

	\item \textbf{Array bounds.\ } ESBMC-GPU performs array-bound checking, in order to ensure that any variable, used as an array index, is within known bounds.

	\item \textbf{Arithmetic under- and overflow.\ } ESBMC-GPU checks whether a sum or product exceeds the memory limits that a variable can handle.%, which can cause an error capable of spreading through the entire execution path.

	\item \textbf{Division by zero.\ } ESBMC-GPU analyzes whether denominators, in arithmetic expressions, lead to a division by zero.

	\item \textbf{User-specified assertions.\ } ESBMC-GPU considers all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
\end{itemize}

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Usage Aspects}
\label{subsec:usage}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

In order to verify CUDA code, users must slightly modify one specific parameter, in a given program. In particular, the call %``{\tt kernel<<< blocks, threads >>>(kernel-parameters)}'' must be changed to ``{\tt ESBMC\_verify\_kernel(<kernel>, <blocks>, <threads>, <kernel-parameters>)}'', where {\tt <kernel>} is the kernel to be verified, {\tt <blocks>} is the number of blocks, {\tt <threads>} is the number of threads, and {\tt <kernel-parameters>} is the kernel function's parameters. After the required amendments, users must call ``\texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}  {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>}'', where {\tt $<$file$>$.cu} is the CUDA program to be verified, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound among all threads, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-CUDA-OM$>$} is the location of the COM library. In addition, the flag {\tt --data-races-check} must be added at the end of the aforementioned command line, in order to check data races.

\begin{center}
\vspace{1 mm}
{\tt kernel<<< blocks, threads >>>(kernel-parameters)}
\vspace{1 mm}
\end{center}

\noindent must be transformed to the following ESBMC-GPU intrinsic function: called {\tt ESBMC\_verify\_kernel()}, which takes as arguments: the kernel to be verified, the number of blocks, the number of threads, and the kernel parameters, respectively.

\begin{center}
\vspace{1 mm}
\noindent {\tt ESBMC\_verify\_kernel(<kernel>, <blocks>,}

\noindent {\tt <threads>, <kernel-parameters>)},
\vspace{1 mm}
\end{center}

\noindent where {\tt <kernel>} is the kernel to be verified, {\tt <blocks>} is the number of blocks, {\tt <threads>} is the number of threads, and {\tt <kernel-parameters>} is the kernel function's parameters. After the required amendments, the user must call the ESBMC-GPU v$2.0$ command-line version, using

\begin{center}
\vspace{1 mm}
\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} % --force-malloc-success}

\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
\vspace{1 mm}
\end{center}

where {\tt $<$file$>$.cu} is the CUDA program to be verified, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound among all threads, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-CUDA-OM$>$} is the location of the COM library. In addition, the flag {\tt --data-races-check} must be added at the end of the aforementioned command line, in order to check data races.

In summary, if no bug is found, up to a $k$-depth unwinding and a $c$ context-bound, ESBMC-GPU reports {\tt VERIFICATION SUCCESSFUL}. Conversely, it reports {\tt VERIFICATION FAILED}, along with a counterexample, which contains all necessary information for identifying and reproducing the respective error. It is worth mentioning that counterexamples are not only recipients that contain the necessary information for reaching an error state, but it can also be used for implementing proper code correction, even automatically.

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Experimental Evaluation}
\label{sec:effectiveness}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

%In order to evaluate ESBMC-GPU's precision and performance, we have developed an automated test suite, which covers all basic functions that are commonly used by real CUDA-based applications. 
The chosen benchmark suite comprises $20$ CUDA kernels from NVIDIA GPU Computing SDK v$2$.$0$, $20$ CUDA kernels from Microsoft C$++$ AMP Sample Projects%~\cite{microsoft:2012}
, and $104$ CUDA-based programs that explore a wide range of CUDA functionalities~\cite{cuda:2012}. In summary, the chosen suite contains $47.4$\% bug-free and $52.6$\% buggy benchmarks, which tackle data race, arithmetic operations, pointer assignment, {\tt \_\_device\_\_} function calls, general ANSI-C functions ({\it e.g.}, \texttt{memset}), general CUDA functions ({\it e.g.}, \texttt{cudaMemcpy}), general libraries in CUDA ({\it e.g.}, \texttt{curand.h}), and the ability to work with variables, type modifiers ({\it e.g.}, \texttt{unsigned}), pointers, type definitions, and intrinsic CUDA variables ({\it e.g.}, \texttt{uint4}).  The present experiments answer two research questions:

%\textcolor{blue}{existe outra informacao sobre a test suite que seria interessante mencionar? falta eu colocar a porcentagem de testes com e sem erro, mas para isso preciso conversar com os membros da equipe}

%\textcolor{red}{seria possivel explicar mais a escolha dos benchmarks? quantos utilizam funcionalidades especificas cuda? quantos permitem avaliar data-race? quantos errados e quantos sem erros?}

\begin{enumerate}

%\item {\bf (sanity check)} Which results does ESBMC-GPU obtain upon verifying benchmarks from the automated test suite?
\item How accurate is ESBMC-GPU when verifying the chosen test suite?

%\textcolor{red}{na verdade, nao deveriamos responder o quao correto o nosso eh e nao apenas se ele eh melhor ou pior?}

%\item {\bf (compare software performance)} What is ESBMC-GPU performance when compared to GKLEE, GPUVerify, PUG, and CIVL?
\item How does ESBMC-GPU compare to other existing verifiers?

%\textcolor{red}{na verdade, uma pergunta seria: o quao ``adequado'' o noss eh? ele jah analisa coisas que os outros ignoram... seria bom colocar isso aqui pois ateh deixaria diferente do artigo}

\end{enumerate}

\noindent In order to answer both questions, all benchmarks were verified with $5$ GPU verifiers (ESBMC-GPU, GKLEE, GPUVerify, PUG, and CIVL), on an otherwise idle Intel Core i$7$-$4790$ CPU $3$.$60$ GHz, with $16$ GB of RAM and a Linux OS. All times were measured by the UNIX time command, in seconds. An overview of the experimental results is shown in Fig.~\ref{figure:experimental}, where {\it True} represents bug-free benchmarks, {\it False} represents buggy benchmarks, {\it Not supported} represents benchmarks that could not be verified, {\it Correct} represents the percentage of benchmarks correctly verified, and {\it Incorrect} represents the percentage of benchmarks incorrectly verified. As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $82\%$, while GKLEE, GPUVerify, PUG, and CIVL reported $70\%$, $57\%$, $35\%$, and $31\%$, respectively. More precisely, ESBMC-GPU support the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($60$\%).

{\bf Limitations.} ESBMC-GPU was unable to correctly verify $27$ benchmarks, which are related to constant memory access ($2$\%), the use of CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when used as kernel call arguments ($8.5$\%). In addition, ESBMC-GPU only reported $1$\% of false positive and $2$\% false negative results.

%\textcolor{red}{aqui ficou meio largado... o nosso eh bom para caralho e isso tem que ficar muito claro... comente mais as duas porcentagens que totalizam os 82\%, no ESBMC-GPU... e qual foi realmente o acerto... dos erros existentes, quantos acertamos?} \textcolor{blue}{Preciso da informacao de quantos estao com bugs, como mencionado acima...}
%In addition, ESBMC-GPU had $23$ benchmarks that were not supported. These are related to constant memory access ($3$), the use of CUDA's specific libraries (\textit{e.g., curand.h}) ($7$), and the use of pointers to functions, structures, and char type variables used as kernel call arguments ($13$).

{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $811$ seconds. Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools:  GKLEE ($128$ sec), GPUVerify ($147$ sec), PUG ($12$ sec), and CIVL ($158$ sec). This is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration. In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.

%\textcolor{red}{tudo bem... mas como melhorar? tem que mencionar que se tempo nao for um problema, isso eh completamente aceitavel... principalmente se for tudo automatico... o desenvolvedor vai deixar lah rolando, almoca e depois volta... entendeu?}

%\textcolor{red}{alem disso, eh bom mencionar quem eh mais adequado a umd ado tipo de erros... ganhamos em tudo ou um dos outros eh melhor em uma dada condicao?... seria bom separar erros por categoria}

\begin{figure}[htb]
  \centering
  \includegraphics[width=5in]{figures/experimental-results.pdf}
  \caption{Experimental evaluation of ESBMC-GPU against other verifiers.}
  \label{figure:experimental}
\end{figure}

\noindent \textbf{Availability of Data and Tools.} The performed experiments are based on a set of publicly available benchmarks. All benchmarks, tools, and results, associated with the current evaluation, are available on {\tt http://esbmc-gpu.org/}.

%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Conclusions and Future Work}
\label{sec:conc}
%=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

ESBMC-GPU is an SMT-based BMC tool that recognizes CUDA directives, further simplifies verification models, and provides fewer incorrect results, if compared with GKLEE, GPUVerify, PUG, and CIVL. In addition, it presents an improved ability to detect array out-of-bounds and data race violations. Future work aims to support stream interleaving and implement further techniques to reduce the number of thread interleavings by taking into account the GPU symmetry.

%\textcolor{red}{aqui eh bom mencionar numeros... quanto eh melhor? apenas resuma o que jah tem acima... e para melhorar a velocidade?}

%\noindent \textbf{Acknowledgements.} This paper is based on research sponsored by the Institute of Development and Technology (INdT) and by the National Council for Scientific and Technological Development (CNPq) under agreement number $475647/2013-0$.

\begin{thebibliography}{25}

\bibitem{cuda:2012}
Cheng, J., Grossman, M., and McKercher, T.
\newblock {\em {Professional CUDA C Programming}}.
\newblock {John Wiley and Sons, Inc.} 2014.

%\bibitem{gpu:2010}
%Kirk, D. and Hwu, W.
%\newblock {\em {Programming Massively Parallel Processors}}.
%\newblock Elsevier, 2010.

\bibitem{betts:2012}
Betts, A., Chong, N., Donaldson, A., Qadeer, S., and Thomson, P.
\newblock {\em{GPUVerify: A Verifier for {GPU} Kernels}}.
\newblock {OOPSLA} 2012; 113--132.

\bibitem{cordeiro:2012}
Cordeiro, L. and {\it et al.} %, Fischer, B., and Marques{-}Silva, J.
\newblock {\em{SMT-Based Bounded Model Checking for Embedded {ANSI-C} Software}}.
\newblock {IEEE} Trans. Software Eng. 2012; 38(4):957--974.

\bibitem{Pereira15}
Pereira, P., Albuquerque, H., Marques, H., Silva, I., Carvalho, C., Santos, V., Ferreira, R., and Cordeiro, L.
\newblock {\em{Verifica\c{c}\~ao de Kernels em Programas CUDA usando Bounded Model Checking.}} 
\newblock {WSCAD-SSC} 2015; 24--35.

\bibitem{Pereira16}
Pereira, P., Albuquerque, H., Marques, H., Silva, I., Carvalho, C., Santos, V., Ferreira, R., and Cordeiro, L.
\newblock {\em{Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.}} 
\newblock {SAC SVT track 2016; 1648--1653.}

\bibitem{pereira:2016}
Pereira, P., Albuquerque, H., Silva, I., Marques, H., Monteiro, F., Ferreira, R., and Cordeiro, L.
\newblock {\em{SMT-Based Context-Bounded Model Checking for CUDA Programs}}.
\newblock Concurrency Computat.: Pract. Exper. 2016; ({\it to appear}).

\bibitem{Li:2010}
Li, G. and Gopalakrishnan, G.
\newblock {\em{Scalable SMT-based Verification of GPU Kernel Functions}}.
\newblock {FSE} 2010; 187--196.

\bibitem{Li:2012}
Li, G., Li, P., Sawaya, G., Gopalakrishnan, G., Ghosh, I., and Rajan, S.
\newblock {\em{GKLEE: Concolic Verification and Test Generation for GPUs.}}
\newblock {PPoPP} 2012; 215--224.

\bibitem{civl:2015}
Zheng, M., Rogers, M., Luo, Z., Dwyer, M., and Siegel, S.
\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
\newblock {ASE} 2015; 830--835.

\bibitem{KahlonWG09}
Kahlon V, Wang C, Gupta A.
\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order
  Reduction Technique}}.
\newblock {CAV} 2009; 398--413.

\bibitem{sesa:2014}
Li, P., Li, G., and Gopalakrishnan, G.
\newblock {\em{Practical Symbolic Race Checking of {GPU} Programs}}.
\newblock {SC} 2014; 179--190.

%\bibitem{Biere:2009}
%Biere, A.
%\newblock {\em{Bounded Model Checking. Handbook of Satisfiability}}. 
%\newblock {IOS Press} 2009; 457--481.

%\bibitem{Barrett:2009}
%Barrett C, Sebastiani R, Seshia S, Tinelli C.
%\newblock {\em{ Satisfiability Modulo Theories. Handbook of Satisfiability}}.
%\newblock {IOS Press} 2009; 825--885.

%\bibitem{cordeiro:2011}
%Cordeiro, L. and Fischer, B.
%\newblock {\em{Verifying Multi-threaded Software using SMT-based Context-Bounded
%  Model Checking}}.
%\newblock {ICSE} 2011; 331--340.

%\bibitem{KahlonWG09}
%Kahlon V, Wang C, Gupta A.
%\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order
%  Reduction Technique}}.
%\newblock {CAV} 2009; 398--413.

%\bibitem{Morse11}
%Morse, J., Cordeiro, L., Nicole, D., and Fischer, B.
%\newblock {\em{Context-Bounded Model Checking of LTL Properties for ANSI-C Software.}} 
%\newblock {SEFM} 2011; 302--317.

%\bibitem{Morse:15}
%Morse, J. and {\it et al.} %, Cordeiro, L., Nicole, D., and Fischer, B.
%\newblock {\em{Model checking LTL properties over ANSI-C programs with bounded traces.}} 
%\newblock {Software and System Modeling} 2015; 14(1): 65--81.

\bibitem{ramalho:2013}
Ramalho, M., Freitas, M., Sousa, F., Marques, H., Cordeiro, L., and Fischer, B.
\newblock {\em{SMT-Based Bounded Model Checking of C++ Programs}}.
\newblock {{ECBS}} 2013; 147--156.

%\bibitem{monteiro:2015}
%Monteiro, F., Cordeiro, L., de Lima Filho, E. 
%\newblock {\em {Bounded Model Checking of C++ Programs Based on the Qt Framework}}. 
%\newblock {GCCE} 2015; 179--447.
%\newblock {DOI=http://dx.doi.org/10.1109/GCCE.2015.7398699}}

\bibitem{garcia:2016}
Garcia, M., Monteiro, F., Cordeiro, L., de Lima Filho, E. 
\newblock {\em {ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
\newblock {SPIN} 2016; 97--103.
%\newblock {DOI=http://dx.doi.org/10.1007/978-3-319-32582-8\_6}}

%\bibitem{posix:2008}
%Institute of Electrical and Electronics Engineers, Inc.
%\newblock {\em {IEEE Standard for Information Technology - Portable Operating System Interface (POSIX) Base Specifications, IEEE Std 1003.1-2008 (Revision of IEEE Std 1003.1-2004)}}.
%\newblock {IEEE} 2008.

%\bibitem{morse:2015}
%Morse, J.
%\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
%\newblock {University of Southampton, PhD Thesis} 2015.

%\bibitem{cudaproguide:2015}
%NVIDIA.
%\newblock {\em {CUDA C Programming Guide}}.
%\newblock NVIDIA Corporation 2015.

%\bibitem{toolkit:2015}
%NVIDIA.
%\newblock {\em {CUDA Toolkit Release Archive}}. 
%\newblock \url{https://developer.nvidia.com/cuda-toolkit-archive}, 2015.

%\bibitem{microsoft:2012}
%Microsoft Corporation.
%\newblock {\em {C++ AMP sample projects for download (MSDN blog)}}. 
%\newblock \url{blogs.msdn.com/b/nativeconcurrency/archive/2012/01/30/c-amp-sample-projects-for-download.aspx}, 2012.

%\bibitem{FIPS:2002}
%Federal Information Processing Standard 180-2. 
%\newblock {\em {Secure Hash Standard.}}
%\newblock National Institute of Standards and Technology, 2002.

\end{thebibliography}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent The presentation plan of our work consists mainly of three parts. First, a brief motivation about general bugs in CUDA-based programs and how our present approach can assist CUDA developers to detect those bugs. Second, an overview about the ESBMC-GPU implementation and its verification approach. Third, a quick demonstration of ESBMC-GPU command-line version, using CUDA kernels from the developed benchmark test suite.
\\
\begin{enumerate}
  \item Introduction and Motivation (20\%)
  \\
  \begin{itemize} 
     \item Brief introduction about CUDA, from the perspective of a parallel computing platform and API model:
        \begin{itemize}
		\item Present the motivation as described by NVIDIA to configure GPUs;
		\item Show initial applications in graphical processing in games applications (especially those that require high computational power);
		\item Demonstrate further applications to biomedicine, air traffic control, and weather simulation. \\
	\end{itemize}

     \item Typical Programming Errors in CUDA:
         \begin{itemize}
		\item Show that arithmetic under- and overflow, buffer overflow, pointer safety, and division by zero are also common source of errors in CUDA programs;
		\item Emphasize that data race conditions, shared memory, and barrier divergence play an important role in CUDA-based program verification.
			\begin{enumerate}
				\item Demonstrate that all those errors lead to incorrect results during the program execution; and
				\item They are hard to detect due to the parallel operations. \\
			\end{enumerate}
	\end{itemize}

     \item Ensure code correctness in (critical) GPU-based applications.
         \begin{itemize}
		\item Exploit SMT-based context-BMC to verify CUDA-based programs.
			\begin{enumerate}
				\item Present an operational model for the CUDA platform, which consists of pre- and post-conditions and simulation features;
				\item Apply SMT-based context-bounded model checking to CUDA-based programs using MPOR, state-hashing, and two-thread analysis;
				\item Show the experimental results of ESBMC-GPU if compared to other state-of-art software verifiers for CUDA programs.\\
			\end{enumerate}
	\end{itemize}

  \end{itemize}

  \item ESBMC-GPU Implementation Aspects (20\%)\\
  \begin{itemize}
     \item Describe all properties that are currently supported by the ESBMC-GPU tool, which are:
     	\begin{itemize}
		\item Data race;
		\item Pointer safety (enabled by default);
		\item Memory leaks;
		\item Array bounds (enabled by default);
		\item Arithmetic under- and overflow;
		\item Division by zero (enabled by default);
		\item User-specified assertions.\\
	\end{itemize}

     \item ESBMC-GPU implementation and architecture:
      \begin{itemize}
      	\item CUDA Operational model;

		\begin{figure}[htb]
  			\centering
  			\includegraphics[width=4in]{figures/demo-om.pdf}
  			\caption{CUDA operational model scheme.}
  			\label{figure:demo-om}
		\end{figure}

		\begin{enumerate}
			\item As shown in Fig.~\ref{figure:demo-om}, operational models add assertions to the verification process, in order to check for specific properties;
			\item Verification model adopts the CPU parallel processing based on the POSIX Threading Library; \\
		\end{enumerate}

	\item Two-threads Analysis;

	Fig.~\ref{figure:demo-two} demonstrates the thread representation, which accesses the GPU shared memory and is based on NVIDIA Fermi GPU architecture. First of all, the numbered white boxes represent the shared memory positions to be accessed, ${t_{r}}^n$ are the thread-readers, and ${t_{w}}^n$ are the thread-writers, where $n$ is the index of each thread. %it is worth noticed that the threads are executed on a streaming multiprocessor. 
	In the first case of Fig.~\ref{figure:demo-two} (\textit{i.e.}, $S_{1}$), the threads, whose kernel is correctly implemented, do not present a data race condition, once they access different memory positions. In the second case of Fig.~\ref{figure:demo-two} (\textit{i.e.}, $S_{2}$), the threads execution of a kernel results in a data race condition, where several threads read or write in the same memory position, at the same lock-step. Finally, the third case of Fig.~\ref{figure:demo-two} is quite similar to the situation occurred in $S_{2}$, however, instead of using multiple threads to analyze data-race conditions, ESBMC-GPU ensures that such errors can be detected through the analysis of the behavior of two threads that are operating over the same memory location ({\it i.e.}, $S_3$).

		\begin{figure}[htb]
  			\centering
  			\includegraphics[width=2in]{figures/two-thread-approach.pdf}
  			\caption{Two-threads analysis.}
  			\label{figure:demo-two}
		\end{figure}

Note that the two-threads analysis affects mostly the data race verification, where program states must be analyzed with respect to their possible thread interleavings, which lead to an execution order of statements that results in error. In our benchmarks, we observed that the consideration of solely two threads attains substantial improvement in performance, whilst keeping true incorrect results at low rates.\\

%In our benchmarks, we observed a substantial performance improvement, considering only two threads, while keeping true incorrect results at low rates.\\

	\item State Hashing;

	Our symbolic state hashing approach computes a summary for a particular state that has been explored, and then index the set of explored state to reduce the generation of redundant states. In particular, given any state computed during the symbolic execution of a specific CUDA kernel, we simply summarize it and efficiently look up whether it has been explored before, along a different computation path; if that happens during the ESBMC-GPU symbolic execution, then the current computation path does not need to be further explored in the reachability tree. So if we reach a state in ESBMC-GPU, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive), and all shared/local variables together with the program counters are similar to another explored node, then ESBMC-GPU just considers one identical node to be further explored, since reachability subtrees associated to identical nodes are also similar. \\

	\item Monotonic Partial Order Reduction;

		\begin{figure}[!]
			\begin{center}
    			\subfloat[Independent transitions\label{subfig-1:independent}]{%
      				\includegraphics[width=0.45\textwidth]{figures/demo-mpor-a-code.pdf}
    			}
    			\hfill
    			\subfloat[Dependent transitions\label{subfig-2:dependent}]{%
      				\includegraphics[width=0.45\textwidth]{figures/demo-mpor-b-code.pdf}
    			}
    			\caption{Code snippets of CUDA kernels.}
    			\label{figure:demo-mpor-codes}
		\end{center}
  		\end{figure}

	In order to understand MPOR algorithm, one can observe the code snippets shown in Fig.~\ref{figure:demo-mpor-codes}. In addition, Fig.~\ref{subfig-1:independent} shows the application of MPOR to a CUDA kernel, where the global variable $a$ is written in a position relative to the thread ID; in particular, this variable determines the program execution state. On the first interleaving, thread $t_{1}$ reaches $v_1$ resulting in $a=[0,0]$. Thread $t_{2}$ runs and reaches $v_2$ resulting in $a=[0,1]$. The MPOR algorithm then checks whether $v_2$ does not exist and whether transition $t_1 \rightarrow t_2$ is defined as independent. Both conditions do not hold, therefore transition $t_1 \rightarrow t_2$ is defined as dependent. ESBMC-GPU thus performs all possible interleavings, and the next execution starts with thread $t_2$ from $v_0$, which changes the array's content to $a=[0,1]$. Thread $t_1$ runs and reaches $v_4$, which is similar to $v_2$. MPOR algorithm then checks whether thread $t_1$ has reached a redundant state. As thread $t_1$ has only reached $v_1$, then the algorithm concludes that the transition $t_2 \rightarrow t_1$ is independent, disregarding it (represented in Fig.~\ref{subfig-1:independent} by dotted lines).

		\begin{figure}[!]
			\begin{center}
    			\subfloat[Independent transitions\label{subfig-1:independent}]{%
      				\includegraphics[width=0.45\textwidth]{figures/demo-mpor-a.pdf}
    			}
    			\hfill
    			\subfloat[Dependent transitions\label{subfig-2:dependent}]{%
      				\includegraphics[width=0.45\textwidth]{figures/demo-mpor-b.pdf}
    			}
    			\caption{MPOR applied to a kernel with independent and dependent transitions.}
    			\label{figure:demo-mpor}
		\end{center}
  		\end{figure}

Fig.~\ref{subfig-2:dependent} shows another example in which the thread execution results in different states. On the first interleaving, the thread execution $t_1$ reaches $v_1$ and the thread execution $t_2$ reaches $v_2$. Since the content of $a[1]$ is different in both reachability tree (RT) nodes, the condition does not hold. On the second interleaving, thread $t_2$ modifies the array $a$ to $a=[0,1]$ and thread $t_1$ accidentally writes to $a[2]$. MPOR algorithm then checks whether $v_4$ exists in the RT. Since the condition does not hold, transition $t_2 \rightarrow t_1$ is then defined as dependent. In this particular case, both interleavings are considered and those two thread execution orders result in an array out-of-bounds violation since the array $a$ is of length $2$ (\textit{e.g.}, $v_4$ in dotted lines).\\

      \end{itemize}
  \end{itemize}
  \item ESBMC-GPU Demonstration (60\%)
\\

%\textcolor{red}{Felipe: There is too much repetition here w.r.t. section 4 Usage Aspects. Can you please focus your description based on a running example from the Microsoft suite? then you can show the code and how we use ESBMC-GPU to verify that program.}
  	%\begin{figure}[htb]
  	%	\centering
  	%	\includegraphics[width=4in]{figures/verification-flow.pdf}
  	%	\caption{ESBMC-GPU's verification flow.}
  	%	\label{figure:verification-flow}
	%\end{figure}

%\begin{figure} [htb]
%\centering
%\begin{minipage}{\textwidth}
%\begin{lstlisting}
%#include <call_kernel.h>
%#include <stdio.h>
%#include <stdlib.h>
%#include <stdio.h>
%#include <cuda.h>
%#include <cuda_runtime_api.h>
%#include <assert.h>
%__global__ void Asum(int *a, int *b, int *c){
%	*c = *a + *b;
%}
%int main(void){
%	int a, b, c;
%	int *dev_a, *dev_b, *dev_c;
%	int size = sizeof(int);
%	cudaMalloc((void**)&dev_a, size);
%	cudaMalloc((void**)&dev_b, size);
%	cudaMalloc((void**)&dev_c, size);
%	a = 2;
%	b = 7;
%	c = 8;	
%	cudaMemcpy(dev_a, &a, size, cudaMemcpyHostToDevice);
%	cudaMemcpy(dev_b, &b, size, cudaMemcpyHostToDevice);
%	// Asum<<<1,2>>>(dev_a, dev_b, dev_c);
%	ESBMC_verify_kernel(Asum, 1, 2, dev_a, dev_b, dev_c);
%	cudaMemcpy(&c, dev_c, size, cudaMemcpyDeviceToHost);
%	assert(c != a + b);	
%	cudaFree(dev_a);
%	cudaFree(dev_b);
%	cudaFree(dev_c);
%	return 0;
%}
%\end{lstlisting}
%\end{minipage}
%\caption{Illustrative CUDA code example.}
%\label{fig:illustrative}
%\end{figure}

\begin{figure} [htb]
\centering
\begin{minipage}{\textwidth}
\begin{lstlisting}
#include <call_kernel.h>
#include <stdio.h>
#include <stdlib.h>
#include <stdio.h>
#include <cuda.h>
#include <cuda_runtime_api.h>
#include <assert.h>

#define BLOCKS 1
#define THREADS 2

__global__ void kernel(int *A) {
	A[threadIdx.x + 1] = threadIdx.x;
}

int main(){
	int *a;
	int *dev_a;
	int size = THREADS*sizeof(int);
	a = (int*)malloc(size);
	cudaMalloc((void**)&dev_a, size);
	for (int i = 0; i < THREADS; i++)
		a[i] = 0;
	cudaMemcpy(dev_a, a, size, cudaMemcpyHostToDevice);
	// kernel<<<BLOCKS,THREADS>>>(dev_a);
	ESBMC_verify_kernel(kernel, BLOCKS, THREADS, dev_a);
	for (int i = 0; i < THREADS; i++)
		assert(a[i]==i);	
	cudaFree(dev_a);
	return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Illustrative CUDA code example.}
\label{fig:illustrative-example}
\end{figure}

In this part, we will demonstrate the ESBMC-GPU usage, based on the CUDA program shown in Fig.~\ref{fig:illustrative-example}. First of all, the user must replace the default kernel call (line $25$) to an intrinsic function of ESBMC-GPU (line $26$), as explained in Sec.~\ref{subsec:usage}. Then, the modified CUDA program can be passed to the command-line version of ESBMC-GPU as follows:

\begin{center}
%\vspace{1 mm}
\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}

\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
%\vspace{1 mm}
\end{center}

%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, respectively. Furthermore, its kernel (lines $8$ to $10$) stores in the variable $c$ the result of the sum of $a$ and $b$, which are passed as input arguments. However, there is a mistake in the assertion (line $26$), as $a + b$ must be equal to $c$, which is correctly detected by ESBMC-GPU detects an assert violation.

As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, and a kernel (lines $12$ to $14$), which assigns thread's index values to an array passed as an input argument. Indeed, its goal is to instantiate array positions according to the thread index. However, there is a mistake in the array index, as the value $1$ is accidentally added to the thread index (line $13$). As shown in the main function, array positions are assigned with value $0$ (line $23$), and after the kernel call (line $26$), it is expected that $a[0] == 0$ and $a[1] == 1$.

In this example, ESBMC-GPU detects an array out-of-bounds violation. Indeed, this CUDA-based program retrieves a memory region that has not been previously allocated, thus, when {\tt threadIdx.x = 1}, the program tries to access the position $a[2]$. Importantly, the {\tt cudaMalloc()} function operational model has a precondition that checks if the memory size to be allocated is greater than zero. In addition, an assertion checks if the result matches to the expected postcondition (line $28$). Therefore, the verification of this program through ESBMC-GPU produces $34$ successful and $6$ failed interleavings. For instance, one possible failed interleaving is represented by the threads executions $t_0 : a[1] = 0$; $t_1 : a[2] = 1$, where $a[2] = 1$ represents an incorrect access to the array index $a$.

It is worth noticing that ESBMC-GPU and GKLEE are able to detect this array out-of-bounds violation, but GPUVerify and PUG fail, as they report a true incorrect result (missed bug). Furthermore, ESBMC-GPU has the following additional command-line options: \\

%\texttt{--no-assertions}: to ignore assertions; \\
%\texttt{--no-bounds-check}: to skip array bounds check; \\
%\texttt{--no-div-by-zero-check}: to skip division by zero check; \\
%\texttt{--no-pointer-check}: to skip pointer check; \\
\texttt{--memory-leak-check}: to enable memory leak check; \\
\texttt{--overflow-check}: to enable arithmetic over- and underflow check; \\
%\texttt{--deadlock-check}: to enable global and local deadlock check with mutex; \\
\texttt{--data-races-check}: to enable data races check; \\
%\texttt{--lock-order-check}: to enable for lock acquisition ordering check; \\
%\texttt{--atomicity-check}: to enable atomicity check at visible assignments; \\
%\texttt{--memlimit}: to configure memory limit, of form ``100MB'' or ``2GB''; \\
%\texttt{--timeout}: to configure time limit, integer followed by \{s, m, or h\}; \\
%\texttt{--force-malloc-success}: to consider that there is always enough memory available in the device; \\
\texttt{-DGPU\_threads=<$number$>}: to define the total number of threads by kernel. \\

The tool, operational model, and all experiments results are available for downloading at: \texttt{http://esbmc-gpu.org/}.\\

Video demonstration at \texttt{https://www.youtube.com/watch?v=Q7pOAr6zY3c}.
\end{enumerate}

\end{document}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/not-supported/assertion/077_value_in_assert/main.cu

@@ -1,0 +1,36 @@
+//pass: o caso deve passar por causa do assert que sempre é verdadeiro
+//--blockDim=16 --gridDim=16 --no-inline
+//a = 12
+//b = 36
+//c = 48
+
+#include <stdio.h>
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <assert.h>
+#include <stdlib.h>
+#define N 2//16
+
+__global__ void example(unsigned a, unsigned b, unsigned c) {
+
+    //__requires(a == 12);
+    //__requires(b == 36);
+
+    //__assert(a + b != c);
+	assert((a+b) != c);
+
+}
+
+int main(){
+
+	unsigned a, b, c;
+
+	a=12;
+	b=36;
+//	c=48;
+
+	example<<<N,N>>>(a,b,c);
+	//ESBMC_verify_kernel_u(example,1,N,a,b,c);
+
+	cudaThreadSynchronize();
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/not-supported/thread-size/094_intragroup_scan/main.cu

@@ -1,0 +1,78 @@
+//pass
+//--blockDim=512 --gridDim=1 --warp-sync=32 --no-inline
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 32//128//512
+
+__device__ static __attribute__((always_inline)) void scan_warp (int* A)
+{
+	unsigned int tid = threadIdx.x;
+	unsigned int lane = tid % 32;
+
+	if (lane >= 1) A[tid] = A[tid - 1] + A[tid]; // this conditional is always true!!
+	if (lane >= 2) A[tid] = A[tid - 2] + A[tid];
+	if (lane >= 4) A[tid] = A[tid - 4] + A[tid];
+	if (lane >= 8) A[tid] = A[tid - 8] + A[tid];
+	if (lane >= 16) A[tid] = A[tid - 16] + A[tid];
+}
+
+__global__ void scan (int* A)
+{
+	unsigned int tid = threadIdx.x;
+	unsigned int lane = tid % 32;
+
+	int temp [32] = {0};
+	scan_warp(A);
+	__syncthreads();
+
+	if (lane == 31)	// ?????????
+		temp[tid / 32] = A[tid];
+	__syncthreads();
+
+	if (tid / 32 == 0)
+		scan_warp(temp);
+	__syncthreads();
+
+	A[tid] += temp[tid/32];
+
+}
+
+
+int main(){
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = i;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	printf("old a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	scan<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(scan, 1, N, dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/not-supported/thread-size/096_shuffle_fail/main.cu

@@ -1,0 +1,56 @@
+//data race
+//--blockDim=512 --gridDim=1 --warp-sync=32 --no-inline
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+#define N 4//512
+
+__global__ void shuffle (int* A)
+{
+	int tid = threadIdx.x;
+	int warp = tid / 2;//32;
+	int* B = A + (warp*2);//32);
+	A[tid] = B[(tid + 1)%2];//32];
+}
+
+int main() {
+	int *a;
+	int *b;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(N*size);
+	b = (int*)malloc(N*size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = i;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	printf("a:  ");
+
+	for (int i = 0; i < N; i++)
+		printf("%d        ", a[i]);
+
+	shuffle<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(shuffle, 1, N, dev_a);
+
+	cudaMemcpy(b,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nFunction Results:\n   ");
+
+	for (int i = 0; i < N; i++){
+		printf("%d        ", b[i]);
+		assert(b[i]==a[i+1]);
+	}
+	free(a);free(b);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/not-supported/thread-size/103_misc1/main.cu

@@ -1,0 +1,42 @@
+//pass
+//--blockDim=512 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 512
+
+__global__ void helloCUDA(int x)
+{
+///    __requires(x == 143);
+    __shared__ float S[256*32];
+    __shared__ float F[256];
+
+    unsigned int idx;
+
+    //initialise data on shared memory
+    for(int i = 0;
+   //         __invariant(__implies(__write(S), ((__write_offset_bytes(S)/sizeof(float)) % blockDim.x) == threadIdx.x)),
+            i < x;
+            i += (blockDim.x/32)) /* translate: i = 0; i < 143; i+=16 , total de iterações: 8*/
+		
+    {
+        if((i+(threadIdx.x/32)) < x){
+            idx = (i+(threadIdx.x/32))*32+(threadIdx.x%32);
+            S[idx] = F[i+(threadIdx.x/32)];
+        }
+    }
+
+}
+int main (){
+
+	helloCUDA <<<1,N>>>(143);
+	//ESBMC_verify_kernel(helloCUDA,1,N, 143);	
+	
+	cudaThreadSynchronize();
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/not-supported/verify-kernel/113_curand_race_fail/main.cu

@@ -1,0 +1,49 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=2 --gridDim=1 --no-inline
+//Write by thread .+kernel\.cu:8:21:
+
+#include <cuda.h>
+#include <curand.h>
+#include <curand_kernel.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 4
+
+__global__ void curand_test(curandState *state, float *A) { // test: replace curandState for curandStateXORWOW_t
+   A[threadIdx.x] = curand_uniform(state);
+}
+
+int main(){
+	float *a;
+	float *dev_a;
+	curandState *dev_state; 
+	
+	int size = N*sizeof(float);
+
+	a = (float*)malloc(size);
+	cudaMalloc ((void**) &dev_a, size);
+
+	printf("old a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%f	", a[i]);
+
+	cudaMalloc ( (void**) &dev_state, N*sizeof( curandState ) );
+
+	curand_test<<<1,N>>>(dev_state, dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a:  ");
+	for (int i = 0; i < N; i++) {
+		printf("%f	", a[i]);
+		assert(a[i] == a[i+1]);
+	}
+
+	free(a);
+	cudaFree(&dev_a);
+	cudaFree(&dev_state);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/assertion/009_testinterprocedural3/main.cu

@@ -1,0 +1,98 @@
+//fail
+//--blockDim=64 --gridDim=64 --no-inline
+#include "cuda.h"
+#include <stdio.h>
+#include <cuda_runtime_api.h>
+
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 1
+
+__device__ void baz (int p []){
+    int a;
+
+    p = &a;
+}
+
+__device__ void bar (int *p){
+
+    int a = 2;
+
+    p = &a;
+}
+
+
+__global__ void foo (int* p, int* q){
+
+    __shared__ int sharedArr  [100];
+
+    __shared__ int sharedArr2 [50];
+
+    bar(p);
+
+    baz (sharedArr);
+
+    bar(q);
+
+    if (*q){
+        baz(sharedArr2);
+    }
+
+    //*p = 23; *q = 23; // remove this comment to see that the __device__ function does not work
+}
+
+int main(){
+	int *a, *b;
+	int *dev_a, *dev_b;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+	cudaMalloc((void**)&dev_b, size);
+
+	a = (int*)malloc(size);
+	b = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	for (int i = 0; i < N; i++)
+		b[i] = 1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+	cudaMemcpy(dev_b,b,size, cudaMemcpyHostToDevice);
+
+	printf("a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	printf("\nb:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", b[i]);
+
+	foo<<<1,2>>>(dev_a,dev_b);
+	//ESBMC_verify_kernel(foo,1,2,dev_a,dev_b);
+	
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+	cudaMemcpy(b,dev_b,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a and new b:\n");
+
+	printf("a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	printf("\nb:  ");
+	for (int i = 0; i < N; i++)
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/data-racer/011_value_causing_race/main.cu

@@ -1,0 +1,48 @@
+//xfail:data-race
+// Write by thread 0
+// Write by thread 1
+// x = 1
+
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+#include <cuda.h>
+
+#define N 2//512
+
+__global__ void example(float * A, int x) {
+
+	//__requires(x = 1); // x deve ser a diferença entre o limite do if1 e do if2
+
+    if(threadIdx.x == 0) {
+        A[threadIdx.x + x] = threadIdx.x; //A[1] = 0;
+    }
+
+    if(threadIdx.x == 1) {
+        A[threadIdx.x] = threadIdx.x; //A[1] = 1;
+   }
+}
+
+int main() {
+	int c=1;
+	float *a;
+	float *dev_a;
+
+	a = (float*)malloc(N*sizeof(float));
+
+	cudaMalloc((void**)&dev_a, N*sizeof(float));
+
+	cudaMemcpy(dev_a, a, N*sizeof(float), cudaMemcpyHostToDevice);
+
+		example<<<1, N>>>(dev_a, c);
+		//ESBMC_verify_kernel_fuintint(example,1, N,dev_a, c);
+
+	cudaMemcpy(a, dev_a, N*sizeof(float), cudaMemcpyDeviceToHost);
+
+	assert(a[1] == 0 || a[1] == 1);	
+
+	free(a);
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/data-racer/012_add_tid/main.cu

@@ -1,0 +1,73 @@
+//xfail:BOOGIE_ERROR:data race
+//--blockDim=2 --gridDim=1 --no-inline
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <assert.h>
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <sm_atomic_functions.h>
+
+#define N 2
+
+__global__ void race_test (unsigned int* i, int* A)
+{
+	int tid = threadIdx.x;
+	int j = atomicAdd(i,tid);
+	A[j] = tid;
+}
+
+int main(){
+
+	unsigned int *i;
+	int *A;
+	unsigned int *dev_i;
+	int *dev_A;
+
+	A = (int*)malloc(N*sizeof(int));
+
+	for (int t = 0; t < N; ++t){
+		A[t] = 0;
+		printf(" %d  ", A[t]);
+	}
+
+	i = (unsigned int*)malloc(sizeof(unsigned int));
+	
+	*i = 0;
+
+	cudaMalloc((void**)&dev_A, N*sizeof(int));
+	cudaMalloc((void**)&dev_i, sizeof(unsigned int));
+
+	cudaMemcpy(dev_A, A, N*sizeof(int), cudaMemcpyHostToDevice);
+	cudaMemcpy(dev_i, i, sizeof(unsigned int), cudaMemcpyHostToDevice);
+
+		race_test<<<1,N>>>(dev_i, dev_A);
+		//ESBMC_verify_kernel_u(race_test,1,N,dev_i,dev_A);
+		
+	cudaMemcpy(A, dev_A, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+/*	printf("\n");
+
+	for (int t=0; t<N; t++){
+		printf ("A[%d]=%d ", t, A[t]);
+	}
+
+	printf("\n\n");
+*/	
+	int tid = 0;
+	for (int t=0; t<N;){
+		printf ("A[%d]=%d ", t, A[t]);
+		//assert(A[t] == tid); // A[t] == x , where x=[0,N]
+		assert(A[t] == 0 || A[t] == 1); // A[t] == x , where x=[0,N-1]		
+		tid++;
+		t = t + tid;
+	}
+
+	free(A);
+	free(i);
+	cudaFree(dev_A);
+	cudaFree(dev_i);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/data-racer/017_equality_abstraction_issue/main.cu

@@ -1,0 +1,44 @@
+//xfail:BOOGIE_ERROR
+//--warp-sync=32 --blockDim=32 --gridDim=1 --equality-abstraction --no-inline
+//kernel.cu:10
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+#define N 2//32
+
+__global__ void foo(int * A) {
+    A[0] = 1;
+    A[1] = 1;
+    A[threadIdx.x] = 0;
+//__assert(A[0] == 1 | A[1] == 1 | A[2] == 1);
+}
+
+int main(){
+
+	int *b;
+	int *dev_b;
+
+	b = (int*)malloc(N*sizeof(int));
+
+	for (int i = 0; i < N; ++i){
+		b[i] = 2;
+	}
+
+	cudaMalloc((void**)&dev_b, N*sizeof(int));
+	cudaMemcpy(dev_b, b, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_b);
+	//ESBMC_verify_kernel(foo, 1, N, dev_b);
+
+	cudaMemcpy(b, dev_b, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; ++i){
+		assert(b[i] == 0 || b[i] == 1);
+	}
+
+	free(b);
+	cudaFree(dev_b);
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/data-racer/085_shared_int/main.cu

@@ -1,0 +1,24 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=64 --gridDim=64 --no-inline
+//
+#include "cuda.h"
+#include <assert.h>
+#include <stdlib.h>
+#define N dim*dim
+#define dim 2
+
+__global__ void foo() {
+
+  __shared__ int a;
+
+  a = threadIdx.x;
+}
+
+int main(){
+
+	foo <<<N,N>>> ();
+	//ESBMC_verify_kernel(foo, dim,dim);
+
+	cudaThreadSynchronize();
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/data-racer/097_loop_unwind/main.cu

@@ -1,0 +1,51 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=512 --gridDim=64 --loop-unwind=2 --no-inline
+//kernel.cu: error: possible write-write race on B
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 2//512
+
+extern "C" {
+
+__global__ void helloCUDA(float *A)
+{
+    __shared__ float B[256];
+
+    for(int i = 0; i < N*2; i ++) {
+        B[i] = A[i];
+    }
+}
+
+}
+
+int main() {
+
+	float *A;
+	float *dev_A;
+
+	float size= N*sizeof(float);
+
+	A=(float*)malloc(size);
+
+	cudaMalloc((void**)&dev_A, size);
+
+	for (int i = 0; i < N; i++)
+		A[i] = 5;
+
+
+	cudaMemcpy(dev_A, A, size, cudaMemcpyHostToDevice);
+	
+	helloCUDA<<<64,N>>>(dev_A);
+	//ESBMC_verify_kernel(helloCUDA, 1, N, dev_A);
+
+	cudaMemcpy(A, dev_A, size, cudaMemcpyDeviceToHost);
+
+	cudaFree(dev_A);
+	free(A);
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/049_pass/main.cu

@@ -1,0 +1,42 @@
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+#include "cuda.h"
+
+#define N 2
+
+__global__ void foo(int* p) {
+    p[threadIdx.x] = 2;
+    __syncthreads();
+}
+
+int main(){
+
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(N*size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(foo,1,N,dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++){
+		assert(a[i]==2);
+	}
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/056_fail/main.cu

@@ -1,0 +1,47 @@
+//fail: assertion
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+#define N 2//1024
+
+__global__ void foo(float *A, int sz) {
+	assert(sz == blockDim.x);
+	for(int i = threadIdx.x; i < 1*sz; i += sz) {
+		A[i] *= 2.0f;
+  }
+}
+
+int main() {
+
+    float *a;
+    float *b;
+    float *dev_a;
+    int size = N*sizeof(float);
+
+    cudaMalloc((void**)&dev_a, size);
+
+    a = (float*)malloc(N*size);
+    b = (float*)malloc(N*size);
+
+    for (int i = 0; i < N; i++)
+        a[i] = i;
+
+    cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	     	foo<<<1,N>>>(dev_a, N);
+		//ESBMC_verify_kernel_fuintint(foo, 1, N, dev_a, N);
+
+    cudaMemcpy(b,dev_a,size,cudaMemcpyDeviceToHost);
+
+    for (int i = 0; i < N; i++){
+        assert(b[i]!=2*a[i]);
+    }
+
+    free(a); free(b);
+    cudaFree(dev_a);
+
+    return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/062_struct/main.cu

@@ -1,0 +1,50 @@
+//pass
+//--blockDim=64 --gridDim=64 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define DIM 2 //64
+#define N 2// DIM*DIM
+
+typedef struct {
+  float x,y,z,w;
+} myfloat4;
+
+__global__ void k(float * i0) {
+  myfloat4 f4;
+  f4.x = 2;
+  i0[threadIdx.x + blockDim.x*blockIdx.x] = f4.x;
+}
+
+int main(){
+	float *a;
+	float *dev_a;
+	int size = N*sizeof(float);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (float*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 5;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+		k<<<DIM,DIM>>>(dev_a);
+		//ESBMC_verify_kernel_f(k,1,N,dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++) {
+		assert(a[i] == 2);
+	}
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/071_atomicMin/main.cu

@@ -1,0 +1,63 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 2 //1024
+
+__global__ void definitions (int* A, unsigned int* B, unsigned long long int* C)
+{
+  atomicMin(A,10);
+
+  atomicMin(B,1);
+
+  atomicMin(C,5);
+}
+
+int main (){
+
+	int a = 5;
+	int *dev_a;
+
+	cudaMalloc ((void**) &dev_a, sizeof(int));
+
+	cudaMemcpy(dev_a, &a, sizeof(int),cudaMemcpyHostToDevice);
+
+	unsigned int b = 5;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	unsigned long long int c = 5;
+	unsigned long long int *dev_c;
+
+	cudaMalloc ((void**) &dev_c, sizeof(unsigned long long int));
+
+	cudaMemcpy(dev_c, &c, sizeof(unsigned long long int),cudaMemcpyHostToDevice);
+
+	definitions <<<1,N>>>(dev_a,dev_b,dev_c);
+	//ESBMC_verify_kernel(definitions,1,N,dev_a,dev_b,dev_c);
+
+	cudaMemcpy(&a,dev_a,sizeof(int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&b,dev_b,sizeof(unsigned int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&c,dev_c,sizeof(unsigned long long int),cudaMemcpyDeviceToHost);
+
+	printf("A: %d\n", a);
+	printf("B: %u\n", b);
+	printf("C: %u\n", c);
+
+	assert(a==5);
+	assert(b==1);
+	assert(c==5);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+	cudaFree(dev_c);
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/074_atomicXor/main.cu

@@ -1,0 +1,63 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#define N 2 //1024
+
+__global__ void definitions (int* A, unsigned int* B, unsigned long long int* C)
+{
+  atomicXor(A,10);//1010 xor 0101 = 1111 /*xor looks for distinct bits*/
+
+  atomicXor(B,7);//0111 xor 0101 = 0010
+
+  atomicXor(C,5);//0101 xor 0101 = 0000
+}
+
+int main (){
+
+	int a = 5;
+	int *dev_a;
+
+	cudaMalloc ((void**) &dev_a, sizeof(int));
+
+	cudaMemcpy(dev_a, &a, sizeof(int),cudaMemcpyHostToDevice);
+
+	unsigned int b = 5;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	unsigned long long int c = 5;
+	unsigned long long int *dev_c;
+
+	cudaMalloc ((void**) &dev_c, sizeof(unsigned long long int));
+
+	cudaMemcpy(dev_c, &c, sizeof(unsigned long long int),cudaMemcpyHostToDevice);
+
+	definitions <<<1,N>>>(dev_a,dev_b,dev_c);
+	//ESBMC_verify_kernel(definitions,1,N,dev_a,dev_b,dev_c);
+
+	cudaMemcpy(&a,dev_a,sizeof(int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&b,dev_b,sizeof(unsigned int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&c,dev_c,sizeof(unsigned long long int),cudaMemcpyDeviceToHost);
+
+	printf("A: %d\n", a);
+	printf("B: %u\n", b);
+	printf("C: %u\n", c);
+
+	assert(a==5);
+	assert(b==5);
+	assert(c==5);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+	cudaFree(dev_c);
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/082_threadfence/main.cu

@@ -1,0 +1,19 @@
+//pass
+//--blockDim=32 --gridDim=2
+
+#include <cuda.h>
+#include <stdlib.h>
+
+__global__ void foo() {
+    __threadfence();
+}
+
+int main(){
+
+	foo<<<1,2>>>();
+	//ESBMC_verify_kernel(foo,1,2);
+
+	cudaThreadSynchronize();
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/088_fail/main.cu

@@ -1,0 +1,62 @@
+//fail
+//--blockDim=1024 --gridDim=1024 --no-inline
+
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <math_functions.h>
+
+#define DIM 2 //1024 in the future
+#define N 2//DIM*DIM
+
+__global__ void mul24_test (int* A, int* B)
+{
+  int idxa          = __mul24(blockIdx.x, blockDim.x) + threadIdx.x;
+  unsigned int idxb = __umul24(blockIdx.x, blockDim.x) + threadIdx.x;
+
+  A[idxa] = idxa;
+  B[idxb] = idxa;
+}
+
+int main (){
+	int *a, *b;
+	int *dev_a, *dev_b;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+	cudaMalloc((void**)&dev_b, size);
+
+	a = (int*)malloc(size);
+	b = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	for (int i = 0; i < N; i++)
+		b[i] = 1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+	cudaMemcpy(dev_b,b,size, cudaMemcpyHostToDevice);
+
+	mul24_test<<<DIM,DIM>>>(dev_a,dev_b);
+	//ESBMC_verify_kernel(mul24_test,1,N,dev_a,dev_b);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+	cudaMemcpy(b,dev_b,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++)
+		assert (a[i] != i);	
+
+	for (int i = 0; i < N; i++) 
+		assert (b[i] != i);	
+
+	free(a); free(b);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/CIVL/supported/pass/088_pass/main.cu

@@ -1,0 +1,62 @@
+//pass
+//--blockDim=1024 --gridDim=1024 --no-inline
+
+#include <stdio.h>
+#include <assert.h>
+#include <stdlib.h>
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <math_functions.h>
+
+#define DIM 2 //1024 in the future
+#define N 2//DIM*DIM
+
+__global__ void mul24_test (int* A, int* B)
+{
+  int idxa          = __mul24(blockIdx.x, blockDim.x) + threadIdx.x;
+  unsigned int idxb = __umul24(blockIdx.x, blockDim.x) + threadIdx.x;
+
+  A[idxa] = idxa;
+  B[idxb] = idxa;
+}
+
+int main (){
+	int *a, *b;
+	int *dev_a, *dev_b;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+	cudaMalloc((void**)&dev_b, size);
+
+	a = (int*)malloc(size);
+	b = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	for (int i = 0; i < N; i++)
+		b[i] = 1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+	cudaMemcpy(dev_b,b,size, cudaMemcpyHostToDevice);
+
+	mul24_test<<<DIM,DIM>>>(dev_a,dev_b);
+	//ESBMC_verify_kernel(mul24_test,1,N,dev_a,dev_b);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+	cudaMemcpy(b,dev_b,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++)
+		assert (a[i] == i);	
+
+	for (int i = 0; i < N; i++) 
+		assert (b[i] == i);	
+
+	free(a); free(b);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/not-supported/cudaMallocToFloat/083_definitions/main.cu

@@ -1,0 +1,103 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <sm_atomic_functions.h>
+#include <stdio.h>
+
+#define N 1 //1024
+
+__global__ void definitions (int* A, unsigned int* B, unsigned long long int* C, float* D)
+{
+	assert(*D == 0.0);
+	//assert(*D == 0.0 || *D == 5);
+	assert(*D == 5.0);
+/**/	atomicAdd(A,10);
+	atomicSub(A,10);
+	atomicExch(A,10);
+	atomicMin(A,10);
+	atomicMax(A,10);
+	atomicAnd(A,10);
+	atomicOr(A,10);
+	atomicXor(A,10);
+  	atomicCAS(A,10,11);
+
+/**/	atomicAdd(B,10);
+	atomicSub(B,10);
+	atomicExch(B,10);
+	atomicMin(B,10);
+	atomicMax(B,10);
+	atomicAnd(B,10);
+	atomicOr(B,10);
+	atomicXor(B,10);
+	atomicInc(B,10);
+	atomicDec(B,10);
+  	atomicCAS(B,10,11);
+
+/**/	atomicAdd(C,10);
+	atomicExch(C,10);
+	atomicMin(C,10);
+	atomicMax(C,10);
+	atomicAnd(C,10);
+	atomicOr(C,10);
+	atomicXor(C,10);
+  	atomicCAS(C,10,11);
+
+	atomicAdd(D,10.0);
+	atomicExch(D,10.0);
+}
+
+int main (){
+
+	int a = 5;
+	int *dev_a;
+
+	cudaMalloc ((void**) &dev_a, sizeof(int));
+
+	cudaMemcpy(dev_a, &a, sizeof(int),cudaMemcpyHostToDevice);
+
+	unsigned int b = 5;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	unsigned long long int c = 5;
+	unsigned long long int *dev_c;
+
+	cudaMalloc ((void**) &dev_c, sizeof(unsigned long long int));
+
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/not-supported/unknown/098_cast_fail/main.cu

@@ -1,0 +1,58 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=32 --gridDim=64 --no-inline
+//error: possible write-write race on
+
+#include <stdio.h>
+#include "cuda.h"
+#include <assert.h>
+#include <cuda_runtime_api.h>
+
+#define M 2//32
+#define N 4//64
+
+__global__ void foo(int* p) {
+  __shared__ unsigned char x[N];
+
+  for (unsigned int i=0; i<(N/4); i++) {
+    ((unsigned int*)x)[i] = 1;//0;
+  }
+/*
+  for (int i = 0; i < N/4; i++) {
+	  p[i] = x[i];
+  }
+*/
+}
+
+int main(){
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 2;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	printf("old a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	foo<<<M,N>>>(dev_a);
+	//ESBMC_verify_kernel(foo,M,N,dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%d	", a[i]);
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/not-supported/verify-kernel/102_basic_argument/main.cu

@@ -1,0 +1,80 @@
+//pass: checka se o parâmetro é passado com sucesso
+//--blockDim=1024 --gridDim=1 --no-inline
+#include <stdio.h>
+#include <stdlib.h>
+
+#define N 8
+
+__device__ float multiplyByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 0.5f;
+}
+
+typedef float(*funcType)(float*, unsigned int);
+
+__global__ void foo(float *v, funcType* f, unsigned int size)
+{
+
+	//*** __requires(f == multiplyByTwo | f == divideByTwo); ****/
+	/************************************************************/
+	assert(*f == divideByTwo || *f == multiplybyTwo);
+	/************************************************************/
+
+	unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+
+    if (tid < size) {
+        v[tid] = (*f)(v, tid);
+    }
+}
+
+int main (){
+
+	float* w;
+	float* dev_w;
+
+	int size = N*sizeof(float);
+
+	w =(float*) malloc(size);
+
+	for (int i = 0; i < N; ++i){
+		w[i] = i;
+	}
+
+	cudaMalloc((void**)&dev_w, size);
+
+	cudaMemcpy(dev_w,w, size,cudaMemcpyHostToDevice);
+
+	funcType* g;
+	funcType* dev_g;
+	g =(funcType*) malloc(sizeof(funcType));
+
+	//*g = multiplyByTwo;
+	*g = divideByTwo;
+
+	cudaMalloc((void**)&dev_g, sizeof(funcType));
+
+	cudaMemcpy(dev_g, g, sizeof(funcType),cudaMemcpyHostToDevice);
+
+	foo <<<1,N>>>(dev_w, dev_g, N );
+
+	cudaMemcpy(w,dev_w,size,cudaMemcpyDeviceToHost);
+
+	cudaMemcpy(g,dev_g,sizeof(funcType),cudaMemcpyDeviceToHost);
+
+	printf("\nw:");
+	for (int i = 0; i < N; ++i){
+		printf(" %f	",	w[i]);
+	}
+
+	//printf ("\n (float) functype: %f", divideByTwo);
+	free(w);
+	cudaFree(&dev_w);
+	cudaFree(&dev_g);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/not-supported/verify-kernel/108_curand_mtgp32/main.cu

@@ -1,0 +1,66 @@
+//pass
+//--blockDim=256 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <curand_kernel.h>
+#include <curand_mtgp32.h>
+#include <stdio.h>
+//#include <curand.h>
+
+#define N 2 //256
+
+__global__ void curand_test(curandStateMtgp32_t *state, float *A) {
+
+	A[threadIdx.x] = curand(&state[threadIdx.x]);
+}
+
+int main() {
+	curandStateMtgp32_t tipo; // Mtgp32_t
+	float *a;
+	float *dev_a;
+	tipo *dev_state;
+	mtgp32_kernel_params *devKernelParams;
+
+	int size = N*sizeof(float);
+
+	a = (float*)malloc(size);
+	cudaMalloc ((void**) &dev_a, size);
+
+	printf("old a:  ");
+	for (int i = 0; i < N; i++)
+	printf("%f	", a[i]);
+
+	cudaMalloc ( (void**) &dev_state, N*sizeof( tipo ) );
+
+	cudaMalloc((void**)&devKernelParams,sizeof(mtgp32_kernel_params));
+
+	curandMakeMTGP32Constants(mtgp32dc_params_fast_11213, devKernelParams); /* Set up constant parameters for the mtgp32 generator */
+		/* mtgp32dc_params_fast_11213 is a constant of the type mtgp32_params_fast, it is a system constant */
+		/* devKernelParams is the destination*/
+
+	curandMakeMTGP32KernelState(dev_state, mtgp32dc_params_fast_11213, devKernelParams,N, 1234); /* Set up initial states for the mtgp32 generator */
+		/*
+		 * \param s - pointer to an array of states in device memory
+		 * \param params - Pointer to an array of type mtgp32_params_fast_t in host memory
+		 * \param k - pointer to a structure of type mtgp32_kernel_params_t in device memory
+		 * \param n - number of parameter sets/states to initialize
+		 * \param seed - seed value
+		 *
+		 * */
+
+//	curand_test<<<1,N>>>(dev_state, dev_a);
+	ESBMC_verify_kernel(curand_test,1,N,dev_state,dev_a);	
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%f	", a[i]);
+
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/not-supported/verify-kernel/113_curand_race_fail/main.cu

@@ -1,0 +1,48 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=2 --gridDim=1 --no-inline
+//Write by thread .+kernel\.cu:8:21:
+
+#include <cuda.h>
+#include <curand.h>
+#include <curand_kernel.h>
+#include <stdio.h>
+#include <assert.h>
+
+#define N 4
+
+__global__ void curand_test(curandState *state, float *A) { // test: replace curandState for curandStateXORWOW_t
+   A[threadIdx.x] = curand_uniform(state);
+}
+
+int main(){
+	float *a;
+	float *dev_a;
+	curandState *dev_state; 
+	
+	int size = N*sizeof(float);
+
+	a = (float*)malloc(size);
+	cudaMalloc ((void**) &dev_a, size);
+
+	printf("old a:  ");
+	for (int i = 0; i < N; i++)
+		printf("%f	", a[i]);
+
+	cudaMalloc ( (void**) &dev_state, N*sizeof( curandState ) );
+
+	curand_test<<<1,N>>>(dev_state, dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nnew a:  ");
+	for (int i = 0; i < N; i++) {
+		printf("%f	", a[i]);
+		assert(a[i] == a[i+1]);
+	}
+
+	free(a);
+	cudaFree(&dev_a);
+	cudaFree(&dev_state);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/assertion/005_soundness_issue_pair/main.cu

@@ -1,0 +1,114 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=128 --gridDim=16 --no-inline
+//assert\(false\)
+
+#include <stdio.h>
+#include <assert.h>
+#include "cuda.h"
+#include "cuda_runtime_api.h"
+
+typedef void(*funcType)(float*);
+
+__device__ void a(float *v)
+{
+	printf ("funcA with p%f = %f", *v, *v);
+}
+__device__ void b(float *v)
+{
+	printf ("funcB with p%f = %f", *v, *v);
+}
+
+__device__ void c(float *v)
+{
+	printf ("funcC with p%f = %f", *v, *v);
+}
+
+__device__ void d(float *v)
+{
+	printf ("funcD with p%f = %f", *v, *v);
+}
+
+__device__ void e(float *v)
+{
+	printf ("funcE with p%f = %f", *v, *v);
+}
+
+__global__ void should_fail(float * __restrict p1, float * __restrict p2, float * __restrict p3, float * __restrict p4, float * __restrict p5, int x, int y)
+{
+	funcType fp = a;
+
+    switch(x) {
+    case 1:
+        fp = &a;
+        break;
+    case 2:
+        fp = &b;
+        break;
+    case 3:
+        fp = &c;
+        break;
+    case 4:
+        fp = &d;
+        break;
+    default:
+        fp = &e;
+        break;
+    }
+
+    switch(y) {
+    case 1:
+        fp(p1);
+        break;
+    case 2:
+        fp(p2);
+        break;
+    case 3:
+        fp(p3);
+        break;
+    case 4:
+        fp(p4);
+        break;
+    default:
+        fp(p5);
+        break;
+    }
+
+   assert(1);
+}
+
+int main (){
+
+	float p1, p2, p3, p4, p5;
+	float *dev_p1, *dev_p2, *dev_p3, *dev_p4, *dev_p5;
+
+	p1 = 1; p2 = 2; p3 = 3; p4 = 4; p5 = 5;
+
+	cudaMalloc((void**)&dev_p1, sizeof(float));
+	cudaMalloc((void**)&dev_p2, sizeof(float));
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/data-racer/014_miscfail3/main.cu

@@ -1,0 +1,59 @@
+//xfail:BOOGIE_ERROR
+//main.cu: error: possible read-write race
+//however, this didn't happen in the tests
+//altough in CUDA providing the inline keyword should still keep a copy of the function around,
+//this kind of access is considered a error by ESBMC
+//ps: the values from A[N-1-offset) to A[N-1] always will receive unpredictable values,
+//because they acess values because they access memory positions that were not initiated
+
+#include <stdio.h>
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#define tid threadIdx.x
+#define N 2//1024
+
+__device__ inline void inlined(int *A, int offset)
+{
+   int temp = A[tid + offset];
+   A[tid] += temp;
+}
+
+__global__ void inline_test(int *A, int offset) {
+  inlined(A, offset);
+}
+
+int main(){
+
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(N*size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = i;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	printf("a:  ");
+	//for (int i = 0; i < N; i++)
+	//	printf("%d	", a[i]);
+
+	inline_test<<<1,N>>>(dev_a, 2);	//you can change this offset for tests
+	//ESBMC_verify_kernel_intt(inline_test, 1, N, dev_a, 2);
+	
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	printf("\nFunction Results:\n   ");
+
+	//for (int i = 0; i < N; i++)
+	//	printf("%d	", a[i]);
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/data-racer/017_equality_abstraction_issue/main.cu

@@ -1,0 +1,43 @@
+//xfail:BOOGIE_ERROR
+//--warp-sync=32 --blockDim=32 --gridDim=1 --equality-abstraction --no-inline
+//kernel.cu:10
+
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <stdio.h>
+#include <assert.h>
+#define N 2//32
+
+__global__ void foo(int * A) {
+    A[0] = 1;
+    A[1] = 1;
+    A[threadIdx.x] = 0;
+//__assert(A[0] == 1 | A[1] == 1 | A[2] == 1);
+}
+
+int main(){
+
+	int *b;
+	int *dev_b;
+
+	b = (int*)malloc(N*sizeof(int));
+
+	for (int i = 0; i < N; ++i){
+		b[i] = 2;
+	}
+
+	cudaMalloc((void**)&dev_b, N*sizeof(int));
+	cudaMemcpy(dev_b, b, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_b);
+	//ESBMC_verify_kernel(foo, 1, N, dev_b);
+
+	cudaMemcpy(b, dev_b, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; ++i){
+		assert(b[i] == 0 || b[i] == 1);
+	}
+
+	free(b);
+	cudaFree(dev_b);
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/data-racer/041_test_copy_between_memory_spaces/main.cu

@@ -1,0 +1,49 @@
+//pass
+//--blockDim=64 --gridDim=1 --equality-abstraction --no-inline
+
+#include "cuda.h"
+#include <stdio.h>
+#include <assert.h>
+#define N 2
+
+__global__ void foo(int* p) {
+
+	__shared__  int A[10];
+
+	A[0] = 1;
+
+	p[0] = A[0];
+
+}
+
+int main(){
+
+	int *b;
+	int *dev_b;
+
+	b = (int*)malloc(N*sizeof(int));
+
+	for (int i = 0; i < N; ++i){
+		b[i] = i+1;
+		printf(" %d; ", b[i]);
+	}
+	printf("\n");
+
+	cudaMalloc((void**)&dev_b, N*sizeof(int));
+
+	cudaMemcpy(dev_b, b, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_b);
+	//ESBMC_verify_kernel(foo,1,N,dev_b);
+
+	cudaMemcpy(b, dev_b, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; ++i){
+		printf(" %d; ", b[i]);
+		assert(b[0]==1);
+	}
+
+	free(b);
+	cudaFree(dev_b);
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/031_pass/main.cu

@@ -1,0 +1,41 @@
+/******************************************* ternarytest2.cu ***************************************/
+/*mostra 0 no índice 0, "c" no índice 1 e nos índices pares, mostra valor lixo nos demais índices */
+
+#include <stdio.h>
+#include "cuda.h"
+#include <assert.h>
+#define N 2 //64
+
+__global__ void foo(float* A, float c) {
+
+  A[threadIdx.x ? 2*threadIdx.x : 1] = c ;
+
+}
+
+int main()
+{
+	float* dev_b;
+	float* b;
+	float c = 2.0f;
+
+	b = (float*)malloc(2*N*sizeof(float)); /* acessível apenas pela CPU função main e funções __host__ */
+
+	cudaMalloc((void**)&dev_b, 2*N*sizeof(float)); /* acessível apenas pela GPU funções __global__ */
+
+	foo<<<1, N>>>(dev_b, c);
+	//ESBMC_verify_kernel_f(foo, 1, N, dev_b, c);
+
+	cudaMemcpy(b, dev_b, 2*N*sizeof(float), cudaMemcpyDeviceToHost);
+
+//	printf("\n");
+
+	for (int i = 0; i < 2*N; ++i){
+//	   printf("%f : ", b[i]);
+	   if((i>0)&&(i%2==0))
+		   assert(b[i] == c);
+	}
+
+	free(b);
+	cudaFree(dev_b);
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/032_return_value_fail/main.cu

@@ -1,0 +1,94 @@
+//fail:assertion
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+#define N 2//1024
+
+typedef float(*funcType)(float*, unsigned int);
+
+__device__ float multiplyByTwo(float *v, unsigned int tid)
+{
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid)
+{
+    return v[tid] * 0.5f;
+}
+
+__device__ funcType grabFunction(int i) {
+  //__requires(i != 0);
+  //__ensures(__return_val_funptr(funcType) == divideByTwo);
+  if (i == 0)
+    return multiplyByTwo;
+  else
+    return divideByTwo;
+}
+
+__global__ void foo(float *v, unsigned int size, int i)
+{
+    //__requires(i != 0);
+    unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+    funcType f = grabFunction(i);
+
+    if (tid < size)
+    {
+    	float x = (*f)(v, tid);
+		x += multiplyByTwo(v, tid);
+		v[threadIdx.x] = x;
+    }
+}
+
+int nondet_int(){
+    int a;
+    return a;
+}
+
+int main(){
+
+	int c = 0; /*define se multiplicará ou dividirá por 2, deve ser 0 ou outro valor, para escolher a função*/
+	float* v;
+	float* a;
+	float* dev_v;
+	funcType fun;
+
+	//fun = (funcType)malloc(sizeof(funcType));
+
+	printf("Digite 0 para multiplicar um vetor por 4 ou\nDigite outro valor para multiplicar um vetor por 2.5: \n");
+    c = nondet_int(); // scanf("%u", &c);
+
+	v = (float*)malloc(N*sizeof(float));
+	a = (float*)malloc(N*sizeof(float));
+
+	for (int i = 0; i < N; ++i){
+		v[i] = rand() %10+1;
+		printf(" %.1f; ", v[i]);
+	}
+
+	printf("\n");
+
+	cudaMalloc((void**)&dev_v, N*sizeof(float));
+
+	cudaMemcpy(dev_v, v, N*sizeof(float), cudaMemcpyHostToDevice);
+
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/033_scanlargelike/main.cu

@@ -1,0 +1,43 @@
+//pass
+//--blockDim=32 --gridDim=64 --no-inline
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+
+#define N 2//32
+
+__device__ void f(float *odata, int* ai) {
+    int thid = threadIdx.x;
+    *ai = thid;
+    odata[*ai] = 2*threadIdx.x;
+}
+
+__global__ void k(float *g_odata) {
+    int ai;
+    f(g_odata,&ai);
+}
+
+
+int main(){
+	float *d;
+	float *dev_d;
+
+	d = (float*)malloc(N*sizeof(float));
+	cudaMalloc ((void**) &dev_d, N*sizeof(float));
+
+	cudaMemcpy(dev_d, d, N*sizeof(float),cudaMemcpyHostToDevice);
+
+	k <<<1,N>>>(dev_d);
+	//ESBMC_verify_kernel(k,1,N,dev_d);
+
+	cudaMemcpy(d,dev_d,N*sizeof(float),cudaMemcpyDeviceToHost);
+
+	printf("D: ");
+	for (int i = 0; i < N; ++i) {
+		printf(" %f		", d[i]);
+		assert(d[i] == 2*i);
+	}
+	cudaFree(dev_d);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/035_test2/main.cu

@@ -1,0 +1,47 @@
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+#define N 2//64
+
+__global__ void foo(int* p) {
+
+  int* q;
+
+  q = p;
+
+  q[threadIdx.x] = 0;
+
+}
+
+int main() {
+	int *c;
+	int *dev_c;
+
+	c = (int*)malloc(N*sizeof(int)); /* acessível apenas pela CPU função main e funções __host__ */
+
+	for (int i = 0; i < N; ++i)
+		c[i] = rand() %10+1;
+
+	for (int i = 0; i < N; ++i)
+		printf(" %d    :", c[i]);
+
+	cudaMalloc((void**)&dev_c, N*sizeof(int)); /* acessível apenas pela GPU funções __global__ */
+
+	cudaMemcpy(dev_c, c, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1, N>>>(dev_c);
+	//ESBMC_verify_kernel(foo,1,N,dev_c);
+
+	cudaMemcpy(c, dev_c, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	printf ("\n");
+
+	for (int i = 0; i < N; ++i){
+		printf(" %d    :", c[i]);
+		assert(c[i] == 0);
+	}
+	free(c);
+	cudaFree(dev_c);
+
+	   return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/036_test4/main.cu

@@ -1,0 +1,49 @@
+//pass: checka a função device (comparar com o cuda69_test2)
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+#define N 2//64
+
+__device__ void bar(int* p) {
+  p[threadIdx.x] = 0;
+  //printf(" %d; ", p[threadIdx.x]);
+}
+
+__global__ void foo(int* p) {
+
+  bar(p);
+
+}
+
+int main() {
+	int *c;
+	int *dev_c;
+	c = (int*)malloc(N*sizeof(int));
+
+	for (int i = 0; i < N; ++i)
+		c[i] = rand() %10+1;
+
+	for (int i = 0; i < N; ++i)
+		printf(" %d; ", c[i]);
+
+	cudaMalloc((void**)&dev_c, N*sizeof(int));
+
+	cudaMemcpy(dev_c, c, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1, N>>>(dev_c);
+	//ESBMC_verify_kernel(foo,1,N,dev_c);
+
+	cudaMemcpy(c, dev_c, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	printf ("\n");
+
+	for (int i = 0; i < N; ++i){
+		printf(" %d; ", c[i]);
+		assert(c[i]==0);
+	}
+
+	free(c);
+	cudaFree(dev_c);
+
+	   return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/039_test7/main.cu

@@ -1,0 +1,55 @@
+/*** substitui os valores aleatórios de determinado vetor de tamanho N por valores ordenados de 0 a N ***/
+#include <stdio.h>
+#include <assert.h>
+#include <cuda.h>
+#define N 2//64
+
+__global__ void foo(int* glob) {
+
+  int a;
+
+  int* p;
+
+  a = 0;
+
+  p = &a;
+
+  *p = threadIdx.x;
+
+  glob[*p] = threadIdx.x;
+}
+
+int main(){
+
+	int* v;
+	int* dev_v;
+
+	/* seta o tamanho de v e inicia com com valores randômicos */
+	v = (int*)malloc(N*sizeof(int)); /* acessível apenas pela CPU função main e funções __host__ */
+
+	for (int i = 0; i < N; ++i)
+		v[i] = rand() %20+1;;
+
+	for (int i = 0; i < N; ++i)
+		printf(" %d    :", v[i]);
+
+	cudaMalloc((void**)&dev_v, N*sizeof(int)); /* acessível apenas pela GPU funções __global__ */
+
+	cudaMemcpy(dev_v, v, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1, N>>>(dev_v);
+	//ESBMC_verify_kernel(foo,1,N,dev_v);
+
+	cudaMemcpy(v, dev_v, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	printf("\n\n\n");
+
+	for (int i = 0; i < N; ++i){
+		printf(" %d    :", v[i]);
+		assert(v[i]==i);
+	}
+	free(v);
+	cudaFree(dev_v);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/047_pass/main.cu

@@ -1,0 +1,40 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include "cuda_runtime_api.h"
+#include <assert.h>
+
+#define N 2//64
+
+__global__ void foo(int *c) {
+  int b, a;
+  a = 2;
+  b = 3;
+  c[threadIdx.x]= a+b;
+  __syncthreads ();
+}
+
+int main(){
+	int *a;
+	int *dev_a;
+
+	a = (int*)malloc(N*sizeof(int));
+
+	cudaMalloc((void**)&dev_a, N*sizeof(int));
+
+	cudaMemcpy(dev_a, a, N*sizeof(int), cudaMemcpyHostToDevice);
+
+	foo<<<1, N>>>(dev_a);
+	//ESBMC_verify_kernel(foo,1,N,dev_a);
+
+	cudaMemcpy(a, dev_a, N*sizeof(int), cudaMemcpyDeviceToHost);
+
+	for (int t=0;t<N;t++){
+		printf ("%d ", a[t]);
+		assert(a[t]==5);
+	}
+
+	cudaFree(dev_a);
+	free(a);
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/048_pass/main.cu

@@ -1,0 +1,25 @@
+//pass
+//--blockDim=64 --gridDim=64 --no-inline
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+
+#define N 2//64
+
+
+__global__ void foo() {
+
+  float x = (float)2;
+
+}
+
+int main(){
+	
+	foo<<<1, N>>>();
+	//ESBMC_verify_kernel(foo,1,N);
+
+	cudaThreadSynchronize();	
+
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/049_pass/main.cu

@@ -1,0 +1,41 @@
+#include <stdio.h>
+#include <assert.h>
+#include "cuda.h"
+
+#define N 2
+
+__global__ void foo(int* p) {
+    p[threadIdx.x] = 2;
+    __syncthreads();
+}
+
+int main(){
+
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(N*size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(foo,1,N,dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++){
+		assert(a[i]==2);
+	}
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/050_fail/main.cu

@@ -1,0 +1,45 @@
+//fail
+//--blockDim=2048 --gridDim=2 --no-inline
+#include <stdio.h>
+#include <assert.h>
+#include <cuda.h>
+
+#define N 2//2048
+
+__constant__ int A[4096];
+__constant__ int B[3] = {0,1,2};
+
+__global__ void kernel(int* x) {
+  x[threadIdx.x] = A[threadIdx.x] + B[0]; //permanece constante por ser muito grande. N < 1024 não permanece
+}
+int main () {
+
+	int *a;
+	int *c;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);	
+
+	a = (int*)malloc(size);
+	c = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = rand() %10+1;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);	
+
+	kernel<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(kernel,1,N,dev_a);
+
+	cudaMemcpy(c,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++){
+		assert(c[i]!=0);
+	}
+	free(a);
+	free(c);
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/058_noraceduetoreturn_fail/main.cu

@@ -1,0 +1,53 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__device__ int bar(float* A) {
+
+  if(threadIdx.x != 0) {
+	return 0;
+  }
+
+  return 1;
+
+}
+
+__global__ void foo(float* A) {
+
+  int y = bar(A);
+  A[threadIdx.x]=y;
+
+}
+
+int main(void){
+	int i;
+	float *A;
+	float *dev_A;
+	float size= N*sizeof(float);
+	A=(float*)malloc(size);
+
+	for(i=0;i<N;i++)
+		A[i]=2;
+			
+	cudaMalloc((void**)&dev_A,size);
+
+	cudaMemcpy(dev_A, A,size, cudaMemcpyHostToDevice);
+
+	foo<<<1,N>>>(dev_A);
+	//ESBMC_verify_kernel_f(foo,1,N,dev_A);
+
+	cudaMemcpy(A,dev_A,size,cudaMemcpyDeviceToHost);
+
+	for(i=0;i<N;i++){
+//		printf("%.2f ",A[i]);
+		if (i!=0)
+			assert(A[i]!=0);
+		
+	}
+	cudaFree(dev_A);
+	free(A);
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/062_struct_fail/main.cu

@@ -1,0 +1,49 @@
+//fail
+//--blockDim=64 --gridDim=64 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <assert.h>
+
+#define DIM  2//64
+#define N 2 //DIM*DIM
+
+typedef struct {
+  float x,y,z,w;
+} myfloat4;
+
+__global__ void k(float * i0) {
+  myfloat4 f4;
+  f4.x = 2;
+  i0[threadIdx.x + blockDim.x*blockIdx.x] = f4.x;
+}
+
+int main(){
+	float *a;
+	float *dev_a;
+	int size = N*sizeof(float);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (float*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 5;
+
+	cudaMemcpy(dev_a,a,size, cudaMemcpyHostToDevice);
+
+		k<<<DIM,DIM>>>(dev_a);
+		//ESBMC_verify_kernel_f(k,1,N,dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++) {
+		assert(a[i] != 2);
+	}
+
+	free(a);
+
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/063_foo/main.cu

@@ -1,0 +1,43 @@
+#include <stdio.h>
+#include "cuda.h"
+#include <assert.h>
+#define N 2 //16
+
+__device__ int bar(int x) {
+
+	return x + 1;
+}
+
+__global__ void foo(int *A) {
+
+	A[threadIdx.x] = bar(threadIdx.x);
+}
+
+
+int main () {
+
+	int *a;
+	int *dev_a;
+	int size = N*sizeof(int);
+
+	cudaMalloc((void**)&dev_a, size);
+
+	a = (int*)malloc(size);
+
+	for (int i = 0; i < N; i++)
+		a[i] = 1;
+
+	foo<<<1,N>>>(dev_a);
+	//ESBMC_verify_kernel(foo, 1, N, dev_a);
+
+	cudaMemcpy(a,dev_a,size,cudaMemcpyDeviceToHost);
+
+	for (int i = 0; i < N; i++){
+		assert(a[i]== (i+1));
+	}
+
+	free(a);
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/064_atomicAdd/main.cu

@@ -1,0 +1,69 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+
+#define N 2 //1024
+
+__global__ void definitions (int* A, unsigned int* B, unsigned long long int* C, float* D)
+{
+	atomicAdd(A,10);
+
+	atomicAdd(B,10);
+
+	atomicAdd(C,10);
+
+	atomicAdd(D,10);
+
+}
+
+int main (){
+
+	int a = 5;
+	int *dev_a;
+
+	cudaMalloc ((void**) &dev_a, sizeof(int));
+
+	cudaMemcpy(dev_a, &a, sizeof(int),cudaMemcpyHostToDevice);
+
+	unsigned int b = -5;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	unsigned long long int c = 0;
+	unsigned long long int *dev_c;
+
+	cudaMalloc ((void**) &dev_c, sizeof(unsigned long long int));
+
+	cudaMemcpy(dev_c, &c, sizeof(unsigned long long int),cudaMemcpyHostToDevice);
+
+	float d = 10;
+	float *dev_d;
+
+	cudaMalloc ((void**) &dev_d, sizeof(float));
+
+	cudaMemcpy(dev_d, &d, sizeof(float),cudaMemcpyHostToDevice);
+
+	definitions <<<1,N>>>(dev_a,dev_b,dev_c,dev_d);
+	//ESBMC_verify_kernel_four(definitions,1,N,dev_a,dev_b,dev_c,dev_d);
+
+	cudaMemcpy(&a,dev_a,sizeof(int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&b,dev_b,sizeof(unsigned int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&c,dev_c,sizeof(unsigned long long int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&d,dev_d,sizeof(float),cudaMemcpyDeviceToHost);
+
+	assert(a==25);
+	assert(b==15);
+	assert(c==20);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+	cudaFree(dev_c);
+	cudaFree(dev_d);
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/067_atomicDec_fail/main.cu

@@ -1,0 +1,37 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <sm_atomic_functions.h>
+#include <cuda_runtime_api.h>
+
+#define N 2 //1024
+
+__global__ void definitions (unsigned int* B)
+{
+  atomicDec(B,7);//0111 -> 1000 -> 0000 -> 0001 -> 0010 -> 0011 -> 0100 -> 0101 -> 0110 ...
+  	  /*the second argument on atomicDec() is a limit for decs. When this limit is reached, B receives <LIM>*/
+}
+
+int main (){
+
+	unsigned int b = 5;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	definitions <<<1,N>>>(dev_b);
+	//ESBMC_verify_kernel_ui(definitions,1,N,dev_b);
+
+	cudaMemcpy(&b,dev_b,sizeof(unsigned int),cudaMemcpyDeviceToHost);
+
+	printf("B: %u\n", b);
+	assert(b!=3);
+
+	cudaFree(dev_b);
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/073_atomicSub/main.cu

@@ -1,0 +1,50 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+
+#define N 2 //1024
+
+__global__ void definitions (int* A, unsigned int* B)
+{
+	atomicSub(A,10);
+
+	atomicSub(B,5);
+
+}
+
+int main (){
+
+	int a = 5;
+	int *dev_a;
+
+	cudaMalloc ((void**) &dev_a, sizeof(int));
+
+	cudaMemcpy(dev_a, &a, sizeof(int),cudaMemcpyHostToDevice);
+
+	unsigned int b = 10;
+	unsigned int *dev_b;
+
+	cudaMalloc ((void**) &dev_b, sizeof(unsigned int));
+
+	cudaMemcpy(dev_b, &b, sizeof(unsigned int),cudaMemcpyHostToDevice);
+
+	definitions <<<1,N>>>(dev_a,dev_b);
+	//ESBMC_verify_kernel(definitions,1,N,dev_a,dev_b);
+
+	cudaMemcpy(&a,dev_a,sizeof(int),cudaMemcpyDeviceToHost);
+	cudaMemcpy(&b,dev_b,sizeof(unsigned int),cudaMemcpyDeviceToHost);
+
+	printf("A: %d\n", a);
+	printf("B: %u\n", b);
+
+	assert(a==-15);
+	assert(b==0);
+
+	cudaFree(dev_a);
+	cudaFree(dev_b);
+
+	return 0;
+
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/076_fail/main.cu

@@ -1,0 +1,37 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__global__ void foo(float* A) {
+
+  if(blockIdx.x == 0)
+	  A[threadIdx.x] = 50.f;
+}
+
+int main() {
+	float *a = 0;
+	float *dev_a;
+
+	a = (float*)malloc(N*sizeof(float));
+
+	cudaMalloc((void**)&dev_a, N*sizeof(float));
+
+	cudaMemcpy(dev_a, a, N*sizeof(float), cudaMemcpyHostToDevice);
+
+		foo<<<1, N>>>(dev_a);
+		//ESBMC_verify_kernel_f(foo, 1, N, dev_a);
+
+	cudaMemcpy(a, dev_a, N*sizeof(float), cudaMemcpyDeviceToHost);
+
+	printf ("%.1f", *a);
+
+		assert(*a != 50.0);
+
+	free(a);
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/076_pass/main.cu

@@ -1,0 +1,37 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__global__ void foo(float* A) {
+
+  if(blockIdx.x == 0)
+	  A[threadIdx.x] = 50.f;
+}
+
+int main() {
+	float *a = 0;
+	float *dev_a;
+
+	a = (float*)malloc(N*sizeof(float));
+
+	cudaMalloc((void**)&dev_a, N*sizeof(float));
+
+	cudaMemcpy(dev_a, a, N*sizeof(float), cudaMemcpyHostToDevice);
+
+		foo<<<1, N>>>(dev_a);
+		//ESBMC_verify_kernel_f(foo, 1, N, dev_a);
+
+	cudaMemcpy(a, dev_a, N*sizeof(float), cudaMemcpyDeviceToHost);
+
+	printf ("%.1f", *a);
+
+		assert(*a == 50.0);
+
+	free(a);
+	cudaFree(dev_a);
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/082_threadfence/main.cu

@@ -1,0 +1,18 @@
+//pass
+//--blockDim=32 --gridDim=2
+
+#include <cuda.h>
+
+__global__ void foo() {
+    __threadfence();
+}
+
+int main(){
+
+	foo<<<1,2>>>();
+	//ESBMC_verify_kernel(foo,1,2);
+
+	cudaThreadSynchronize();
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GKLEE/supported/pass/086_unknown_function/main.cu

@@ -1,0 +1,41 @@
+//pass
+//blockDim=1024 --gridDim=1 --no-inline
+
+#include <stdio.h>
+#include <cuda.h>
+#include <cuda_runtime_api.h>
+#include <math_functions.h>
+
+typedef double(*funcType)(double);
+
+__device__ double bar(double x) {
+  return sin(x);
+}
+
+__global__ void foo(double x, int i)
+{
+  funcType f;
+
+  if (i == 0)
+    f = bar;
+  else
+    f = cos;
+
+  double z = f(x);
+	assert(z != NULL);
+
+  printf("z: %f ", z);
+}
+
+int main(){
+
+	int select_function = 1; // 1= sen; 0=cos
+	double angle = 1.57; //0;
+
+	foo <<<1,2>>>(angle, select_function);
+	//ESBMC_verify_kernel_c (foo,1,2,angle, select_function);
+
+	cudaThreadSynchronize();
+
+	return 0;
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/cuda-functions/081_localarrayaccess/main.cu

@@ -1,0 +1,22 @@
+//pass 
+//--blockDim=10 --gridDim=64 --no-inline
+
+#include <stdio.h>
+#include <cuda.h>
+
+#include <assert.h>
+
+#define N 2
+
+__global__ void foo() {
+
+  __shared__ int A[10];
+
+  A[threadIdx.x] = 2;
+
+  __syncthreads (); //evita corrida de dados
+
+  int x = A[threadIdx.x + 1];
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/unknown/098_cast_fail/main.cu

@@ -1,0 +1,25 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=32 --gridDim=64 --no-inline
+//error: possible write-write race on
+
+#include <stdio.h>
+#include "cuda.h"
+#include <assert.h>
+
+
+#define M 2//32
+#define N 4//64
+
+__global__ void foo(int* p) {
+  __shared__ unsigned char x[N];
+
+  for (unsigned int i=0; i<(N/4); i++) {
+    ((unsigned int*)x)[i] = 1;//0;
+  }
+/*
+  for (int i = 0; i < N/4; i++) {
+	  p[i] = x[i];
+  }
+*/
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/101_funcptr_to_ptr/main.cu

@@ -1,0 +1,47 @@
+//pass
+//--blockDim=1024 --gridDim=1 --boogie-file=${KERNEL_DIR}/axioms.bpl --no-inline
+#include <stdio.h>
+#include <stdlib.h>
+#include <assert.h>
+#include <cuda.h>
+
+#define N 8
+
+typedef float(*funcType)(float*, unsigned int);
+
+__device__ float multiplyByTwo(float *v, unsigned int tid)
+{
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid)
+{
+    return v[tid] * 0.5f;
+}
+
+// Static pointers to device functions
+
+	__device__ funcType p_mul_func = multiplyByTwo;
+
+	__device__ funcType p_div_func = divideByTwo;
+
+
+__global__ void foo(float *v, funcType f, unsigned int size, int i)
+{
+    __assert(f == divideByTwo);
+	__assert(i != 0);
+
+	uint tid = blockIdx.x * blockDim.x + threadIdx.x;
+
+    void *x = (void*)f; /*ptr_to_ptr*/
+
+    if (i == 0)
+      x = NULL;
+
+    funcType g = (funcType)x;
+
+    if (tid < size)
+    {
+        v[tid] = (*g)(v, tid);
+    }
+}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/102_basic_argument/main.cu

@@ -1,0 +1,34 @@
+//pass: checka se o parâmetro é passado com sucesso
+//--blockDim=1024 --gridDim=1 --no-inline
+#include <stdio.h>
+#include <stdlib.h>
+
+#define N 8
+
+__device__ float multiplyByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 0.5f;
+}
+
+typedef float(*funcType)(float*, unsigned int);
+
+__global__ void foo(float *v, funcType* f, unsigned int size)
+{
+
+	__requires(f == multiplyByTwo | f == divideByTwo);
+	/************************************************************/
+	__assert(*f == divideByTwo || *f == multiplybyTwo);
+	/************************************************************/
+
+	unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+
+    if (tid < size) {
+        v[tid] = (*f)(v, tid);
+    }
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/102_basic_argument_fail/main.cu

@@ -1,0 +1,30 @@
+//fail: o argumento não é passado com sucesso
+//--blockDim=1024 --gridDim=1 --no-inline
+#include <stdio.h>
+#include <stdlib.h>
+
+#define N 8
+
+__device__ float multiplyByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 0.5f;
+}
+
+typedef float(*funcType)(float*, unsigned int);
+
+__global__ void foo(float *v, funcType* f, unsigned int size)
+{
+	unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+
+	__assert(*f == divideByTwo || *f == multiplybyTwo);
+
+    if (tid < size) {
+        v[tid] = (*f)(v, tid);
+    }
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/108_curand_mtgp32/main.cu

@@ -1,0 +1,16 @@
+//pass
+//--blockDim=256 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <curand_kernel.h>
+#include <curand_mtgp32.h>
+#include <stdio.h>
+//#include <curand.h>
+
+#define N 2 //256
+
+__global__ void curand_test(curandStateMtgp32_t *state, float *A) {
+
+	A[threadIdx.x] = curand(&state[threadIdx.x]);
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/109_curand_test/main.cu

@@ -1,0 +1,44 @@
+#include <iostream>
+#include <cuda.h>
+#include <curand.h>
+#include <curand_kernel.h>
+
+#define N 4
+
+int n = 20; //it defines the range of the random number
+using namespace std;
+
+__device__ float generate( curandState* globalState, int ind ) // ind varies from 0 to N
+{
+    //int ind = threadIdx.x;
+    curandState localState = globalState[ind];
+    float RANDOM = curand_uniform( &localState ); // float curand_uniform(curandStateXORWOW_t *state
+    											// Return a uniformly distributed float between \p 0.0f and \p 1.0f
+    globalState[ind] = localState; // localState received a new value based on its own 'seed'
+    return RANDOM;
+}
+
+__global__ void setup_seed ( curandState * state, unsigned long seed )
+{
+    int id = threadIdx.x;
+    curand_init ( seed, id, 0, &state[id] );
+
+    // curand_init generate random numbers that will be stored on state[id]
+    // seed must be a random number
+}
+
+__global__ void kernel(float* N3, curandState* globalState, int n) // n = 2
+{
+    // generate random numbers
+    for(int i=0;i<N;i++)
+    {					// globalState received the 'seeds'
+        int k = generate(globalState, i) * (10*N/4); // float generate (curandState* globalState, int ind) is a __device__ function
+        while(k > n*n-1) // k >3 THIS WHILE DEFINES THE RANGE OF THE RANDOM NUBER, TO n=2 and N=4 deinine the range (0,3]
+//        ESTE WHILE DEFINE O RANGE DO NÚMERO ALEATÓRIO, neste caso o limita a (0,3]
+        {
+            k-=(n*n-1); // k = k -3
+        }
+        N3[i] = k; //  10 -> 1; 9 -> 3; 8 -> 2; 7 -> 1; etc
+    }
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/not-supported/verify-kernel/112_curand_mtgp32_race/main.cu

@@ -1,0 +1,18 @@
+//pass
+//--blockDim=256 --gridDim=1 --no-inline
+
+#include <cuda.h>
+
+#include <curand.h>
+#include <curand_kernel.h>
+//#include <curand_precalc.h>
+//#include <curand_mtgp32_host.h>
+#include <stdio.h>
+
+#define N 256
+
+__global__ void curand_test(curandStateMtgp32_t *state, float *A) {
+
+	A[threadIdx.x] = curand(&state[threadIdx.x]);
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/assertion/009_testinterprocedural3/main.cu

@@ -1,0 +1,41 @@
+//fail
+//--blockDim=64 --gridDim=64 --no-inline
+#include "cuda.h"
+#include <stdio.h>
+
+#define N 1
+
+__device__ void baz (int p []){
+    int a;
+
+    p = &a;
+}
+
+__device__ void bar (int *p){
+
+    int a = 2;
+
+    p = &a;
+}
+
+
+__global__ void foo (int* p, int* q){
+
+	__requires (*q == 1)
+    __shared__ int sharedArr  [100];
+
+    __shared__ int sharedArr2 [50];
+
+    bar(p);
+
+    baz (sharedArr);
+
+    bar(q);
+
+    if (*q){
+        baz(sharedArr2);
+    }
+
+    //*p = 23; *q = 23; // remove this comment to see that the __device__ function does not work
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/data-racer/015_miscfail4/main.cu

@@ -1,0 +1,25 @@
+//xfail:BOOGIE_ERROR
+//main.cu: error: possible read-write race
+//however, this didn't happen in the tests
+// In CUDA providing static and __attribute__((always_inline)) SHOUD NOT
+// keep a copy of inlined function around.
+//ps: the values from A[N-1-offset] to A[N-1] always will receive unpredictable values,
+//because they acess values because they access memory positions that were not initiated
+
+#include <stdio.h>
+#include <cuda.h>
+
+#define tid threadIdx.x
+#define N 2//32
+
+__device__ static __attribute__((always_inline))
+void inlined(int *A, int offset)
+{
+   int temp = A[tid + offset];
+   A[tid] += temp;
+}
+
+__global__ void inline_test(int *A, int offset) {
+  inlined(A, offset);
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/data-racer/093_bronken_shuffle/main.cu

@@ -1,0 +1,18 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=1024 --gridDim=1 --warp-sync=16 --no-inline
+//It should show only the values from B[0] to B[31], but it exceeds.
+
+#include <cuda.h>
+
+#include <stdio.h>
+
+#define N 2//32//1024
+
+__global__ void shuffle (int* A)
+{
+	int tid = threadIdx.x;
+	int warp = tid / 32;
+	int* B = A + (warp*32);
+	A[tid] = B[(tid + 1)%32];
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/data-racer/097_loop_unwind/main.cu

@@ -1,0 +1,24 @@
+//xfail:BOOGIE_ERROR
+//--blockDim=512 --gridDim=64 --loop-unwind=2 --no-inline
+//kernel.cu: error: possible write-write race on B
+
+#include <cuda.h>
+
+#include <stdio.h>
+#include <assert.h>
+
+#define N 2//512
+
+extern "C" {
+
+__global__ void helloCUDA(float *A)
+{
+    __shared__ float B[256];
+
+    for(int i = 0; i < N*2; i ++) {
+        B[i] = A[i];
+    }
+}
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/null-pointer/020_basic_assignment_pair/main.cu

@@ -1,0 +1,40 @@
+#include <stdio.h>
+#include "cuda.h"
+
+#define N 2 //64
+
+__device__ float multiplyByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 2.0f;
+}
+
+__device__ float divideByTwo(float *v, unsigned int tid) {
+
+    return v[tid] * 0.5f;
+}
+
+typedef float(*funcType)(float*, unsigned int);
+
+__global__ void foor(float *v, unsigned int size, unsigned int i) {
+
+    unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+    funcType f;
+__requires (i == 1 || i == 2)
+	/*** ESBMC_assert (i = 1 || i = 2); ***/
+		__assert(i == 1 || i == 2);
+		
+    if (i == 1)
+      f = multiplyByTwo;
+    else if (i == 2)
+      f = divideByTwo;
+    else
+      f = NULL;
+
+    if (tid < size)
+    {
+        float x = (*f)(v, tid);
+        x += multiplyByTwo(v, tid);
+		v[tid] = x;
+    }
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/null-pointer/022_basic_statement_pair/main.cu

@@ -1,0 +1,53 @@
+//pass: checka a declaração de uma variável do tipo "ponteiro pra função" dentro do kernel
+//--blockDim=1024 --gridDim=1 --no-inline
+#include <cuda.h>
+
+#include <stdio.h>
+#include <assert.h>
+
+#define tid (blockIdx.x * blockDim.x + threadIdx.x)
+#define N 2//1024
+
+__device__ void multiplyByTwo(float *v, unsigned int index)
+{
+    /*__requires(__ptr_offset_bytes(v) == 0);
+    __requires(index == tid);
+    __requires(__read_implies(v, __read_offset_bytes(v)/sizeof(float) == index));
+    __requires(__write_implies(v, __write_offset_bytes(v)/sizeof(float) == index));
+    __ensures(__read_implies(v, __read_offset_bytes(v)/sizeof(float) == index));
+    __ensures(__write_implies(v, __write_offset_bytes(v)/sizeof(float) == index));*/
+    
+	v[index] = v[index] * 2.0f;
+}
+
+__device__ void divideByTwo(float *v, unsigned int index)
+{
+    /*__requires(__ptr_offset_bytes(v) == 0);
+    __requires(index == tid);
+    __requires(__read_implies(v, __read_offset_bytes(v)/sizeof(float) == index));
+    __requires(__write_implies(v, __write_offset_bytes(v)/sizeof(float) == index));
+    __ensures(__read_implies(v, __read_offset_bytes(v)/sizeof(float) == index));
+    __ensures(__write_implies(v, __write_offset_bytes(v)/sizeof(float) == index));*/
+    
+    v[index] = v[index] * 0.5f;
+}
+
+typedef void(*funcType)(float*, unsigned int);
+
+__global__ void foo(float *v, unsigned int i)
+{
+    __requires(i == 1 | i == 2);
+    	__assert(i == 1 | i == 2);
+    	
+    funcType f;
+
+    if (i == 1)
+      f = multiplyByTwo;
+    else if (i == 2)
+      f = divideByTwo;
+    else
+      f = NULL;
+
+    (*f)(v, tid);
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/030_fail/main.cu

@@ -1,0 +1,16 @@
+/******************************************* ternarytest.cu **************************************/
+/*mostra 0 no índice 0, "c" no índice 1 e nos índice pares, mostra valor lixo nos demais índices */
+//fail: assert()	
+
+#include <stdio.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__global__ void foo(float* A, float c) {
+
+	A[threadIdx.x == 0 ? 1 : 2*threadIdx.x] = c;
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/035_test2/main.cu

@@ -1,0 +1,15 @@
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+#define N 2//64
+
+__global__ void foo(int* p) {
+
+  int* q;
+
+  q = p;
+
+  q[threadIdx.x] = 0;
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/036_test4/main.cu

@@ -1,0 +1,17 @@
+//pass: checka a função device (comparar com o cuda69_test2)
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+#define N 2//64
+
+__device__ void bar(int* p) {
+  p[threadIdx.x] = 0;
+  //printf(" %d; ", p[threadIdx.x]);
+}
+
+__global__ void foo(int* p) {
+
+  bar(p);
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/038_test6/main.cu

@@ -1,0 +1,20 @@
+#include <stdio.h>
+#include <cuda.h>
+#include <assert.h>
+
+#define N 2//64
+
+__device__ int* bar(int* p) {
+
+	//__ensures(__implies(__enabled(), __return_val_ptr() == p));
+	return p;
+}
+
+__global__ void foo(int* p) {
+
+  //bar(p)[threadIdx.x] = 0;
+  *(bar(p)+threadIdx.x) = 2;
+  //printf(" %d; ", bar(p)[threadIdx.x]);
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/047_fail/main.cu

@@ -1,0 +1,15 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__global__ void foo(int *c) {
+  int b, a;
+  a = 2;
+  b = 3;
+  c[threadIdx.x]= a+b;
+  __syncthreads ();
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/049_fail/main.cu

@@ -1,0 +1,11 @@
+#include <stdio.h>
+#include <assert.h>
+#include "cuda.h"
+
+#define N 2
+
+__global__ void foo(int* p) {
+    p[threadIdx.x] = 2;
+    __syncthreads();
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/050_fail/main.cu

@@ -1,0 +1,15 @@
+//fail
+//--blockDim=2048 --gridDim=2 --no-inline
+#include <stdio.h>
+#include <assert.h>
+#include <cuda.h>
+
+#define N 2//2048
+
+__constant__ int A[4096];
+__constant__ int B[3] = {0,1,2};
+
+__global__ void kernel(int* x) {
+  x[threadIdx.x] = A[threadIdx.x] + B[0]; //permanece constante por ser muito grande. N < 1024 não permanece
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/050_pass/main.cu

@@ -1,0 +1,15 @@
+//pass
+//--blockDim=2048 --gridDim=2 --no-inline
+#include <stdio.h>
+#include <assert.h>
+#include <cuda.h>
+
+#define N 2//2048
+
+__constant__ int A[4096];
+__constant__ int B[3] = {0,1,2};
+
+__global__ void kernel(int* x) {
+  x[threadIdx.x] = A[threadIdx.x] + B[0]; //permanece constante por ser muito grande. N < 1024 não permanece
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/058_noraceduetoreturn_fail/main.cu

@@ -1,0 +1,24 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include "cuda.h"
+#include <assert.h>
+
+#define N 2//64
+
+__device__ int bar(float* A) {
+
+  if(threadIdx.x != 0) {
+	return 0;
+  }
+
+  return 1;
+
+}
+
+__global__ void foo(float* A) {
+
+  int y = bar(A);
+  A[threadIdx.x]=y;
+
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/065_atomicAnd/main.cu

@@ -1,0 +1,17 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+
+#define N 2 //1024
+
+__global__ void definitions (int* A, unsigned int* B, unsigned long long int* C)
+{
+  atomicAnd(A,10);
+
+  atomicAnd(B,1);
+
+  atomicAnd(C,5);
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/067_atomicDec_fail/main.cu

@@ -1,0 +1,16 @@
+//pass
+//--blockDim=1024 --gridDim=1 --no-inline
+
+#include <cuda.h>
+#include <stdio.h>
+#include <sm_atomic_functions.h>
+
+
+#define N 2 //1024
+
+__global__ void definitions (unsigned int* B)
+{
+  atomicDec(B,7);//0111 -> 1000 -> 0000 -> 0001 -> 0010 -> 0011 -> 0100 -> 0101 -> 0110 ...
+  	  /*the second argument on atomicDec() is a limit for decs. When this limit is reached, B receives <LIM>*/
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/075_reduced_strength_blockwise/main.cu

@@ -1,0 +1,51 @@
+//pass
+//--blockDim=256 --gridDim=2 -DWIDTH=2064 --no-inline
+#include <cuda.h>
+
+#include <stdio.h>
+
+#define GRIDDIM 1
+#define BLOCKDIM 2//256
+#define WIDTH 2//2048
+#define N WIDTH
+/*
+ * This kernel demonstrates a blockwise strength-reduction loop.
+ * Each block is given a disjoint partition (of length WIDTH) of A.
+ * Then each thread writes multiple elements in the partition.
+ * It is not necessarily the case that WIDTH%blockDim.x == 0
+ */
+
+__global__ void k(int *A) {
+//  __assert(blockDim.x <= WIDTH);
+//#ifdef BLOCK_DIVIDES_WIDTH
+//  //__assert(__mod_pow2(WIDTH, blockDim.x) == 0);
+//#endif
+
+  for (int i=threadIdx.x; i<WIDTH; i+=blockDim.x) {
+
+//#ifndef BLOCK_DIVIDES_WIDTH
+//    // working set(1) using global invariants
+//    /*A*/__global_invariant(__write_implies(A, (blockIdx.x*WIDTH) <= __write_offset_bytes(A)/sizeof(int))),
+//    /*B*/__global_invariant(__write_implies(A,                       __write_offset_bytes(A)/sizeof(int) < (blockIdx.x+1)*WIDTH)),
+//    /*C*/__invariant(threadIdx.x <= i),
+//    /*D*/__invariant(               i <= WIDTH+blockDim.x),
+//         __invariant(i % blockDim.x == threadIdx.x),
+//         __global_invariant(__write_implies(A, (((__write_offset_bytes(A)/sizeof(int)) % WIDTH) % blockDim.x) == threadIdx.x)),
+//#else
+//    // working set(2) iff WIDTH % blockDim.x == 0
+//    /*A*/__invariant(__write_implies(A, (blockIdx.x*WIDTH) <= __write_offset_bytes(A)/sizeof(int))),
+//    /*B*/__invariant(__write_implies(A,                       __write_offset_bytes(A)/sizeof(int) < (blockIdx.x+1)*WIDTH)),
+//    /*C*/__invariant(threadIdx.x <= i),
+//    /*D*/__invariant(               i <= WIDTH+blockDim.x),
+//         __invariant(__uniform_int((i-threadIdx.x))),
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/GPUVerify/supported/pass/088_fail/main.cu

@@ -1,0 +1,22 @@
+//fail
+//--blockDim=1024 --gridDim=1024 --no-inline
+
+#include <stdio.h>
+#include <assert.h>
+
+#include <cuda.h>
+
+#include <math_functions.h>
+
+#define DIM 2 //1024 in the future
+#define N 2//DIM*DIM
+
+__global__ void mul24_test (int* A, int* B)
+{
+  int idxa          = __mul24(blockIdx.x, blockDim.x) + threadIdx.x;
+  unsigned int idxb = __umul24(blockIdx.x, blockDim.x) + threadIdx.x;
+
+  A[idxa] = idxa;
+  B[idxb] = idxa;
+}
+

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/not-supported/verify-kernel/101_funcptr_to_ptr/main.cu

@@ -1,0 +1,27 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(float *v, funcType f, unsigned int size, int i)
+{
+    assert(f == divideByTwo);
+	assert(i != 0);
+
+	uint tid = blockIdx.x * blockDim.x + threadIdx.x;
+
+    void *x = (void*)f; /*ptr_to_ptr*/
+
+    if (i == 0)
+      x = NULL;
+
+    funcType g = (funcType)x;
+
+    if (tid < size)
+    {
+        v[tid] = (*g)(v, tid);
+    }
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/not-supported/verify-kernel/110_curand_init_race_fail/main.cu

@@ -1,0 +1,18 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(curandState *state, unsigned int *A) {
+   curand_init(0, 0, 0, state);
+
+   __syncthreads();
+
+   A[threadIdx.x] =  curand(&state[threadIdx.x]);
+//   if (threadIdx.x == 0) {
+  //   A[0] = curand(state);
+   //}
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/not-supported/verify-kernel/112_curand_mtgp32_race/main.cu

@@ -1,0 +1,12 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(curandStateMtgp32_t *state, float *A) {
+
+	A[threadIdx.x] = curand(&state[threadIdx.x]);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/array-bounds/104_misc8_fail/main.cu

@@ -1,0 +1,17 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(char *A, char *B, char* c)
+{
+
+  char *choice1 = *c ? A : B;	//It Makes choice1 receives A
+  char *choice2 = *c ? B : A;	//It Makes choice2 receives B
+
+  bar(&choice1, &choice2);
+  bar(&choice1, &choice2);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/data-racer/010_test3/main.cu

@@ -1,0 +1,13 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(int* p) {
+  int* q;
+  q = p + 1;
+  p[threadIdx.x] = q[threadIdx.x];
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/data-racer/016_bad_inter_group/main.cu

@@ -1,0 +1,13 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(int* A) {
+
+   A[ blockIdx.x*blockDim.x + threadIdx.x ] += (A[ (blockIdx.x + 1)*blockDim.x + threadIdx.x ]);
+
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/data-racer/042_test_copy_between_pointers/main.cu

@@ -1,0 +1,25 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(int* p) {
+
+  __shared__ int A[10];
+
+  int* x;
+
+  x = p;
+
+  	assert(*p <2);
+
+  x[0] = 0;
+	
+  x = A;
+
+  x[0] = 0;
+
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/data-racer/084_race_on_shared/main.cu

@@ -1,0 +1,13 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel()
+{
+	__shared__ int A[8];
+	A[0] = threadIdx.x;
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/data-racer/099_test8/main.cu

@@ -1,0 +1,14 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "cutil.h"
+
+__global__ void kernel(int	*p, int *ptr_a) {
+
+	ptr_a = p + threadIdx.x;
+
+//	assert(*ptr_a == 2);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/null-pointer/019_basic_assignment_fail/main.cu

@@ -1,0 +1,27 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(float *v, unsigned int size, unsigned int i)
+{
+    unsigned int tid = blockIdx.x * blockDim.x + threadIdx.x;
+    funcType f;
+
+    if (i == 1)
+      f = multiplyByTwo;
+    else if (i == 2)
+      f = divideByTwo;
+    else
+      f = NULL;
+
+    if (tid < size)
+    {
+        float x = (*f)(v, tid);
+        x += multiplyByTwo(v, tid);
+        v[tid] = x;
+    }
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/025_fail/main.cu

@@ -1,0 +1,12 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(void *data) {
+	unsigned int *idata = (unsigned int*) data;
+	idata[threadIdx.x] = bitreverse1(idata[threadIdx.x]);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/039_test7_fail/main.cu

@@ -1,0 +1,22 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(int* glob) {
+
+  int a;
+
+  int* p;
+
+  a = 0;
+
+  p = &a;
+
+  *p = threadIdx.x;
+
+  glob[*p] = threadIdx.x;
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/050_fail/main.cu

@@ -1,0 +1,11 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(int* x) {
+  x[threadIdx.x] = A[threadIdx.x] + B[0]; //permanece constante por ser muito grande. N < 1024 não permanece
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/051_fail/main.cu

@@ -1,0 +1,14 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel() {
+
+  __shared__ int A[10][10];
+  A[threadIdx.y][threadIdx.x] = 2;
+  assert(A[threadIdx.y][threadIdx.x]!=2);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/053_misc2/main.cu

@@ -1,0 +1,13 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(volatile int* p)
+{
+    //__assert(__no_read(p));
+    p[threadIdx.x] = threadIdx.x;
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/058_noraceduetoreturn/config.h

@@ -1,0 +1,15 @@
+
+/* Global configuration parameter */
+
+#ifndef _CONFIG_H_
+#define _CONFIG_H_
+
+// should be power of two
+#define  MAX_THREADS_BLOCK                256
+
+#define  MAX_SMALL_MATRIX                 512
+#define  MAX_THREADS_BLOCK_SMALL_MATRIX   512
+
+#define  MIN_ABS_INTERVAL                 0 // 5.0e-37
+
+#endif // #ifndef _CONFIG_H_

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/058_noraceduetoreturn_fail/main.cu

@@ -1,0 +1,14 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel(float* A) {
+
+  int y = bar(A);
+  A[threadIdx.x]=y;
+
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/059_nestedinline/main.cu

@@ -1,0 +1,11 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel() {
+  g();
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/PUG/supported/pass/066_atomicCAS/main.cu

@@ -1,0 +1,16 @@
+#ifndef _KERNEL_H_ 
+#define _KERNEL_H_
+
+#include "config.h"
+#include "my_cutil.h"
+
+__global__ void kernel (int* A, unsigned int* B, unsigned long long int* C)
+{
+  atomicCAS(A,2,11);
+
+  atomicCAS(B,5,1);
+
+  atomicCAS(C,7,3);
+}
+
+#endif

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-civl-v1.cpp

@@ -1,0 +1,185 @@
+#include <algorithm>
+#include <array>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <memory>
+#include <regex>
+#include <string>
+#include <vector>
+#include <cstdio>
+#include <dirent.h>
+#include <unistd.h>
+
+using namespace std;
+
+#define BILLION 1E9
+#define BUFFER_SIZE 128
+#define CURRENT_FOLDER_SYMLINK "."
+#define CIVL_COMMAND "civl verify main.cu"
+#define PARENT_FOLDER_SYMLINK ".."
+#define PATH_SEPARATOR "/"
+#define PIPE_READ_MODE "r"
+#define PROPERTY_HOLDS "The standard properties hold for all executions."
+#define RESULT_DESCRIPTION "=== Result ==="
+#define TEST_DESCRIPTION_DELIMITER '^'
+#define TEST_DESCRIPTION_FILE "test.desc"
+#define VERIFICATION_FAILED_MESSAGE "VERIFICATION FAILED"
+#define VERIFICATION_SUCCESSFUL_MESSAGE "VERIFICATION SUCCESSFUL"
+
+vector<string> execute_command(const char *command) {
+    array<char, BUFFER_SIZE> buffer;
+    vector<string> result;
+    shared_ptr<FILE> pipe(popen(command, PIPE_READ_MODE), pclose);
+    if (pipe)
+    {
+        while (!feof(pipe.get()))
+        {
+            if (fgets(buffer.data(), BUFFER_SIZE, pipe.get()) != nullptr)
+            {
+                auto data = string(buffer.data());
+                result.push_back(data.substr(0, data.length() - 1));
+            }
+        }
+    }
+    return result;
+}
+
+vector<string> list_directories()
+{
+    DIR *dir;
+    struct dirent *ent;
+    vector<string> directories;
+    if ((dir = opendir(CURRENT_FOLDER_SYMLINK)) != nullptr)
+    {
+        while ((ent = readdir(dir)) != nullptr)
+        {
+            if (ent->d_type == DT_DIR)
+            {
+                auto current_directory = string(ent->d_name);
+                if (current_directory.compare(CURRENT_FOLDER_SYMLINK) != 0 &&
+                    current_directory.compare(PARENT_FOLDER_SYMLINK) != 0)
+                    directories.push_back(current_directory);
+            }
+        }
+        closedir(dir);
+        sort(directories.begin(), directories.end());
+    }
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-civl-v2.cpp

@@ -1,0 +1,175 @@
+/*******************************************************************
+
+ Author: Felipe R. Monteiro
+
+ Date: Jan 2017
+
+ Description: Verify ESBMC-GPU test suite with CIVL and get its
+              verification time
+
+ \*******************************************************************/
+
+#include <iostream>
+#include <string>
+#include <fstream>
+#include <list>
+#include <time.h>
+#include <stdlib.h>
+#include <unistd.h>
+using namespace std;
+
+#define BILLION 1E9
+#define SUCCESSFUL 1
+#define FAILED 2
+
+unsigned short int CORRECT = 0;
+unsigned short int INCORRECT = 0;
+unsigned short int FALSE_CORRECT = 0;
+unsigned short int FALSE_INCORRECT = 0;
+unsigned short int NOT_SUPPORTED = 0;
+
+/*
+   Function: getDirectoryList()
+   Description: Get a list of directory names
+*/
+list<string> getDirectoryList(){
+
+   string dir;
+	ifstream directories;
+   list<string> listOfDirectories;
+
+   system("ls >> directories.txt");
+   
+   directories.open("directories.txt");
+   while(!(directories.eof()))
+   {
+	   getline(directories,dir);
+      if(dir != "directories.txt" && dir != "" && dir != "make.sh" && dir != "running-civl.cpp" && dir != "run" && dir != "running-civl.cpp~")
+	      listOfDirectories.push_back(dir);
+   }
+	directories.close();
+   system("rm directories.txt");
+
+   return listOfDirectories;
+}
+
+/*
+   Function: run_CIVL()
+   Description: Analyse verification results
+*/
+void run_CIVL(){
+   string res;
+   string testDesc;
+	ifstream results;
+   int tempResults = -1;
+
+   system("civl verify main.cu >> results.txt"); // to run CIVL
+   
+   results.open("results.txt");
+   while(!(results.eof()))
+   {
+	   getline(results,res);
+         
+      if(res == "=== Result ==="){
+         getline(results,res);
+         if(res == "The standard properties hold for all executions."){
+            tempResults = SUCCESSFUL;
+         } else {
+            tempResults = FAILED;
+	 }
+	 break;
+      }
+   }
+   results.close();
+   res.clear();
+   system("rm results.txt");
+
+   if (tempResults == FAILED)
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-esbmc-gpu-v2.cpp

@@ -1,0 +1,192 @@
+#include <algorithm>
+#include <array>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <memory>
+#include <regex>
+#include <string>
+#include <vector>
+#include <cstdio>
+#include <dirent.h>
+#include <unistd.h>
+
+using namespace std;
+
+#define BILLION 1E9
+#define BUFFER_SIZE 128
+#define CURRENT_FOLDER_SYMLINK "."
+#define ESBMC_GPU_RUN_COMMAND_BASE "esbmc-gpu main.cu "
+#define LIBRARIES_INPUT_PARAMETER "-I ~/libraries/" 
+#define PARENT_FOLDER_SYMLINK ".."
+#define PATH_SEPARATOR "/"
+#define PIPE_READ_MODE "r"
+#define TEST_DESCRIPTION_DELIMITER '^'
+#define TEST_DESCRIPTION_FILE "test.desc"
+#define VERIFICATION_FAILED_MESSAGE "VERIFICATION FAILED"
+#define VERIFICATION_SUCCESSFUL_MESSAGE "VERIFICATION SUCCESSFUL"
+
+vector<string> execute_command(const char *command) {
+    array<char, BUFFER_SIZE> buffer;
+    vector<string> result;
+    shared_ptr<FILE> pipe(popen(command, PIPE_READ_MODE), pclose);
+    if (pipe)
+    {
+        while (!feof(pipe.get()))
+        {
+            if (fgets(buffer.data(), BUFFER_SIZE, pipe.get()) != nullptr)
+            {
+                auto data = string(buffer.data());
+                result.push_back(data.substr(0, data.length() - 1));
+            }
+        }
+    }
+    return result;
+}
+
+vector<string> list_directories()
+{
+    DIR *dir;
+    struct dirent *ent;
+    vector<string> directories;
+    if ((dir = opendir(CURRENT_FOLDER_SYMLINK)) != nullptr)
+    {
+        while ((ent = readdir(dir)) != nullptr)
+        {
+            if (ent->d_type == DT_DIR)
+            {
+                auto current_directory = string(ent->d_name);
+                if (current_directory.compare(CURRENT_FOLDER_SYMLINK) != 0 &&
+                    current_directory.compare(PARENT_FOLDER_SYMLINK) != 0)
+                    directories.push_back(current_directory);
+            }
+        }
+        closedir(dir);
+        sort(directories.begin(), directories.end());
+    }
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-esbmc-gpu-v3.cpp

@@ -1,0 +1,193 @@
+#include <algorithm>
+#include <array>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <memory>
+#include <regex>
+#include <string>
+#include <vector>
+#include <cstdio>
+#include <dirent.h>
+#include <unistd.h>
+
+using namespace std;
+
+#define BILLION 1E9
+#define BUFFER_SIZE 128
+#define CURRENT_FOLDER_SYMLINK "."
+#define ESBMC_GPU_RUN_COMMAND_BASE "esbmc-gpu 2>/dev/null main.cu --z3 "
+#define LIBRARIES_INPUT_PARAMETER "-I ~/libraries/" 
+#define PARENT_FOLDER_SYMLINK ".."
+#define PATH_SEPARATOR "/"
+#define PIPE_READ_MODE "r"
+#define TEST_DESCRIPTION_DELIMITER '^'
+#define TEST_DESCRIPTION_FILE "test.desc"
+#define VERIFICATION_FAILED_MESSAGE "VERIFICATION FAILED"
+#define VERIFICATION_SUCCESSFUL_MESSAGE "VERIFICATION SUCCESSFUL"
+
+vector<string> execute_command(const char *command) {
+    array<char, BUFFER_SIZE> buffer;
+    vector<string> result;
+    shared_ptr<FILE> pipe(popen(command, PIPE_READ_MODE), pclose);
+    if (pipe)
+    {
+        while (!feof(pipe.get()))
+        {
+            if (fgets(buffer.data(), BUFFER_SIZE, pipe.get()) != nullptr)
+            {
+                auto data = string(buffer.data());
+                result.push_back(data.substr(0, data.length() - 1));
+            }
+        }
+    }
+    return result;
+}
+
+vector<string> list_directories()
+{
+    DIR *dir;
+    struct dirent *ent;
+    vector<string> directories;
+    if ((dir = opendir(CURRENT_FOLDER_SYMLINK)) != nullptr)
+    {
+        while ((ent = readdir(dir)) != nullptr)
+        {
+            if (ent->d_type == DT_DIR)
+            {
+                auto current_directory = string(ent->d_name);
+                if (current_directory.compare(CURRENT_FOLDER_SYMLINK) != 0 &&
+                    current_directory.compare(PARENT_FOLDER_SYMLINK) != 0)
+                    directories.push_back(current_directory);
+            }
+        }
+        closedir(dir);
+        sort(directories.begin(), directories.end());
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-gklee-v4.cpp

@@ -1,0 +1,255 @@
+#include <algorithm>
+#include <array>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <memory>
+#include <regex>
+#include <string>
+#include <vector>
+#include <cstdio>
+#include <dirent.h>
+#include <unistd.h>
+
+using namespace std;
+
+#define BILLION 1E9
+#define BUFFER_SIZE 128
+#define NOT_SUPPORTED 0
+#define USE_KLEE 1
+#define USE_UCLIBC 2
+#define CURRENT_FOLDER_SYMLINK "."
+#define ERRORS_GENERATED_REGEX ".*\\d+ error[s]* generated."
+#define GKLEE_PRE_COMPILE_COMMAND_KLEE "gklee-nvcc main.cu -libc=klee"
+#define GKLEE_PRE_COMPILE_COMMAND_UCLIBC "gklee-nvcc main.cu -libc=uclibc"
+#define GKLEE_RUN_COMMAND "gklee main"
+#define KLEE_ERROR_REGEX "KLEE: ERROR:"
+#define PARENT_FOLDER_SYMLINK ".."
+#define PATH_SEPARATOR "/"
+#define PIPE_READ_MODE "r"
+#define SEGMENTATION_FAULT_REGEX ".*Segmentation fault      \\(core dumped\\) klee.*"
+#define TEST_DESCRIPTION_DELIMITER '^'
+#define TEST_DESCRIPTION_FILE "test.desc"
+#define VERIFICATION_FAILED_MESSAGE "VERIFICATION FAILED"
+#define VERIFICATION_SUCCESSFUL_MESSAGE "VERIFICATION SUCCESSFUL"
+
+vector<string> execute_command(const char *command) {
+    array<char, BUFFER_SIZE> buffer;
+    vector<string> result;
+    shared_ptr<FILE> pipe(popen(command, PIPE_READ_MODE), pclose);
+    if (pipe)
+    {
+        while (!feof(pipe.get()))
+        {
+            if (fgets(buffer.data(), BUFFER_SIZE, pipe.get()) != nullptr)
+            {
+                auto data = string(buffer.data());
+                result.push_back(data.substr(0, data.length() - 1));
+            }
+        }
+    }
+    return result;
+}
+
+vector<string> list_directories()
+{
+    DIR *dir;
+    struct dirent *ent;
+    vector<string> directories;
+    if ((dir = opendir(CURRENT_FOLDER_SYMLINK)) != nullptr)
+    {
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/benchmarks/scripts/old/running-pug-v2.cpp

@@ -1,0 +1,176 @@
+#include <algorithm>
+#include <array>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <memory>
+#include <regex>
+#include <string>
+#include <vector>
+#include <cstdio>
+#include <dirent.h>
+#include <unistd.h>
+
+using namespace std;
+
+#define BILLION 1E9
+#define BUFFER_SIZE 128
+#define CONFLICT_FOUND "!!!!!!! Found conflict: "
+#define CURRENT_FOLDER_SYMLINK "."
+#define NO_ERRORS "No races or synchronization errors were found."
+#define PUG_COMMAND "pug main.cu"
+#define PARENT_FOLDER_SYMLINK ".."
+#define PATH_SEPARATOR "/"
+#define PIPE_READ_MODE "r"
+#define TEST_DESCRIPTION_DELIMITER '^'
+#define TEST_DESCRIPTION_FILE "test.desc"
+#define VERIFICATION_FAILED_MESSAGE "VERIFICATION FAILED"
+#define VERIFICATION_SUCCESSFUL_MESSAGE "VERIFICATION SUCCESSFUL"
+
+vector<string> execute_command(const char *command) {
+    array<char, BUFFER_SIZE> buffer;
+    vector<string> result;
+    shared_ptr<FILE> pipe(popen(command, PIPE_READ_MODE), pclose);
+    if (pipe)
+    {
+        while (!feof(pipe.get()))
+        {
+            if (fgets(buffer.data(), BUFFER_SIZE, pipe.get()) != nullptr)
+            {
+                auto data = string(buffer.data());
+                result.push_back(data.substr(0, data.length() - 1));
+            }
+        }
+    }
+    return result;
+}
+
+vector<string> list_directories()
+{
+    DIR *dir;
+    struct dirent *ent;
+    vector<string> directories;
+    if ((dir = opendir(CURRENT_FOLDER_SYMLINK)) != nullptr)
+    {
+        while ((ent = readdir(dir)) != nullptr)
+        {
+            if (ent->d_type == DT_DIR)
+            {
+                auto current_directory = string(ent->d_name);
+                if (current_directory.compare(CURRENT_FOLDER_SYMLINK) != 0 &&
+                    current_directory.compare(PARENT_FOLDER_SYMLINK) != 0)
+                    directories.push_back(current_directory);
+            }
+        }
+        closedir(dir);
+        sort(directories.begin(), directories.end());
+    }
[…]

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixing typos on introduction
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
 %% The amssymb package provides various useful mathematical symbols
 \usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
 
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
 \title{ESBMC-GPU \\ A Context-Bounded Model Checking Tool \\ to Verify CUDA Programs}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \address[label1]{}
-%% \address[label2]{}
 
@@
 \begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
 GPU verification \sep formal verification \sep model checking \sep CUDA
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
 
@@
 A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit state-space exploration~\cite{cordeiro:2011,cordeiro:2012}.
-In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, space-state exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}.
+In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, state-space exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}.
 As a consequence, redundant interleavings are eliminated, without ignoring a program's behavior.

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixing typo
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-ESBMC-GPU builds on top of ESBMC, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.
+ESBMC-GPU is built on top of ESBMC, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.
 

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/comments/comments.tex
updating comments
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
   \begin{itemize}
-  \item space-state exploration $\rightarrow$ state-space exploration;
-  \item ESBMC-GPU builds on top of ESBMC $\rightarrow$ ESBMC-GPU is built on top of ESBMC;
+  \item \sout{space-state exploration $\rightarrow$ state-space exploration;}
+  \item \sout{ESBMC-GPU builds on top of ESBMC $\rightarrow$ ESBMC-GPU is built on top of ESBMC;}
   \end{itemize}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixing typos on page 3
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 CUDA Operational Model.} An operational model for CUDA libraries that provides support to CUDA functionalities, in conjunction with ESBMC, as shown in Fig.~\ref{figure:arch}.
-Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, consists in an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf.
+Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, consists of an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf.
 Sec.~\ref{sec:features}).
@@
 Two-threads Analysis.} Similarly to GPUVerify~\cite{betts:2012} and PUG~\cite{Li:2010}, ESBMC-GPU also reduces the number of threads (to only two elements), during the verification of CUDA programs, by considering a NVIDIA Fermi GPU architecture, in order to improve verification time and avoid the state-space explosion problem.
-Besides, in CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check.
+In CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check.
 Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.
@@
 \noindent \textbf{3.
-State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also reduce state space, based on SHA256 hashes \cite{FIPS:2002}.
+State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also reduce the state space, based on SHA256 hashes \cite{FIPS:2002}.
 In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states.

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixing typos on page 4
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node.
-Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings has been systematically explored~\cite{pereira:2016}.
+Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings have been systematically explored~\cite{pereira:2016}.
 
@@
 In addition, an assertion checks if the result matches to the expected postcondition (line $19$).
-Therefore, the verification of this program through ESBMC-GPU produces $54$ successful and $3$ failed interleavings.
+The verification of this program through ESBMC-GPU produces $54$ successful and $3$ failed interleavings.
 For instance, one possible failed interleaving is represented by the threads executions $t_0 : a[1] = 0$; $t_1 : a[2] = 1$, where $a[2] = 1$ represents an incorrect access to the array index $a$.

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/comments/comments.tex
updating comments
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \begin{itemize}
-\item Page 2
+\item \sout{Page 2}
   \begin{itemize}
@@
   \end{itemize}
-\item Page 3
+\item \sout{Page 3}
   \begin{itemize}
-  \item consists in $\rightarrow$ consists of;
-  \item Besides, in: remove ``Besides'';
-  \item reduce state space $\rightarrow$ reduce the state space;
+  \item \sout{consists in $\rightarrow$ consists of};
+  \item \sout{Besides, in: remove ``Besides'';}
+  \item \sout{reduce state space $\rightarrow$ reduce the state space;}
   \end{itemize}
@@
   \item if multiple threads perform unsynchronized 90 access to shared-memory locations: to the same memory locations, it is crucial to mention that;
-  \item has been systematically explored $\rightarrow$ have been systematically explored;
-  \item Therefore: the next conclusion does not follow logically from the previous statements, so ``therefore'' should be removed here;
+  \item \sout{has been systematically explored $\rightarrow$ have been systematically explored;}
+  \item \sout{Therefore: the next conclusion does not follow logically from the previous statements, so ``therefore'' should be removed here;}
   \end{itemize}
@@
   \begin{itemize}
-  \item Figure 2: between lines 17 and 18, the contents of {\tt dev\_a} should be copied to a with a {\tt cudaMemcpy}, otherwise the subsequent check is quite useless;
+  \item Figure 2: between lines 17 and 18, the contents of {\tt dev\_a} should be copied to {\tt a} with a {\tt cudaMemcpy}, otherwise the subsequent check is quite useless;
   \end{itemize}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixing typos on page 6
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.
-In addition, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
+On the other hand, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
 
@@
 ESBMC-GPU marks the first application of an SMT-based context-BMC tool that recognizes CUDA directives~\cite{pereira:2016}.
-Besides, it further simplifies verification models and provides fewer incorrect results, if compared with GKLEE, GPUVerify, and CIVL.
+Besides, it further simplifies verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.
 Finally, it presents improved ability to detect array out-of-bounds and data race violations.

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
uploading reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
\usepackage{enumerate}
\usepackage[nodayofweek]{datetime}
\usepackage{xcolor}
\usepackage{natbib}
\newcommand{\Real}{\mathbb{R}}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{Manuscript Reference Number: SCICO-D-17-00062}
\rhead{Page \thepage\ of \pageref{LastPage}}
\lfoot{}
\rfoot{}

\pagenumbering{roman}
\begin{document}
\setcounter{page}{1}
\thispagestyle{empty}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{logo_ufam.png}}
\end{minipage}}} \hspace*{0.5cm}
\begin{minipage}{11.0cm}{\sc \large \textbf{FEDERAL UNIVERSITY OF AMAZONAS}}\\
{\sc Faculty of Technology\\}
{\sc Department of Electronics and Computing}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Mr. Felipe R. Monteiro\\
Federal University of Amazonas\\
Department of Electronics and Computing\\
Avenida General Rodrigo Ot\'avio, 6200, Coroado\\
69077-000 - Manaus - AM - BRAZIL\\
{\em e-mail: felipemonteiro@ufam.edu.br}
\end{minipage}
\end{flushright}

\vi \vi
\hspace*{\fill}{\small Manaus, \today.}
\vi \vi

\hoffset -1.5cm \voffset .5cm

\begin{flushleft}
Dr. Mark van den Brand\\
   Editor-in-Chief \\
   Science of Computer Programming Software Track
    \end{flushleft}
\vi

\begin{flushleft}
\textbf{REF.:} SCICO-D-17-00062
    \end{flushleft}
\vi \vi

\indent Dear Dr. Mark van den Brand,
\vi

We thank you for asking us to comply with the Referees' reports. We are submitting the revised version of our paper entitled as ``{\em ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs},'' by Felipe R. Monteiro, Erickson H. da S. Alves, Isabela S. Silva, Hussama A. Ismail, Lucas C. Cordeiro, and Eddie B. de Lima Filho.

The details of all changes made to our paper and also our response to each comment raised by the reviewers have been included at the end of this letter.

\vi\vi

\indent
~~~~~~~~~~~~~~~~~~~~~~~~~~~Sincerely,\\

\begin{quote}
\begin{quote}
\begin{flushright}

\vi
\vi

Felipe R. Monteiro~~~~~~~
\end{flushright}
\end{quote}
\end{quote}

\hoffset -1.5cm \voffset .5cm

% =============================================
\newpage
\subsection*{Authors' Response to the Review Comments on manuscript ``{ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}'' -- Manuscript Reference Number: SCICO-D-17-00062}

\vi

The authors thank the anonymous reviewers for their valuable and constructive comments, which helped us significantly improve the quality of our manuscript.

\vi

Our manuscript has been carefully revised and the detailed changes are highlighted in \textcolor{blue}{\bf blue color}, along the original text. Our responses to all comments, which were raised by the reviewers, are given in the sequel, using \textcolor{blue}{\bf blue color} as well.

\vi

We hope that these revisions in the manuscript and our responses will be sufficient to make our manuscript suitable for publication in the Journal of Science of Computer Programming (Software Track).

% =============================================
\newpage

\vi
\subsection*{Reviewer \#1 comments followed by the authors' answers:}

\begin{quote}

{\bf Comments to the Authors}

The authors present an extension of the ESBMC tool, to verify CUDA GPU kernels by means of bounded model checking. This is an interesting and relevant topic, since GPUs are used increasingly often to speed up applications significantly. However, programming GPU software can be very tedious and error-prone. A formal verifier of GPU code can make the development process much more efficient.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Response.}
\end{quote}

{\bf Reviews:}

{\bf (1)} The present submission provides an ok overview of the capabilities of the tool, in my opinion, but the authors never really address the details to fully understand what the tool can and cannot do. For instance, when discussing the experimental evaluation, it is not explained what the source of the bugs is in the various benchmarks.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Here.}
\end{quote}

  \begin{itemize}
  \item Are there clear differences in the types of bugs that the different tools being compared can detect?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item Which types could ESBMC-GPU handle in particular better than the other tools, and why?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item To what extent does ESBMC-GPU support CUDA?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item Can ESBMC-GPU handle variables in different types of memory?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item (Shared, global, thread registers?), and inter-warp communication, for instance?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item In that respect, what does COM cover, and what remains to be added?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \item Is the memory model for the different types of memory complete?

  \begin{quote}
  \textcolor{blue}{\textbf{Response:} Here.}
  \end{quote}

  \end{itemize}

{\bf (2)} Another paragraph that raises questions is the second on page 3. It seems to say that COM uses pthread functions to simulate GPU kernels. Is that so, and why is it necessary to simulate them? Is this safe to do, because some features in CUDA are not available in C/C++. How do the authors handle those, or do they not?

\begin{quote}
\textcolor{blue}{\textbf{Response:} Here.}
\end{quote}

{\bf (3)} At some point, the authors mention in passing that Microsoft C++ AMP projects have been used for the evaluation. Does this mean that this language is also supported, or are the benchmarks translated?

\begin{quote}
\textcolor{blue}{\textbf{Response:} Here.}
\end{quote}

{\bf (4)} Concluding, I would like to see an overall more detailed discussion of the capabilities of the tool.

\begin{quote}
\textcolor{blue}{\textbf{Response:} Here.}
\end{quote}

{\bf (5)} Detailed comments:
\begin{itemize}
\item Page 2
  \begin{itemize}
  \item space-state exploration $\rightarrow$ state-space exploration;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item ESBMC-GPU builds on top of ESBMC $\rightarrow$ ESBMC-GPU is built on top of ESBMC;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \end{itemize}
\item Page 3
  \begin{itemize}
  \item consists in $\rightarrow$ consists of;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item Besides, in: remove ``Besides'';

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item reduce state space $\rightarrow$ reduce the state space;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \end{itemize}
\item Page 4
  \begin{itemize}
  \item if multiple threads perform unsynchronized 90 access to shared-memory locations: to the same memory locations, it is crucial to mention that;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item has been systematically explored $\rightarrow$ have been systematically explored;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item Therefore: the next conclusion does not follow logically from the previous statements, so ``therefore'' should be removed here;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \end{itemize}
\item Page 5
  \begin{itemize}
  \item Figure 2: between lines 17 and 18, the contents of {\tt dev\_a} should be copied to {\tt a} with a {\tt cudaMemcpy}, otherwise the subsequent check is quite useless;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \end{itemize}
\item Page 6
  \begin{itemize}
  \item In addition $\rightarrow$ On the other hand;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item Besides, it further simplifies verification models: further compared to what? How does it make it easier?

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item if compared: remove ``if'';

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item stream interleaving: what is this?

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \item  to reduce the number of thread interleavings: in what respect? Not reducing the number of threads, right? You already use the 2-thread simplification;

    \begin{quote}
    \textcolor{blue}{\textbf{Response:} Response}
    \end{quote}

  \end{itemize}
\end{itemize}

\end{quote}

\label{LastPage}

% \printbibliography

\end{document}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} Response.}
+\textcolor{blue}{\textbf{Response:} We really appreciate the reviewer's summary about our work, which fairly describes the main contributions of our manuscript.
+We also appreciate the reviewer's comments and also his/her effort to provide relevant comments to improve our paper.
+We have carefully revised our manuscript, considering all valuable and constructive comments/suggestions.}
 \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} We have fixed this typo.}
     \end{quote}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
pointing out changes in manuscript
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit state-space exploration~\cite{cordeiro:2011,cordeiro:2012}.
+A distinct feature of ESBMC-GPU, when compared with other approaches \cite{betts:2012,Li:2010,Li:2012,civl:2015}, is the use of Bounded Model Checking (BMC)~\cite{Biere:2009} allied to Satisfiability Modulo Theories (SMT)~\cite{Barrett:2009}, with explicit \textcolor{blue}{state-space exploration}~\cite{cordeiro:2011,cordeiro:2012}.
 In summary, concurrency problems are tackled, up to an unwinding bound, while each interleaving itself is symbolically handled; however, even with BMC, state-space exploration may become a very time-consuming task, which is alleviated through state hashing and Monotonic Partial Order Reduction (MPOR)~\cite{KahlonWG09}.
@@
 
-ESBMC-GPU is built on top of ESBMC, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.
+\textcolor{blue}{ESBMC-GPU is built on top of ESBMC}, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.
 
@@
 CUDA Operational Model.} An operational model for CUDA libraries that provides support to CUDA functionalities, in conjunction with ESBMC, as shown in Fig.~\ref{figure:arch}.
-Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, consists of an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf.
+Such an approach, which was previously attempted in the verification of C++ programs~\cite{ramalho:2013, monteiro:2015, garcia:2016, monteiro:2017}, \textcolor{blue}{consists of} an abstract representation that reliably approximates the CUDA library's semantics; however, COM incorporates pre- and post-conditions into verification processes, which enables ESBMC-GPU to verify specific properties (cf.
 Sec.~\ref{sec:features}).
@@
 \noindent \textbf{3.
-State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also reduce the state space, based on SHA256 hashes \cite{FIPS:2002}.
+State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and also \textcolor{blue}{reduce the state space}, based on SHA256 hashes \cite{FIPS:2002}.
 In particular, its symbolic state hashing approach computes a summary for a particular state that has already been explored and then indexes the resulting set, in order to reduce the generation of redundant states.
@@
 In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node.
-Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings have been systematically explored~\cite{pereira:2016}.
+Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings \textcolor{blue}{have been systematically explored}~\cite{pereira:2016}.
 
@@
 %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.
-On the other hand, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
+\textcolor{blue}{On the other hand}, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
 

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating conclusion
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 ESBMC-GPU marks the first application of an SMT-based context-BMC tool that recognizes CUDA directives~\cite{pereira:2016}.
-Besides, it further simplifies verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.
-Finally, it presents improved ability to detect array out-of-bounds and data race violations.
+\textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations.
 Future work aims to support stream interleaving and implement further techniques to reduce the number of thread interleavings, by taking into account GPU symmetry.

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>
adding comments to reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} We have fixed this typo.}
+    \textcolor{blue}{\textbf{Response:} We have implemented as suggested.}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} We explicitly indicate now that ESBMC-GPU applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models.}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: which reference would you suggest to use here?}}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: what would be the next step to reduce the number of thread interleavings? Improve MPOR?}}
     \end{quote}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating section Functionalities
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access to shared-memory locations; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
+Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access \textcolor{blue}{to the same memory locations}; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
 

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
   \begin{itemize}
-  \item if multiple threads perform unsynchronized 90 access to shared-memory locations: to the same memory locations, it is crucial to mention that;
+  \item if multiple threads perform unsynchronized access to shared-memory locations: to the same memory locations, it is crucial to mention that;
 
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} We have implemented as suggested.}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} Response}
+    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: To be honest, I didn't get it.
+Do you think this is correct?
+Could professor Ricardo Pereira help us answer this question appropriately?}}
     \end{quote}

### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
adding comments to reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} Here.}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: haven't we explained that in {\bf Limitations} subsection on page 6?}}
 \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Hussama: could you answer this question?
+I believe we need the list of properties that each tool can verify and then we can see the differences between them.
+We should display such information on the website, in addition to the manuscript.}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: you have more experience with the experimental evaluation. What do you think?}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: would you be able to tell me that?}}
   \end{quote}
 
-  \item Can ESBMC-GPU handle variables in different types of memory?
+  \item Can ESBMC-GPU handle variables in different types of memory? (Shared, global, thread registers?), and inter-warp communication, for instance?
 
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
-  \end{quote}
-
-  \item (Shared, global, thread registers?), and inter-warp communication, for instance?
-
-  \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: Shall we reference your work about verifying multithread ANSI-C programs?}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: any insights about coverage?}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} Here.}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: I didn't get it. Did you?}}
   \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} Here.}
+\textcolor{blue}{\textbf{Response:}}
 \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} Here.}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: does he mean C++? Shall we reference the paper from ECBS 2013 about ESBMC++?}}
 \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} Here.}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: don't we explain that on Functionalities section? Any ideia what else we could add?}}
 \end{quote}

### 2017-05-12 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
added answers to Felipe's questions
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: haven't we explained that in {\bf Limitations} subsection on page 6?}}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: haven't we explained that in {\bf Limitations} subsection on page 6?} The reviewer wants to know more details about the type of bugs that we have found in the benchmarks.
+May I ask you to take a look at our journal paper?
+I think we describe the type of bugs there.
+}
 \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: Shall we reference your work about verifying multithread ANSI-C programs?}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: Shall we reference your work about verifying multithread ANSI-C programs? Yes, please. Also reference our journal paper about esbmc-gpu.}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: I didn't get it. Did you?}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: I didn't get it.
+Did you?
+Can you cite our TACAS 2014 paper?
+I suggest you to read that paper which briefly explains our memory model.
+Then you can properly reply to the reviewer and cite that paper.}}
   \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: does he mean C++? Shall we reference the paper from ECBS 2013 about ESBMC++?}}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: does he mean C++?
+Shall we reference the paper from ECBS 2013 about ESBMC++?
+Yes, please.
+Mention that we support C++ and cite our ECBS 2013 paper.}}
 \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: don't we explain that on Functionalities section? Any ideia what else we could add?}}
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: don't we explain that on Functionalities section?
+Any ideia what else we could add?
+The problem is that we have page limit, right?
+Can you please mail the editor and check if we can exceed the page limit to properly handle the reviewer comments?}}
 \end{quote}
@@
 Do you think this is correct?
-Could professor Ricardo Pereira help us answer this question appropriately?}}
+Could professor Ricardo Pereira help us answer this question appropriately?
+Can you please check this with Isabela?
+Isabela had worked quite hard on the experimental evaluation and examples of our papers.}}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: which reference would you suggest to use here?}}
+    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: which reference would you suggest to use here?
+Add a reference to some book on GPU.
+You can also briefly explain the term and then cite the reference.}}
     \end{quote}
@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: what would be the next step to reduce the number of thread interleavings? Improve MPOR?}}
+    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: what would be the next step to reduce the number of thread interleavings?
+Improve MPOR?
+You can mention that we intend to use Abstract Interpretation to further reduce the model and the number of threads.
+Then you can cite our TACAS 2017 paper}}
     \end{quote}

### 2017-05-12 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
fixed typo
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec).
-This is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}.
+\textcolor{blue}{On the one hand}, this is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}.
 %In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
adding Eddie's comments
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-We thank you for asking us to comply with the Referees' reports.
-We are submitting the revised version of our paper entitled as ``{\em ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs},'' by Felipe R.
+We thank you for asking us to comply with the Referees' reports, which allowed us to submit a revised version of our paper, entitled as ``{\em ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs},'' by Felipe R.
 Monteiro, Erickson H.
@@
 
-We hope that these revisions in the manuscript and our responses will be sufficient to make our manuscript suitable for publication in the Journal of Science of Computer Programming (Software Track).
+We hope that those revisions in the manuscript and our responses will be sufficient to make our manuscript suitable for publication in the Journal of Science of Computer Programming (Software Track).
 
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} We really appreciate the reviewer's summary about our work, which fairly describes the main contributions of our manuscript.
-We also appreciate the reviewer's comments and also his/her effort to provide relevant comments to improve our paper.
-We have carefully revised our manuscript, considering all valuable and constructive comments/suggestions.}
+\textcolor{blue}{\textbf{Response:} We really appreciate the reviewer's summary about our work, which fairly describes its main contributions.
+We also appreciate the reviewer's comments and also his/her effort to provide a relevant discussion, with the goal of improving our paper.
+As a consequence, we have carefully revised our manuscript, considering all suggestions.}
 \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: haven't we explained that in {\bf Limitations} subsection on page 6?} The reviewer wants to know more details about the type of bugs that we have found in the benchmarks.
+\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: haven't we explained that in {\bf Limitations} subsection on page 6?} \textcolor{green}{Eddie: It seems he meant a detailed explanation about each bug.
+Do you remember what we did in our former articles?
+We described each benchmark, which included clarifications about existing bugs...
+We did not do that here...
+We only said ``covers basic functions commonly used by real CUDA applications'', but we did not clearly showed what...
+Maybe a brief explanation of each pat of our test set...} The reviewer wants to know more details about the type of bugs that we have found in the benchmarks.
 May I ask you to take a look at our journal paper?
@@
 I believe we need the list of properties that each tool can verify and then we can see the differences between them.
-We should display such information on the website, in addition to the manuscript.}}
+We should display such information on the website, in addition to the manuscript.}\textcolor{green}{Eddie: We only do that for our approach (a brief discussion), at the beginning of page 6, above section performance...
+We need to do that for the other tools...
+Do we have room for that?}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: you have more experience with the experimental evaluation. What do you think?}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: you have more experience with the experimental evaluation.
+What do you think?}\textcolor{green}{Eddie: Maybe, the comment above answers that}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: would you be able to tell me that?}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: would you be able to tell me that?} \textcolor{green}{Eddie: A simple paragraph, just commenting on that and preferably giving a percentage...}}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: any insights about coverage?}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Isabela: any insights about coverage?} \textcolor{green}{Eddie: A simple text, just commenting on that and preferably giving a percentage...}}
   \end{quote}
@@
 I suggest you to read that paper which briefly explains our memory model.
-Then you can properly reply to the reviewer and cite that paper.}}
+Then you can properly reply to the reviewer and cite that paper.}\textcolor{green}{Eddie: Maybe he is talking about shared, register, global and local...
+We need to show that our memory model correctly represent such types...}}
   \end{quote}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:}}
+\textcolor{blue}{\textbf{Response:}\textcolor{green}{Eddie: We have to comment on kernels, blocks and threads, in order to clarify that, showing that it it feasible ...}}
 \end{quote}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating code in Fig. 2
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 #define THREADS 2
+
 __global__ void kernel(int *A) {
@@
 }
+
 int main(){
@@
 	ESBMC_verify_kernel(kernel, BLOCKS, THREADS, dev_a);
+	cudaMemcpy(a, dev_a , size , cudaMemcpyDeviceToHost);
 	for (int i = 0; i < THREADS; i++)
@@
 	cudaFree(dev_a);
+	free(a);
 	return 0;

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: To be honest, I didn't get it.
-Do you think this is correct?
-Could professor Ricardo Pereira help us answer this question appropriately?
-Can you please check this with Isabela?
-Isabela had worked quite hard on the experimental evaluation and examples of our papers.}}
+    \textcolor{blue}{\textbf{Response:} Indeed, there was a code line missing from the figure, we have fixed it now.}
     \end{quote}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating conclusion
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations.
-Future work aims to support stream interleaving and implement further techniques to reduce the number of thread interleavings, by taking into account GPU symmetry.
+Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques to reduce the number of thread interleavings, by taking into account GPU symmetry.
 

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: which reference would you suggest to use here?
-Add a reference to some book on GPU.
-You can also briefly explain the term and then cite the reference.}}
+    \textcolor{blue}{\textbf{Response:} We have replaced this term by ``CUDA (parallel) streams and events'' and indicated ``Professional CUDA C Programming'' book as reference.}
     \end{quote}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating conclusion
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

--- 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
@@
 \textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations.
-Future work aims to support stream interleaving and implement further techniques to reduce the number of thread interleavings, by taking into account GPU symmetry.
+Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques \textcolor{blue}{(e.g., Abstract Interpretation~\cite{Rocha2017})} to reduce the number of thread interleavings, by taking into account GPU symmetry.
 
@@
 \newblock {\em{SMT-Based Context-Bounded Model Checking for CUDA Programs}}.
-\newblock Concurrency Computat.: Pract. Exper. 2016 (to appear).
+\newblock Concurrency Computat.: Pract. Exper. 2016.
 
@@
 
+\bibitem{Rocha2017}
+\textcolor{blue}{Rocha, W., Rocha, H., Ismail, H., Cordeiro, L., and Fischer, B.}
+\newblock {\em{\textcolor{blue}{DepthK: A k-Induction Verifier Based on Invariant Inference for C Programs}}}.
+\newblock \textcolor{blue}{Tools and Algorithms for the Construction and Analysis of Systems: 23rd International Conference, TACAS 2017, Held as Part of the European Joint Conferences on Theory and Practice of Software, ETAPS 2017, Uppsala, Sweden.}
+\newblock \textcolor{blue}{{Springer Berlin Heidelberg}, 360--364, 2017.}
+
 \end{thebibliography}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
     \begin{quote}
-    \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: what would be the next step to reduce the number of thread interleavings?
-Improve MPOR?
-You can mention that we intend to use Abstract Interpretation to further reduce the model and the number of threads.
-Then you can cite our TACAS 2017 paper}}
+    \textcolor{blue}{\textbf{Response:} We intend to use Abstract Interpretation to further reduce the number of thread interleavings (i.e., the model to be verified).}
     \end{quote}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
minor changes
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-\vspace{-1mm}
+%\vspace{-1mm}
 
@@
 
-\vspace{-3mm}
+%\vspace{-3mm}
 
@@
 
-\vspace{-2mm}
+%\vspace{-2mm}
 
@@
 Thus, ESBMC-GPU applies a combination of processing methods used by Central Processing Units (CPUs) and the POSIX library, where thread instructions can interleave to create execution paths.
-Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).
+Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\
 %Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.
@@
 In CUDA programs, whilst threads execute the same parametrized kernel, only two of them are necessary for conflict check.
-Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.
+Thus, such an analysis ensures that errors ({\it e.g.}, data races) detected between two threads, in a given subgroup and due to unsynchronized accesses to shared variables, are enough to justify a property violation~\cite{pereira:2016}.\\
 
@@
 When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT).
-This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.% \textcolor{red}{similares?
+This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\% \textcolor{red}{similares?
 entao, qual eh o impacto de se verificar apenas um?
@@
 
-\vspace{-4mm}
+%\vspace{-4mm}
 
@@
 First of all, users must replace the default kernel call (line $16$) by an intrinsic function of ESBMC-GPU (line $17$).
-Then, the resulting CUDA program can be passed to the command-line version of ESBMC-GPU, as follows: \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},  where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.
+Then, the resulting CUDA program can be passed to the command-line version of ESBMC-GPU, as follows: %\texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>} {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
 
-%\begin{center}
+\begin{center}
 %\vspace{-2 mm}
-%\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}
+\noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}
 
-%\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
+\noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
 %\vspace{-2 mm}
-%\end{center}
+\end{center}
+
+where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.
 
@@
 
-\vspace{-7mm}
+%\vspace{-7mm}
 
@@
 As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}.
-More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).
-%\textcolor{red}{e os outros?
+More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\ %\textcolor{red}{e os outros?
 quais sao as vantagens?
 esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}
-
-{\bf Limitations.} ESBMC-GPU was unable to correctly verify $24$ benchmarks, which are related to constant memory access ($2$\%), CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when passed as kernel call arguments ($4.5$\%).
-In addition, it only reported $3$\% of incorrect true and $1$\% of incorrect false results.
-
-%In addition, ESBMC-GPU had $23$ benchmarks that were not supported.
-These are related to constant memory access ($3$), the use of CUDA's specific libraries (\textit{e.g., curand.h}) ($7$), and the use of pointers to functions, structures, and char type variables used as kernel call arguments ($13$).
-
-{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $789.6$ sec.
-Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec).
-\textcolor{blue}{On the one hand}, this is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}.
-%In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.
-\textcolor{blue}{On the other hand}, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
-
-\vspace{-2mm}
 
@@
 
-\vspace{-8mm}
+{\bf Limitations.} ESBMC-GPU was unable to correctly verify $24$ benchmarks, which are related to constant memory access ($2$\%), CUDA's specific libraries ($4.5$\%), and the use of pointers to functions, structures, and {\tt char} type variables, when passed as kernel call arguments ($4.5$\%).
+In addition, it only reported $3$\% of incorrect true and $1$\% of incorrect false results.\\
+
+{\bf Performance.} MPOR resulted in a performance improvement of approximately $80\%$, by decreasing the verification time from $16$ to $3$ hours, while the two-threads analysis further reduced that to $789.6$ sec.
+Although such techniques have considerably improved the ESBMC-GPU's performance, it still takes longer than the other evaluated tools: GPUVerify ($98.36$ sec), GKLEE ($108.32$ sec), and CIVL ($708.52$ sec).
+\textcolor{blue}{On the one hand}, this is due to thread interleavings, which combine symbolic model checking with explicit state-space exploration~\cite{pereira:2016}.
+%In addition, even though ESBMC-GPU reports the highest verification time, it still presents the highest accuracy, with less than $6$ seconds per benchmark.%, which shows that it is a feasible tool regarding the verification of CUDA-based applications.
+\textcolor{blue}{On the other hand}, ESBMC-GPU still presents the highest accuracy, with less than $6$ seconds per benchmark.
 
@@
 %=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-%\vspace{-1mm}
 

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating section 2 and references
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-\textcolor{blue}{ESBMC-GPU is built on top of ESBMC}, which is an open source context-bounded model checker based on SMT solvers for ANSI-C/C$++$ programs~\cite{cordeiro:2011,cordeiro:2012}, and adds four essential models, as described below.
+\textcolor{blue}{ESBMC-GPU is built on top of ESBMC}, which is an open source context-bounded model checker based on SMT solvers for \textcolor{blue}{ANSI-C/C$++$ programs~\cite{cordeiro:2011, cordeiro:2012, ramalho:2013}}, and adds four essential models, as described below.
 
@@
 
+\bibitem{ramalho:2013}
+\textcolor{blue}{Ramalho, M., Freitas, M., Sousa, F., Marques, H., Cordeiro, L., and Fischer, B.}
+\newblock \textcolor{blue}{{\em{SMT-Based Bounded Model Checking of C++ Programs}}.}
+\newblock \textcolor{blue}{Proceedings of the $20^{th}$ Annual IEEE International Conference and Workshops on the Engineering of Computer Based Systems, ECBS'13.}
+\newblock \textcolor{blue}{{IEEE} Computer Society: Washington, DC, USA, 147--156, 2013.}
+
 \bibitem{Pereira15}
@@
 
-\bibitem{ramalho:2013}
-Ramalho, M., Freitas, M., Sousa, F., Marques, H., Cordeiro, L., and Fischer, B.
-\newblock {\em{SMT-Based Bounded Model Checking of C++ Programs}}.
-\newblock {{ECBS}} 2013; 147--156.
+%\bibitem{ramalho:2013}
+%Ramalho, M., Freitas, M., Sousa, F., Marques, H., Cordeiro, L., and Fischer, B.
+%\newblock {\em{SMT-Based Bounded Model Checking of C++ Programs}}.
+%\newblock {{ECBS}} 2013; 147--156.
 

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: does he mean C++?
-Shall we reference the paper from ECBS 2013 about ESBMC++?
-Yes, please.
-Mention that we support C++ and cite our ECBS 2013 paper.}}
+\textcolor{blue}{\textbf{Response:} Since 2013, ESBMC supports C++. We have added a reference to the paper ``SMT-Based Bounded Model Checking of C++ Programs'', which describes such work.}
 \end{quote}

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
clean up
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-%\vspace{-1mm}
-
 The Compute Unified Device Architecture (CUDA) is a development framework that makes use of the architecture and processing power of Graphics Processing Units (GPUs)~\cite{cuda:2012}.
@@
 Obviously, it results in higher verification times, but more errors can then be identified and later corrected, in another development cycle.
-
-%\textcolor{red}{seria interessante colocar um resumo sobre o que serah visto no resto do artigo?} \textcolor{blue}{Acho que nao temos espaco para isso} \textcolor{green}{Melhor nao colocarmos este resumo...}
 
@@
 
-%\vspace{-3mm}
-
 %=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 
-%\vspace{-1mm}
-
 \textcolor{blue}{ESBMC-GPU is built on top of ESBMC}, which is an open source context-bounded model checker based on SMT solvers for \textcolor{blue}{ANSI-C/C$++$ programs~\cite{cordeiro:2011, cordeiro:2012, ramalho:2013}}, and adds four essential models, as described below.
-
-%\vspace{-2mm}
 
@@
   \centering
-  \includegraphics[width=4.8in]{../figures/arch.pdf}
+  \includegraphics[width=5.5in]{../figures/arch.pdf}
   \caption{Overview of ESBMC-GPU's architecture.}
@@
 \end{figure}
-
-%\vspace{-2mm}
 
@@
 Particularly, COM simulates the behavior of kernel calls using pthread functions ({\it e.g.}, {\tt pthread\_create}) and combines that with ESBMC, in order to check data race and specific C/C++ programming language failures ({\it e.g.}, array out-of-bounds and pointer safety).\\
-%Additionally, ESBMC-GPU employs symbolic state hashing~\cite{Morse11}, monotonic partial order reduction (MPOR)~\cite{morse:2015}, and the two-thread analysis~\cite{pereira:2016}, in order to prune the state space exploration.
 
@@
 
-%\noindent \textbf{State Hashing.} ESBMC-GPU applies state hashing to further eliminate redundant interleavings and reduce state space, based on SHA256 hashes.\\%, due to the low the likelihood regarding collisions.
-\\
-%
-
 \noindent \textbf{3.
@@
 When this behavior is confirmed, which happens during the ESBMC-GPU's symbolic-execution procedure, then the current computation path does not need to be further explored in the associated reachability tree (RT).
-This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\% \textcolor{red}{similares?
-entao, qual eh o impacto de se verificar apenas um?
-eh bom colocar uma ref aqui e melhorar o texto} \textcolor{blue}{professor Lucas, devemos citar aqui a tese do Jeremy?} \textcolor{green}{Felipe podes citar o nosso journal do CPE e a tese do Jeremy}.
+This way, if ESBMC-GPU reaches such a state, {\it i.e.}, where a context switch can be taken ({\it e.g.}, before a global variable or synchronization primitive) and all shared/local variables and program counters are similar to another explored node, then ESBMC-GPU just considers that an identical node to be further explored, since reachability subtrees associated to them are also similar~\cite{pereira:2016,morse:2015}.\\
 
@@
 As a consequence, it is possible to determine whether interleaving pairs always lead to the same state and then remove duplicates in a reachability tree, without ignoring any program's behavior~\cite{morse:2015}.
-
-%In addition to both techniques,
-
-%\vspace{-4mm}
 
@@
 
-%\vspace{-1mm}
-
 Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access \textcolor{blue}{to the same memory locations}; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
@@
 Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings \textcolor{blue}{have been systematically explored}~\cite{pereira:2016}.
-
-%\textcolor{red}{eu nao sei como a disponibilidade de espaco estah, mas seria bom aqui um fechamento de como BMC checa tudo...} \textcolor{blue}{acredito que nao teremos espaco para isso} \textcolor{green}{Felipe, podes referenciar o nosso journal do CPE informando que la explicamos como o BMC checa tudo com mais detalhes.}
-
-%\vspace{-4mm}
 
@@
 
-%\vspace{-1mm}
-
 In this part, ESBMC-GPU usage is demonstrated, by using the CUDA program shown in Fig.~\ref{fig:illustrative-example}.
@@
 \begin{center}
-%\vspace{-2 mm}
 \noindent \texttt{esbmc-gpu} \emph{\tt <file>.cu} {\tt --unwind <$k$>}  {\tt --context-switch <$c$>}
@@
 \noindent {\tt --state-hashing} {\tt -I <path-to-CUDA-OM>},
-%\vspace{-2 mm}
 \end{center}
@@
 where {\tt $<$file$>$.cu} is the CUDA program, {\tt <$k$>} is the maximum loop unrolling, {\tt <$c$>} is a context-switch bound, {\tt --state-hashing} reduces redundant interleavings, and {\tt $<$path-to-} {\tt CUDA-OM$>$} is the location of the COM library.
-
-%\vspace{-0.1cm}
-
-%\textcolor{red}{esta parte ficou solta, pois nao houve explicacao de BMC e, por conseguinte, k (unwind)} \textcolor{blue}{concorco com o comentario, mas como nao temos espaco, estamos considerando que o revisor ja tera conhecimento de tecnicas mais conhecidas, como BMC} \textcolor{green}{sugiro novamente referenciar o journal do CPE.}
-%As one may see in Fig.~\ref{fig:illustrative-example}, this CUDA-based program has 1 block and 2 threads, respectively.
-Furthermore, its kernel (lines $8$ to $10$) stores in the variable $c$ the result of the sum of $a$ and $b$, which are passed as input arguments.
-However, there is a mistake in the assertion (line $26$), as $a + b$ must be equal to $c$, which is correctly detected by ESBMC-GPU detects an assert violation.
 
@@
 As shown in the main function, array positions are assigned with value $0$ (line $23$), and after the kernel call (line $26$), it is expected that $a[0] == 0$ and $a[1] == 1$.
-\textcolor{red}{$<$---- poderiamos deletar esse paragrafo?}
 
@@
 It is worth noticing that CIVL, ESBMC-GPU, and GKLEE are also able to detect this array out-of-bounds violation, but GPUVerify fails, as it reports a true incorrect result (missed bug).
-
-%\vspace{-3mm}
 
@@
 
-%\vspace{-7mm}
-
 %=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 As one may notice, the present experimental results show that ESBMC-GPU reached a successful verification rate of approximately $85\%$, while GKLEE, GPUVerify, and CIVL reported $72\%$, $50\%$, and $35\%$, respectively\footnote{All experimental results are available at {\tt http://esbmc.org/gpu/}}.
-More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\ %\textcolor{red}{e os outros?
-quais sao as vantagens?
-esse paragrafo ficou incompleto} \textcolor{blue}{estamos focando apenas nos aspectos do ESBMC porque nao temos espaco para discutir sobre os demais, isso estara na pagina online} \textcolor{green}{Concordo em colocar o texto explicativo na pagina online e referenciar.}
+More precisely, ESBMC-GPU supports the verification of benchmarks related to array bounds ($3$\%), assertive statements ($5$\%), data race ($11$\%), {\tt NULL} pointers ($3$\%), and other specific CUDA functionalities ($63$\%).\\
 
@@
   \centering
-  \includegraphics[width=3.9in]{../figures/experimental-results.pdf}
+  \includegraphics[width=4.5in]{../figures/experimental-results.pdf}
   \caption{Experimental evaluation of ESBMC-GPU against other verifiers.}
@@
 \textcolor{blue}{Besides, it also applies MPOR, two-thread analysis, and state hashing in order to further simplify verification models and provides fewer incorrect results, compared with GKLEE, GPUVerify, and CIVL.} Finally, it presents improved ability to detect array out-of-bounds and data race violations.
-Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques \textcolor{blue}{(e.g., Abstract Interpretation~\cite{Rocha2017})} to reduce the number of thread interleavings, by taking into account GPU symmetry.
+Future work aims to support \textcolor{blue}{CUDA (parallel) streams and events~\cite{cuda:2012}} and implement further techniques \textcolor{blue}{(e.g., Abstract Interpretation~\cite{Rocha2017}) in order to prune the state space exploration}, by taking into account GPU symmetry.
 
@@
 \newblock {GCCE} 2015; 179--447.
-%\newblock {DOI=http://dx.doi.org/10.1109/GCCE.2015.7398699}}
 
@@
 \newblock {SPIN} 2016; 97--103.
-%\newblock {DOI=http://dx.doi.org/10.1007/978-3-319-32582-8\_6}}
 

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/source/manuscript.tex
updating Funtionalities section
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 
-Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access \textcolor{blue}{to the same memory locations}; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
+%Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties: data-race conditions, in order to detect if multiple threads perform unsynchronized access \textcolor{blue}{to the same memory locations}; pointer safety, {\it i.e.}, whether {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid; array bounds, in order to ensure that array indices are within known bounds; arithmetic under- and overflow, which happens when a sum or product exceeds the memory limits that a variable can handle; division by zero, which takes place when denominators, in arithmetic expressions, lead to a division by zero; and user-specified assertions, {\it i.e.}, all assertions specified by users, which is essential to a thorough verification process.%, as some specific possible violations must be explicitly pointed out.
+
+%In order to check the aforementioned properties, ESBMC-GPU explicitly explores the possible interleavings (up to the given context bound) and calls the single-threaded BMC procedure on each one, whenever it reaches an RT leaf node.
+Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings \textcolor{blue}{have been systematically explored}~\cite{pereira:2016}.
+
+\textcolor{blue}{Through the integration of COM into ESBMC ({\it i.e.}, ESBMC-GPU), one is able to analyze CUDA programs and verify the following properties:}
+
+\begin{itemize}
+	\item[--] \textcolor{blue}{\textbf{Data race.} ESBMC-GPU checks data race conditions, in order to detect if multiple threads perform unsynchronized access \textcolor{blue}{to the same memory locations};}
+
+	\item[--] \textcolor{blue}{\textbf{Pointer safety.} ESBMC-GPU also ensures that {\it (i)} a pointer offset does not exceed object bounds and {\it (ii)} a pointer is neither {\tt NULL} nor invalid;}
+
+	\item[--] \textcolor{blue}{\textbf{Array bounds.} ESBMC-GPU performs array-bound checking, in order to ensure that any variable, used as an array index, is within known bounds;}
+
+	\item[--] \textcolor{blue}{\textbf{Arithmetic under- and overflow.} ESBMC-GPU checks whether a sum or product exceeds the memory limits that a variable can handle, which can cause an error capable of spreading through the entire execution path;}
+
+	\item[--] \textcolor{blue}{\textbf{Division by zero.} ESBMC-GPU analyzes whether denominators, in arithmetic expressions, lead to a division by zero;}
+
+	\item[--] \textcolor{blue}{\textbf{User-specified assertions.\ } ESBMC-GPU considers all assertions specified by users, which is essential to a thorough verification process, as some specific possible violations must be explicitly pointed out.}
+\end{itemize}
 
@@
 Then, the mentioned procedure will stop if it finds a bug or when all possible RT interleavings \textcolor{blue}{have been systematically explored}~\cite{pereira:2016}.
+\textcolor{blue}{Furthermore, ESBMC-GPU has the following additional command-line options:
+\\
+\texttt{--no-assertions}: to ignore assertions; \\
+\texttt{--no-bounds-check}: to skip array bounds check; \\
+\texttt{--no-div-by-zero-check}: to skip division by zero check; \\
+\texttt{--no-pointer-check}: to skip pointer check; \\
+\texttt{--memory-leak-check}: to enable memory leak check; \\
+\texttt{--overflow-check}: to enable arithmetic over- and underflow check; \\
+\texttt{--deadlock-check}: to enable global and local deadlock check with mutex; \\
+\texttt{--data-races-check}: to enable data races check; \\
+\texttt{--lock-order-check}: to enable for lock acquisition ordering check; \\
+\texttt{--atomicity-check}: to enable atomicity check at visible assignments; \\
+\texttt{--memlimit}: to configure memory limit, of form ``100MB'' or ``2GB''; \\
+\texttt{--timeout}: to configure time limit, integer followed by \{s, m, or h\}; \\
+\texttt{--force-malloc-success}: to consider that there is always enough memory available in the device; \\
+\texttt{-DGPU\_threads=<$number$>}: to define the total number of threads by kernel. \\}
 

### 2017-05-15 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
updating reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: don't we explain that on Functionalities section?
-Any ideia what else we could add?
-The problem is that we have page limit, right?
-Can you please mail the editor and check if we can exceed the page limit to properly handle the reviewer comments?}}
+\textcolor{blue}{\textbf{Response:} We have improved Functionalities section, in order to include more details about ESBMC-GPU's capabilities.}
 \end{quote}

### 2017-05-16 Publications/Original Software Publications/Science of Computer Programming/reply letter/letter.tex
adding comments to reply letter
Signed-off-by: Felipe R. Monteiro <rms.felipe@gmail.com>

@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: Shall we reference your work about verifying multithread ANSI-C programs? Yes, please. Also reference our journal paper about esbmc-gpu.}}
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Erickson: Could you implement this comment, please?} Lucas: Shall we reference your work about verifying multithread ANSI-C programs?
+Yes, please.
+Also reference our journal paper about esbmc-gpu.}
   \end{quote}
@@
   \begin{quote}
-  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Lucas: I didn't get it.
-Did you?
-Can you cite our TACAS 2014 paper?
+  \textcolor{blue}{\textbf{Response:} \textcolor{red}{Erickson: Could you implement this comment, please?} Lucas: Can you cite our TACAS 2014 paper?
 I suggest you to read that paper which briefly explains our memory model.
-Then you can properly reply to the reviewer and cite that paper.}\textcolor{green}{Eddie: Maybe he is talking about shared, register, global and local...
+Then you can properly reply to the reviewer and cite that paper.\textcolor{green}{Eddie: Maybe he is talking about shared, register, global and local...
 We need to show that our memory model correctly represent such types...}}
@@
 \begin{quote}
-\textcolor{blue}{\textbf{Response:}\textcolor{green}{Eddie: We have to comment on kernels, blocks and threads, in order to clarify that, showing that it it feasible ...}}
+\textcolor{blue}{\textbf{Response:}\textcolor{red}{Erickson: Could you implement this comment, please?} \textcolor{green}{Eddie: We have to comment on kernels, blocks and threads, in order to clarify that, showing that it it feasible ...}}
 \end{quote}
